### 2019-10-28 scripts/laggedScatterplot.Rmd
Lagged scatterplot .Rmd

---
title: "NLA17 Dissolved Gas Spatial Autocorrelation"
author: "J. Beaulieu"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

During the 2017 National Lakes Assessment, dissolved gas samples and air samples were collected from the index site at each waterbody.  Samples were analyzed for CO~2~, CH~4~, and N~2~O concentration via gas chromatography.  The following is an analysis of spatial autocorrelation via lagged scatterplots

## Analysis

This .rmd assumes 'DissolvedGasNla.prj' working environment.  Data is available as an sf and sp object.
```{r}
# Unique samples are identified by site.id and visit.no.  Some sites were sampled twice during the NLA17 index period.
# sample.source column discriminates between 'AIR' and dissolved gas 'DG" samples.  "AIR" samples were removed in
# readDissolvedGas.R.  Dissolved gas concentrations are reported in mol/L.

# Inspect sf object
dg.sf[ , c("site.id", "visit.no", "sample.source", "dissolved.co2", "dissolved.ch4", "dissolved.n2o")]

# Inspect sp object
dg.sp[ , c("site.id", "visit.no", "sample.source", "dissolved.co2", "dissolved.ch4", "dissolved.n2o")]
```

Bubble plots are a useful tool for visualizing spatial patterns.  Dissolved CH4 and N2O don't display obvious spatial patterns, but dissolved CO2 seems to greater along east coast, upper midwest, and a SW by NE transect between Texas and Wisconsin.  Pretty interesting.

```{r}
# bubble function requires sp object and can't handle missing values.
bubble(dg.sp[!is.na(dg.sp$dissolved.ch4), "dissolved.ch4"]) #
bubble(dg.sp[!is.na(dg.sp$dissolved.co2), "dissolved.co2"])
bubble(dg.sp[!is.na(dg.sp$dissolved.n2o), "dissolved.n2o"])
```

Lagged scatterplots are a useful tool for detecting spatial autocorrelation.  Evidence for spatial autocorrelation is stronger correltation among close vs. distant pairs of observations.  This pattern is not evident for any of the gases.

```{r}
# Need to select appropriate distance bins.
# What are distance between points in the data?
gdis <- spDists(dg.sp)
max(gdis) # 5617, this must be km.  5617 km = 3490 miles, AK to FL

hscat(dissolved.ch4 ~ 1,  
      dg.sp[!is.na(dg.sp$dissolved.ch4), ], # remove missing
      # 5617275m = 3490 miles. AK to FL
      c(0, 1000, 2000, 3000, 4000, 5000, 6000))

hscat(dissolved.co2 ~ 1,  
      dg.sp[!is.na(dg.sp$dissolved.co2), ], # remove missing
      # 5617275m = 3490 miles. AK to FL
      c(0, 1000, 2000, 3000, 4000, 5000, 6000))

hscat(dissolved.n2o ~ 1,  
      dg.sp[!is.na(dg.sp$dissolved.n2o), ], # remove missing
      # 5617275m = 3490 miles. AK to FL
      c(0, 1000, 2000, 3000, 4000, 5000, 6000))

```

Perhaps spatial autocoreltation is more apparent within particular ecoregions.  Can we do lagged scatterplots by ecoregion?
```{r}

# Loop to create lagged scatterplot by ecoregion x gas.  Not working yet....
# Need to customize size bins per ecoregion, I think.
# for(i in 1:length(unique(dg.sp@data$WSA9_NAME))) {
#   WSA9.i <- unique(dg.sp@data$WSA9_NAME)[i] # unique ecoregion
#   dg.sp.i <- filter(dg.sp, WSA9_NAME == WSA9.i) # subset spatial object for ecoregion of interest
#   gDis.1 <- spDists(dg.sp.i) # matrix of distances
#   gDis.2 <- seq(from = 0, to = max(gDis.1), by = max(gDis.1) / 6) # input to breaks argument in hscat()
#   
#   for(j in 1:3) { # for each gas
#     gas.j <- c("dissolved.ch4", "dissolved.co2", "dissolved.n2o")[j] # gas of interest
#     dg.sp.i.j <- filter(dg.sp.i, !is.na(gas.j)) # remove missing
#     
#     hscat(gas.j ~ 1, # not interpretting 'gas.j' correctly.  Works if "dissolved.ch4' is specified.
#           dg.sp.i, 
#           gDis.2) 
#   }
#   
# }
```