### 2020-10-05 scripts/dataMunge.Rmd
Added code to calculate upper mixed layer DO and temp.

---
title: "Data Munging"
authors: "J. Beaulieu, R. Martin, and M. McManus"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    fig_caption: yes
    depth: 2
    number_sections: true
editor_options: 
  chunk_output_type: console
---
```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
# load libraries
library(plyr) # for ddply, load before tidyverse
library(tidyverse) # dplyr, ggplot
library(forcats) # fct_explicit_na()
library(rLakeAnalyzer) # for buoyancy frequency
library(zoo) # interpolate NA temp/DO

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")
```

# Purpose
The purpose of this .rmd is to prepare the data set for subsequent modeling.  The data will be written as .RData object and stored at the shared documents library associated with this project: https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas

The dissolved gas data were currated in the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-02-24.txt.  Data on waterbody surface area and design weights were provided by Karen Blocksom on 12/13/2019.  These data files are stored in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.

# Data
## Read data
Below we read in the data files.

```{r}
# Read dissolved gas data file
dg <- read.table(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/nla17gasDataAggregated_2020-02-24.txt"),
                 header = TRUE, sep = "\t", as.is = TRUE)

# Read site weights and waterbody area data
wt.area <- read.csv(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/NLA17_SiteWeights_Areas.csv"),
                    as.is = TRUE)

# Read in temp profile data
temp.o2 <- read.csv(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/NLA17_Profile_Data_Updated_11Feb2020.csv"),
                    as.is = TRUE)
```

## Merge dissolved gas with survey weights + area
Unique waterbodies are identified by 'site id'.  Some waterbodies were visited twice, therefore each unique sampling trip is identified by a combination of 'site id' and visit number.

``` {r}
dg %>% distinct(site.id, visit.no)
```

Survey weights and lake area are invariant across site visits, therefore the weights + lake area file does not contain a visit number field:
``` {r}
# no visit number in this file
names(wt.area)
```

The two data sets will be merged on the 'site id' fields.  The dissolved gas file contains data from sites in Alaska that are not present in the weights + lake area file.  Alaska sites are not an official component of the NLA and will removed from this analysis.
``` {r}
# Will merge datasets on SITE_ID field.  Are all site.id values in dg present in wt.area?
dg %>% filter(!(site.id %in% wt.area$SITE_ID)) %>% distinct(site.id) # no.

# Remove AK sites
dg <- dg %>% filter(!(grepl("AK", site.id))) # exlclude alaska
```

There are 23 observations in the weights + lake area file that are absent from the dissolved gas file.  One observation is from Arizona.  This corresponds to notes in the chain of custody form indicating the vials arrived unlabeled and were not run.  The remaining sites are from OR. Karen Blocksom confirmed that these sites were part of the OR state survey, but collected 'most' of the NLA measurements.  They were included in the final NLA dataset for 'the purpose of adjusting weights'.  They did not collect gas samples and can be disregarded for this analysis.

``` {r}
# Any site id values in wt.area, but not present in dg?
wt.area %>% filter(!(SITE_ID %in% dg$site.id)) %>% select(SITE_ID, EVAL_CAT, SITETYPE) # 23 sites w/out dissolved gas data
```

Finally, lets merge these two files on site id.  Use left_join to retain all dissolved gas data, but disregard OR and AZ sites where we don't have gas data.
``` {r}
dim(dg) # 2369, 364

dg <- left_join(dg, wt.area, by = c("site.id" = "SITE_ID"))

dim(dg) # 2369, 369;  good retained all DG observations, added a few columns.
```

## Manipulate
Implement unit conversion and create source/sink column:
``` {r}
dg <- dg %>%
  # unit conversion
  mutate(dissolved.ch4 = dissolved.ch4 * 1000000, # mol/L -> umol/L
         dissolved.co2 = dissolved.co2 * 1000000, # mol/L -> umol/L,
         dissolved.n2o = dissolved.n2o * 1000000000, # mol/L -> nmol/L
         # add source/sink column
         co2.src.snk = ifelse(co2.sat.ratio > 1, "source", "sink"),
         ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"))
```

The survey design is stratified by State.  Within each state, an unequal probability category based on 5 lake size classes is used.  Use lake area data provided by Karen to define lake area categories.
```{r}
dg <- dg %>%
  mutate(
    # add state
    state = substr(site.id, 7, 8),
    size_cat = factor( cut( AREA_HA,
                            breaks = c( -Inf, 4, 10, 20, 50, Inf ),
                            labels = c( "min_4", "4_10", "10_20", "20_50", "50_max") ) )
    )

```
All observations have a value for the size_cat design variable.

``` {r}
table(dg$size_cat, useNA = "ifany")
```

## Calculate buoyancy frequency
Water temperature and dissolved oxygen profiles were measured at the index site in each waterbody.  A variety of metrics describing the degree of thermal stratification can be extracted from these data.  Webb et al. used the Brunt–Väisälä buoyancy frequency as an indicator of thermal stratification strength.  We will do the same here.

Three site visits contain only NA for temperature.  Four site visits contain only NA for DO.  Two of the incidents include notes that sonde was forgotten or malfunctioning. Two of the three sites were revisited later in the study and temperature data were collected.  Lets assign data from the second site visit to the first visit when no temperature data were collected.  This leaves only NLA17_CA-10082, visit 1, without temperature data.  NLA17_ID-10084 has no DO data and was not resampled.  Data sheet comment "Probe not measuring LDO for all depths".

```{r}
# Sites that only contain NAs.
missing <- temp.o2 %>% group_by(SITE_ID, VISIT_NO) %>% 
  summarise(all.temp = all(is.na(TEMPERATURE)),
            all.do = all(is.na(OXYGEN))) %>% 
  filter(all.temp == TRUE | all.do == TRUE) %>%
  select(SITE_ID, VISIT_NO, all.temp, all.do)

# View sites with missing temp/do
missing

# Two of the sites have data from second site visit
temp.o2 %>% filter(SITE_ID %in% missing$SITE_ID) %>%
  select(SITE_ID, VISIT_NO, TEMPERATURE, OXYGEN)

# Extract data from second visit to assign to first visit
replacementData <- temp.o2 %>% filter(SITE_ID %in% missing$SITE_ID, VISIT_NO == 2) %>%
  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>%
  mutate(VISIT_NO = 1) # change visit no to one for join

# Join the data
temp.o21 <- full_join(temp.o2, replacementData)

# Inspect results of join.  Looks good.  
# Only one site missing all temperature and DO data. (NLA17_CA-10082)
# One site missing all DO data (NLA17_ID-10084)
temp.o21 %>% filter(SITE_ID %in% missing$SITE_ID) %>%
  select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>% 
  arrange(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN)

# NLA17_ID-10001 visit 1 only has temp/DO measured at surface depth.  Notes indicate probe malfunction.
# Second site visit shows similar surface DO/temp and no thermal stratification.  Replace visit 1 with visit 2
# data
temp.o21[temp.o21$SITE_ID == "NLA17_ID-10001" & temp.o21$"VISIT_NO" == 1, c("OXYGEN", "TEMPERATURE")] = 
  temp.o21[temp.o21$SITE_ID == "NLA17_ID-10001" & temp.o21$"VISIT_NO" == 2, c("OXYGEN", "TEMPERATURE", "DEPTH")] %>%
  group_by(DEPTH) %>% # Need to aggregate duplicate measurements at depth = 0
  summarise(OXYGEN = mean(OXYGEN), TEMPERATURE = mean(TEMPERATURE)) %>%  # mean by depth
  select(-DEPTH) # remove depth

```

The field crews measured temperature/DO starting near the water surface, then at increasing depths.  A second measurement was made at the water surface before the sonde was removed from the water, therefore the file contains duplicate measurements at the shallowest depths. Duplicate measurements corrupt the buoyancy frequency function and will therefore be aggregated. 
``` {r}
# Aggregate across SITE_ID, VISIT_NO, and DEPTH.  I don't like the dplyr solution for this (group_map) and I'm not familiar with the purr::map_ functions.  I'm going old-school and busting out the trusty plyr::ddply.
temp.o2.2 <- temp.o21 %>% select(SITE_ID, VISIT_NO, DEPTH, TEMPERATURE, OXYGEN) %>%
  plyr::ddply(., .(SITE_ID, VISIT_NO, DEPTH), summarize, 
              TEMPERATURE.M = mean(TEMPERATURE, na.rm = TRUE),
              OXYGEN.M = mean(OXYGEN, na.rm = TRUE))

dim(temp.o21) # 11003
dim(temp.o2.2) # 10133, 870 observations aggregated
```

`r temp.o2.2 %>% filter(is.na(DEPTH)) %>% nrow()` observations report NA for DEPTH, DO, and TEMPERATURE.  These will be removed.

``` {r}
# Missing depth values
# All also have NA for DO and Temp.  Remove from dataset
temp.o2.2 %>% 
  filter(is.na(DEPTH))

dim(temp.o2.2) # 10133, 5
temp.o2.2 <- temp.o2.2 %>% filter(!is.na(DEPTH))
dim(temp.o2.2) # 10085, 5
```

`r temp.o2.2 %>% filter(is.na(TEMPERATURE.M) & is.na(OXYGEN.M)) %>% nrow()` observations report NA for temperature and DO .  These are always at the shallowest (0m) or deepest depths.  When at shallowest, there is a 0.1m measurement.  When at deepest, I'm guessing no measurements were made at that depth.  Remove these observations.

``` {r}
# NA for temp and DO
fewMissing <- temp.o2.2 %>% 
  filter(is.na(TEMPERATURE.M) & is.na(OXYGEN.M)) %>%
  select(SITE_ID, VISIT_NO) %>%
  distinct(SITE_ID, VISIT_NO)

temp.o2.2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% paste0(fewMissing$SITE_ID, fewMissing$VISIT_NO))

dim(temp.o2.2) # 10085, 5
temp.o2.2 <- temp.o2.2 %>% filter(!(is.na(TEMPERATURE.M) & is.na(OXYGEN.M)))
dim(temp.o2.2) # 10081, 5
```

`r temp.o2.2 %>% filter(is.na(TEMPERATURE.M) | is.na(OXYGEN.M)) %>% nrow()` observations report NA for depth or temperature.  If these are fairly random within the depth profile, they can be imputed.  No DO data for NLA17_ID-10002 (see above).  The remaining cases are fairly random.  Will impute via linear interpolation.

```{r}
fewMissing <- temp.o2.2 %>% filter(is.na(TEMPERATURE.M) | is.na(OXYGEN.M)) %>%
  select(SITE_ID, VISIT_NO) %>%
  distinct(SITE_ID, VISIT_NO)

temp.o2.2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% paste0(fewMissing$SITE_ID, fewMissing$VISIT_NO))

temp.o2.2 <- temp.o2.2 %>% 
  group_by(SITE_ID, VISIT_NO) %>%
  mutate(OXYGEN.M = na.approx(OXYGEN.M, DEPTH, na.rm = FALSE, rule = 2), #rule = 2 assigns value closest to boundary 
         TEMPERATURE.M = na.approx(TEMPERATURE.M, DEPTH, na.rm = FALSE, rule = 2)) %>%
  ungroup()

dim(temp.o2.2) #10081, 5
```

Finally, calculate the buoyancy frequency depth profile for each waterbody and print to buoyancyFrequency.pdf for inspection.

``` {r}
# The buoyancy frequency is calculated at multiple depths, but at fewer depths than the where 
# temperature was measured.  The output of the buoyancy frequency function is shorter 
# than the input, which is incompatible with dplyr mutate and summarize.  The dplyr alternative
# is group_map which produces a list, but the grouping variables are stripped from the list
# elements.  I really don't like this.  I think this can be addressed by combining group_map
# with purr::map_ functions, but I'm not familiar with purr.  Rather, I'll use the old school
# split -lapply - do(rbind) approach.

# First, strip out any NA, NaN, Inf...etc

bf <- temp.o2.2 %>%
  filter(is.finite(TEMPERATURE.M)) %>% # This completely removes the CA site with only NA for temp data.
  group_split(SITE_ID, VISIT_NO) %>% # split into list of df
  lapply(function(x) {
    data.frame(BF = buoyancy.freq(wtr = x$TEMPERATURE.M, depths = x$DEPTH), # calculate buoyancy frequency
               SITE_ID = unique(x$SITE_ID), # add site ID
               VISIT_NO = unique(x$VISIT_NO), # add visit no
               stringsAsFactors = FALSE)}) %>% 
  lapply(function(x) x %>% mutate(DEPTHS.BF = attr(BF, "depths"))) %>% # depths stored as attribute, assign to column
  do.call("rbind", .) # collapse to df

# there are two non-finite bf values, lets have a look
bf.na <- bf %>% filter(!is.finite(BF)) %>%
  select(SITE_ID, VISIT_NO) %>%
  mutate(SITE.VISIT = paste0(SITE_ID, VISIT_NO))

bf.na

# These site have a single temp measurement made at shallow depths (<=1m, NLA17_MD-10004, NLA17_ND-10189).
# Presumably these are shallow sites, therefore unlikely to be stratified.  Assign a bf very close to 0 (see below).
bf %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT) 
temp.o2.2 %>% filter(paste0(SITE_ID, VISIT_NO) %in% bf.na$SITE.VISIT)

# plot buoyancy frequency depth profile
# (http://stackoverflow.com/questions/8018961/connect-points-in-qplot-by-adjacent-y-value-not-x-value)
# https://stackoverflow.com/questions/29034863/apply-a-ggplot-function-per-group-with-dplyr-and-set-title-per-group
bf.plot <- bf %>%
  group_by(SITE_ID, VISIT_NO) %>%
  do(plots = ggplot(data = .) +
       aes(DEPTHS.BF, BF) +
    geom_point() +
    geom_line() +
    scale_x_reverse() +
    coord_flip() +
    ggtitle(paste(unique(.$SITE_ID), "VISIT_NO = ", unique(.$VISIT_NO))))

# pdf("output/figures/buoyancyFrequency.pdf", onefile = TRUE)
# bf.plot$plots
# dev.off()
```

Webb et al. used the maximum buoyancy frequency value as an index of thermal stratification.  The maximum value should occur near the thermocline.  Here is an example of a classical pattern:
```{r}
bf.plot$plots[92]
```

Extract the max buoyancy frequency value for each waterbody after converting negative BF values to 0, per Webb.
``` {r}
# Finally, extract max buoyancy for each SITE_ID x VISIT_NO
bf.max <- bf %>% 
  mutate(BF = ifelse(BF < 0,
                     0,
                     BF)) %>%
  group_by(SITE_ID, VISIT_NO) %>%
  summarise(MAX.BF = max(BF, na.rm = TRUE)) %>%
  ungroup()
```

`r bf.max %>% summarize(n.na = sum(!is.finite(bf.max$MAX.BF))) %>% pull()` of the site have a non-finite value for maximum buoyancy frequency value.  These are the two shallow sites with only a single temperature measurement.  Assign these the minimum bf.max value.
``` {r}
bf.max %>% filter(!is.finite(MAX.BF))

min.max.bf <- bf.max %>% 
  filter(is.finite(MAX.BF), MAX.BF > 0) %>%
  summarize(min.max.bf = min(MAX.BF))

bf.max <- bf.max %>%
  mutate(MAX.BF = replace(MAX.BF, 
                          !is.finite(MAX.BF), # if not finite
                          min.max.bf$min.max.bf)) # then min

# that took care of it
bf.max %>% filter(!is.finite(MAX.BF))
```

Finally, merge buoyancy frequency data with dg.  The buoyancy frequency data contains information from the extra Oregon sites (see above) and the AZ site where the gas samples were unlabeled and not analyzed.  This is as expected.
The buoyancy frequency data do not contain observations for the California site (CA-10082) where the sonde was reported missing/malfunctioning.
```{r}
# are all SITE_ID values in buoyancy frequency also present in dg?
bf.max %>% filter(!(SITE_ID %in% dg$site.id)) %>% # no, OR and AZ-10007
print(n= Inf)

# are all site.id values in dg also present in bf.max?
# Just CA site, as expected.
dg %>% filter(!(site.id %in% bf.max$SITE_ID)) %>% # no
  select(site.id, visit.no)

# dim(bf.max) # 1209, 3
# dim(dg) # 2369, 374

# Left join: retain all dg values, exclude non-matching bf values (i.e. OR and AZ)
dg <- left_join(dg, bf.max, by = c("site.id" = "SITE_ID", "visit.no" = "VISIT_NO"))  
# dim(dg) # 2369, 375, good!
```

The California site with the broken sonde has no bf data.  Nothing we can do about it.
```{r}
dg %>% filter(is.na(MAX.BF)) %>%
  select(site.id, visit.no)
```

## Calculate thermocline depth

Dissolved gas concentrations collected near the water surface may be more related to mean conditions in the upper mixed layer than conditions near the air-water interface (e.g. depth = 0.1).  Below we calculate depth to thermocline.
The thermocline data contains information from the extra Oregon sites (see above) and the AZ site where the gas samples were unlabeled and not analyzed.  This is as expected.  No thermocline data for the California site (CA-10082) where the sonde was reported missing/malfunctioning.

``` {r}
therm <- temp.o2.2 %>%
  group_by(SITE_ID, VISIT_NO) %>%
  summarise(thermD = thermo.depth(TEMPERATURE.M, DEPTH))

summary(therm$thermD) # 395 sites with no evidence of stratification
dim(therm) #1209, 3 collapsed temp.o2.2 into one observation per site x visit

# Merge therm depth into temp.o2.2 for calcs below.  A single therm depth repeated for each row in unique site x visit
dim(temp.o2.2) # 10081, 5
dim(therm) #1209, 3
temp.o2.2 <- left_join(temp.o2.2, therm)
dim(temp.o2.2) # 10081, 6
# ggplot(dg, aes(thermD, MAX.BF)) + geom_point() + ylim(0,0.04)
```

## Surface mixed layer DO and temperature

Average all values between surface and thermocline.  The surface mixed layer data contains information from the extra Oregon sites (see above) and the AZ site where the gas samples were unlabeled and not analyzed.  This is as expected.  No mixed layer data for the California site (CA-10082) where the sonde was reported missing/malfunctioning.

```{r}
mixed.layer.do.temp <- temp.o2.2 %>%
  group_split(SITE_ID, VISIT_NO) %>% # split into list of dataframes
  lapply(function(x) { # apply to each element of list
    data.i <- if(any(is.na(x$thermD))) { # if not stratified (thermD is NA or NaN)
      x # then all data
    } else { 
      filter(x, DEPTH <= thermD) # if stratified, then only data at depths <= thermD
    }
    OXYGEN.EPI <- mean(data.i$OXYGEN.M, na.rm = TRUE) # calculate mean DO
    TEMP.EPI <- mean(data.i$TEMPERATURE.M, na.rm = TRUE) # calculate mean Temp
    # dump into dataframe
    data.frame(SITE_ID = unique(x$SITE_ID), # extract site id
               VISIT_NO = unique(x$VISIT_NO), # extract visit no
               OXYGEN.EPI = OXYGEN.EPI,
               TEMP.EPI = TEMP.EPI,
               thermD = unique(x$thermD), # identical values within each df, extract unique
               stringsAsFactors = FALSE)
  }
  ) %>% bind_rows() # collapse to df.  dplyr implementation of do.call(rbind, list)

# are all SITE_ID values in mixed.layer.do.temp also present in dg?
mixed.layer.do.temp %>% filter(!(SITE_ID %in% dg$site.id)) %>% # no, OR and AZ-10007, as expected
distinct(SITE_ID, VISIT_NO)

# are all site.id values in dg also present in mixed.layer.do.temp?
# Just CA site, as expected.
dg %>% filter(!(site.id %in% mixed.layer.do.temp$SITE_ID)) %>% # just CA site, as expected
  select(site.id, visit.no)

# merge with dg
dim(dg) # 2369, 375
dim(mixed.layer.do.temp) #1209, 5
dg <- left_join(dg, mixed.layer.do.temp, by = c("site.id" = "SITE_ID", "visit.no" = "VISIT_NO"))
dim(dg) # 2369, 378
```

## Calculate surface and bottom DO/temp
Webb et al. used the 'surface' and 'bottom' water DO percent saturation and temperature as predictors, but didn't give more details on how these indices were calculated.  We will do something similar and define the 'shallowest' and 'deepest' measurements as 'surface' and 'bottom', respectively, but will use mg.l for dissolved oxygen.
```{r}
# o2 surface (s) and bottom (b)
o2.s.b <- 
temp.o2.2 %>% 
  group_by(SITE_ID, VISIT_NO) %>%
  slice(c(which.min(DEPTH), which.max(DEPTH))) %>% # min and max depth by group
  arrange(SITE_ID, VISIT_NO, DEPTH) %>%
    mutate(DEPTH = ifelse(row_number() == 1,
                        "surface",
                        "bottom")) %>%
  pivot_wider(id_cols = c(SITE_ID, VISIT_NO), 
              names_from = DEPTH, 
              values_from = c(TEMPERATURE.M, OXYGEN.M)) %>%
  rename(temp.surf = TEMPERATURE.M_surface,
         temp.bottom = TEMPERATURE.M_bottom,
         o2.surf = OXYGEN.M_surface,
         o2.bottom = OXYGEN.M_bottom)

o2.s.b

```

These calculations produced 1 site visit with at least one NA for surface/bottom temp/DO.  This is NL17_ID-10084 with no DO data (see above)

```{r}
o2.s.b %>% filter(is.na(temp.surf) | is.na(temp.bottom) | is.na(o2.surf) | is.na(temp.bottom))
```

Merge bottom and surface DO and temperature with dg data.

```{r}
# are all SITE_ID values in o2 and temp data also present in dg?
o2.s.b %>% filter(!(SITE_ID %in% dg$site.id)) %>% # no, OR and AZ-10007, as expected
print(n= Inf)

# are all site.id values in dg also present in o2 and temp data?
# Just CA site, as expected.
dg %>% filter(!(site.id %in% o2.s.b$SITE_ID)) %>% # no
  select(site.id, visit.no)

# dim(o2.s.b) # 1209, 6
# dim(dg) # 2369, 378

# Left join: retain all dg values, exclude non-matching bf values (i.e. OR and AZ)
dg <- left_join(dg, o2.s.b, by = c("site.id" = "SITE_ID", "visit.no" = "VISIT_NO"))  
# dim(dg) # 2369, 382, good!
```

## Fix chemistry
The file contains `r dg %>% dplyr::select(nitrate.n.result) %>% filter(is.na(nitrate.n.result)) %>% summarize(na.no3 = nrow(.)) %>% pull()` missing nitrate values.  Inspection of raw data sheet indicates that:
* North Dakota reported NITRATE_NITRITE_N_RESULT rather than NITRATE.N.RESULT.  Most of the ND sites were 'non detect (ND)' with a detection limit of 0.03 mg /L.
* WI sites were ND with a detection limit of 0.019.

Will replace these values with NITRATE_NITRITE_N_RESULT when available, else 1/2 detection limit.

``` {r}
dg <- dg %>% 
  mutate(nitrate.n.result = replace(nitrate.n.result, 
                                    grepl("WI", site.id) & is.na(nitrate.n.result), # WI na sites
                                    0.019/2), #1/2 detection limit
         # ND sites without nitrate.n and nitrate.nitrite
         nitrate.n.result = replace(nitrate.n.result, 
                                    grepl("ND", site.id) & 
                                      is.na(nitrate.n.result) & 
                                      is.na(nitrate.nitrite.n.result), 
                                    0.03/2), # 1/2 detection limit
# ND sites with a nitrate.nitrite.n.result
         nitrate.n.result = replace(nitrate.n.result, 
                                    grepl("ND", site.id) & 
                                      is.na(nitrate.n.result) & 
                                      !is.na(nitrate.nitrite.n.result), # ND na sites
                                    0.03)) # two sites reporting 0.03 mg/L
```

The file contains `r dg %>% filter(is.na(chla.result.vol)) %>% summarize(na.chl = nrow(.)) %>% pull()` missing chlorophyll values.  Inspection reveals that these are all from North Dakota sites.  It looks like the lab failed to record the volume filtered, so we can't calculate ug/L chl from these sites.  These will remain missing.

## Write out data
Write out final .RData object.
```{r}
save(dg, file = paste0(localPath, 
                       "/Environmental Protection Agency (EPA)/",
                       "ORD NLA17 Dissolved Gas - Documents/",
                       "inputData/dg.", Sys.Date(), ".RData"))
```