### 2020-10-07 scripts/dataMunge.Rmd
Transferred most data manipulation code from dataMunge.Rmd to NLA repo.

---
title: "Data Munging"
authors: "J. Beaulieu, R. Martin, and M. McManus"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    fig_caption: yes
    depth: 2
    number_sections: true
editor_options: 
  chunk_output_type: console
---
```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
# load libraries
library(tidyverse) # dplyr, ggplot

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")
```

# Purpose
The purpose of this .rmd is to prepare the data set for subsequent modeling.  The data will be written as .RData object and stored at the shared documents library associated with this project: https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas

The dissolved gas data were currated in the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-10-07.txt in the documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  When synced to a local computer, the documents library can be read directly with R, after identifying the local directory path.

# Data
## Read data
Below we read in the data files.

```{r}
# Read dissolved gas data file
dg <- read.table(file = paste0(localPath, 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "inputData/nla17gasDataAggregated_2020-10-07.txt"),
                 header = TRUE, sep = "\t", as.is = TRUE) 
```

## Manipulate
Implement unit conversion and create source/sink column:
``` {r}
dg <- dg %>%
  # unit conversion
  mutate(dissolved.ch4 = dissolved.ch4 * 1000000, # mol/L -> umol/L
         dissolved.co2 = dissolved.co2 * 1000000, # mol/L -> umol/L,
         dissolved.n2o = dissolved.n2o * 1000000000, # mol/L -> nmol/L
         # add source/sink column
         co2.src.snk = ifelse(co2.sat.ratio > 1, "source", "sink"),
         ch4.src.snk = ifelse(ch4.sat.ratio > 1, "source", "sink"),
         n2o.src.snk = ifelse(n2o.sat.ratio > 1, "source", "sink"))
```

## Write out data
Write out final .RData object.
```{r}
save(dg, file = paste0(localPath, 
                       "/Environmental Protection Agency (EPA)/",
                       "ORD NLA17 Dissolved Gas - Documents/",
                       "inputData/dg.", Sys.Date(), ".RData"))

```