### 2020-11-30 scripts/CDF_Estimates_Jb.Rmd
Added N2O flux calculation to dataMunge

---
title: "Cumulative Distribution Function (CDF) Estimates for the NLA 2017 Survey"
author: "Tom Kincaid originally, modified by Jake Beaulieu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

## Purpose
Calculate CDF estimates for the NLA 2017 survey.

## History
Created: November 5, 2019

Revised: December 12, 2019 to calculate estimates using the SRS variance
         estimator in addition to the local mean variance estimator

         November 2020, expanded to include estimates of N2O emission rates.

```{r, echo=FALSE}

# Attach the Hmisc and spsurvey packages
library(Hmisc)
library(spsurvey)

# Print the date and session information
cat(date(), "\n\n")
cat("Session Information:\n\n")
sessionInfo()

# Read the file containing data for CDF estimates
#cdf <- read.csv("L:/Priv/ARM Data/NLA 2017/Data Prep/Dissolved Gases/NLA17_DissolvedGases_forPopEst.csv")

# modified file to include N2O emission rates.  See "Users\JBEAULIE\OneDrive - Environmental Protection Agency (EPA)\gitRepository\DissolvedGasNla\scripts\dataMunge.Rmd" for details
# updated file path
cdf <- read.csv(file = paste0(localPath, 
                              "/Environmental Protection Agency (EPA)/",
                              "ORD NLA17 Dissolved Gas - Documents/",
                              "inputData/populationEstimates/",
                              "NLA17_DissolvedGases_forPopEst.JB.csv"))
nr <- nrow(cdf)

# Reorder levels for the lake size class (AREA_CAT6) variable
levels(cdf$AREA_CAT6) <- list(
   "(1,4]" = "(1,4]",
   "(4,10]" = "(4,10]",
   "(10,20]" = "(10,20]",
   "(20,50]" = "(20,50]",
   ">50" = ">50")

# Reorder levels for the lake U.S. Forest Service (FS_EW) variable
levels(cdf$FS_EW) <- list(
   "East" = "East",
   "West" = "West",
   "Not_USFS" = "Not_USFS")

# Print the initial six lines in the cdf data frame
cat("\n\nThe initial six lines in the cdf data frame follow:\n\n")
head(cdf)

# Print the final six lines in the survey cdf data frame
cat("\n\nThe final six lines in the cdf data frame follow:\n\n")
tail(cdf)

# Print summary information for the response variables
cat("\n\nSummary Information for Response Variables\n")
temp <- match(c("CH4_SAT_RATIO", "CO2_SAT_RATIO", "DELTA_13_CH4_C",
   "DELTA_13_CO2_C", "DISSOLVED_CH4", "DISSOLVED_CO2", "DISSOLVED_N2O",
    "N2O_SAT_RATIO", "F_N2O", "F_N2O_D"), names(cdf))
for(c in temp) {
   cat("\n")
   print(describe(cdf[, c], descript=names(cdf)[c]))
}

#
# Calculate CDF estimates using the local mean variance estimator
#

# Create the sites data frame, which identifies sites to use in the analysis
sites <- data.frame(siteID=cdf$SITE_ID,
                    Use=rep(TRUE, nr))

# Create the subpop data frame, which defines populations and subpopulations for
# which estimates are desired
subpop <- data.frame(siteID=cdf$SITE_ID,
                     National=rep("National", nr),
                     LAKE_ORGN=cdf$LAKE_ORGN,
                     AREA_CAT6=cdf$AREA_CAT6,
                     #  > 1 hour with all variables and subpopulations
                     # AG_ECO3_NM=cdf$AG_ECO3_NM,
                      AG_ECO9_NM=cdf$AG_ECO9_NM
                     # EPA_REG=cdf$EPA_REG,
                     # MAJ_BAS_NM=cdf$MAJ_BAS_NM,
                     # HUC2_NM=cdf$HUC2_NM,
                     # OWN_NARS=cdf$OWN_NARS,
                     # FS_EW=cdf$FS_EW
)

# Create the design data frame, which identifies the weight, x-coordinate, and
# y-coordinate for each site ID
design <- data.frame(siteID=cdf$SITE_ID,
                     wgt=cdf$WGT_TP,
                     xcoord=cdf$XCOORD,
                     ycoord=cdf$YCOORD)

# Create the data.cont data frame, which specifies the variables to use in the
# analysis.  > 1 hour with all variables.
data.cont <- data.frame(siteID=cdf$SITE_ID,
                        # CH4_SAT_RATIO=cdf$CH4_SAT_RATIO,
                        # CO2_SAT_RATIO=cdf$CO2_SAT_RATIO,
                        # DELTA_13_CH4_C=cdf$DELTA_13_CH4_C,
                        # DELTA_13_CO2_C=cdf$DELTA_13_CO2_C,
                        # DISSOLVED_CH4=cdf$DISSOLVED_CH4,
                        # DISSOLVED_CO2=cdf$DISSOLVED_CO2,
                        # DISSOLVED_N2O=cdf$DISSOLVED_N2O,
                        # N2O_SAT_RATIO=cdf$N2O_SAT_RATIO,
                        F_N2O = cdf$F_N2O) # diffusive N2O nmol N2O m-2 day-1

```

```{r, echo=FALSE, results="hide"}

# Calculate the estimates
cat("\n\nCalculate CDF estimates using the local mean variance estimator\n")
if(exists("warn.df")) rm("warn.df")
CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont) # < 1 hour with all variables. 1 minute with F_N2O

```

```{r, echo=FALSE}

# Check for warning messages and print them if any exist
if(exists("warn.df")) {
   cat("\nWarning messages generated during the call to cat.analysis:\n")
   warnprnt()
} else {
   cat("\nNo warning messages were generated during the call to cat.analysis.\n")
}

# # Write CDF estimates as a comma-separated value (csv) file
# eval(parse(text = paste0("write.csv(CDF_Estimates$CDF, 'NLA_2017_CDF_Estimates_Local_Mean_",
#    gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))
# 
# # Create a PDF file containing plots of the CDF estimates
# eval(parse(text = paste0("cont.cdfplot('CDF_Estimates_Local_Mean_", gsub("-", "", Sys.Date()), ".pdf', CDF_Estimates$CDF)")))

# Write percentile estimates as a csv file
write.csv(CDF_Estimates$Pct, 
          paste0(localPath, 
                 "/Environmental Protection Agency (EPA)/",
                 "ORD NLA17 Dissolved Gas - Documents/",
                 "inputData/populationEstimates/",
                 "NLA_2017_Percentile_Estimates_Local_Mean_F_N2O_JB",
                 gsub("-", "", Sys.Date()), ".csv"),
          row.names=FALSE)

```

#
# Calculate CDF estimates using the SRS variance estimator
#

```{r, echo=FALSE, results="hide"}

# Calculate the estimates
cat("\n\nCalculate CDF estimates using the SRS variance estimator\n")
if(exists("warn.df")) rm("warn.df")
CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont,
   vartype = "SRS")

```

```{r, echo=FALSE}

# Check for warning messages and print them if any exist
if(exists("warn.df")) {
   cat("\nWarning messages generated during the call to cat.analysis:\n")
   warnprnt()
} else {
   cat("\nNo warning messages were generated during the call to cat.analysis.\n")
}

# Write CDF estimates as a comma-separated value (csv) file
eval(parse(text = paste0("write.csv(CDF_Estimates$CDF, 'NLA_2017_CDF_Estimates_SRS_",
   gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))

# Create a PDF file containing plots of the CDF estimates
eval(parse(text = paste0("cont.cdfplot('CDF_Estimates_SRS_", gsub("-", "", Sys.Date()), ".pdf', CDF_Estimates$CDF)")))

# Write percentile estimates as a csv file
eval(parse(text = paste0("write.csv(CDF_Estimates$Pct, 'NLA_2017_Percentile_Estimates_SRS_",
   gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))

```