### 2021-01-21 scripts/CDF_Estimates_Jb.Rmd

---
title: "Cumulative Distribution Function (CDF) Estimates for the NLA 2017 Survey"
author: "Tom Kincaid originally, modified by Jake Beaulieu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

## Purpose
Calculate CDF estimates for the NLA 2017 survey.

## History
Created: November 5, 2019

Revision history:
December 12, 2019 to calculate estimates using the SRS variance estimator in addition to the local mean variance estimator.

November 2020, expanded to include estimates of N2O emission rates.

December 3, 2020 added extent estimate of N~2~O source/sink

```{r, echo=FALSE}

# Attach the Hmisc and spsurvey packages
library(Hmisc)
library(spsurvey)

# Print the date and session information
cat(date(), "\n\n")
cat("Session Information:\n\n")
sessionInfo()

# Setup file path
localPath <- Sys.getenv("USERPROFILE")

# Read the file containing data for CDF estimates
#cdf <- read.csv("L:/Priv/ARM Data/NLA 2017/Data Prep/Dissolved Gases/NLA17_DissolvedGases_forPopEst.csv")

# modified file to include N2O emission rates.  See "Users\JBEAULIE\OneDrive - Environmental Protection Agency (EPA)\gitRepository\DissolvedGasNla\scripts\dataMunge.Rmd" for details
# updated file path
cdf <- read.csv(file = paste0(localPath, 
                              "/Environmental Protection Agency (EPA)/",
                              "ORD NLA17 Dissolved Gas - Documents/",
                              "inputData/populationEstimates/",
                              "NLA17_DissolvedGases_forPopEst.JB.csv"))
nr <- nrow(cdf)

# Reorder levels for the lake size class (AREA_CAT6) variable
levels(cdf$AREA_CAT6) <- list(
   "(1,4]" = "(1,4]",
   "(4,10]" = "(4,10]",
   "(10,20]" = "(10,20]",
   "(20,50]" = "(20,50]",
   ">50" = ">50")

# Reorder levels for the lake U.S. Forest Service (FS_EW) variable
levels(cdf$FS_EW) <- list(
   "East" = "East",
   "West" = "West",
   "Not_USFS" = "Not_USFS")

# Print the initial six lines in the cdf data frame
cat("\n\nThe initial six lines in the cdf data frame follow:\n\n")
head(cdf)

# Print the final six lines in the survey cdf data frame
cat("\n\nThe final six lines in the cdf data frame follow:\n\n")
tail(cdf)

# Print summary information for the response variables
cat("\n\nSummary Information for Response Variables\n")
temp <- match(c("CH4_SAT_RATIO", "DISSOLVED_CH4", "DELTA_13_CH4_C", "E_CH4_UMOL_D", "F_CH4_KM_D",
                "CO2_SAT_RATIO", "DISSOLVED_CO2", "DELTA_13_CO2_C", "E_CO2_UMOL_D", "F_CO2_KM_D",
                "N2O_SAT_RATIO", "DISSOLVED_N2O", "E_N2O_NMOL_D", "F_N2O_M_D"), names(cdf))
for(c in temp) {
  cat("\n")
  print(describe(cdf[, c], descript=names(cdf)[c]))
}

#
# Calculate CDF estimates using the local mean variance estimator
#

# Create the sites data frame, which identifies sites to use in the analysis
sites <- data.frame(siteID=cdf$SITE_ID,
                    Use=rep(TRUE, nr))

# Create the subpop data frame, which defines populations and subpopulations for
# which estimates are desired
subpop <- data.frame(siteID=cdf$SITE_ID,
                     National=rep("National", nr),
                     LAKE_ORGN=cdf$LAKE_ORGN,
                     AREA_CAT6=cdf$AREA_CAT6,
                     #AG_ECO3_NM=cdf$AG_ECO3_NM,
                     AG_ECO9_NM=cdf$AG_ECO9_NM
                     #EPA_REG=cdf$EPA_REG,
                     #MAJ_BAS_NM=cdf$MAJ_BAS_NM,
                     #HUC2_NM=cdf$HUC2_NM,
                     #OWN_NARS=cdf$OWN_NARS,
                     #FS_EW=cdf$FS_EW
                     )

# Create the design data frame, which identifies the weight, x-coordinate, and
# y-coordinate for each site ID
design <- data.frame(siteID=cdf$SITE_ID,
                     wgt=cdf$WGT_TP,
                     xcoord=cdf$XCOORD,
                     ycoord=cdf$YCOORD)

# Create the data.cont data frame, which specifies the variables to use in the
# analysis.  30 minutes on VM.
data.cont <- data.frame(siteID=cdf$SITE_ID,
                        CH4_SAT_RATIO=cdf$CH4_SAT_RATIO,
                        CO2_SAT_RATIO=cdf$CO2_SAT_RATIO,
                        N2O_SAT_RATIO=cdf$N2O_SAT_RATIO,
                        DELTA_13_CH4_C=cdf$DELTA_13_CH4_C,
                        DELTA_13_CO2_C=cdf$DELTA_13_CO2_C,
                        DISSOLVED_CH4=cdf$DISSOLVED_CH4,
                        DISSOLVED_CO2=cdf$DISSOLVED_CO2,
                        DISSOLVED_N2O=cdf$DISSOLVED_N2O,
                        E_CH4_UMOL_D = cdf$E_CH4_UMOL_D, # umol CH4 m-2 day-1
                        F_CH4_KM_D = cdf$F_CH4_KM_D, # kmol CH4 day-1
                        E_CO2_UMOL_D = cdf$E_CO2_UMOL_D, # umol CO2 m-2 day-1
                        F_CO2_KM_D = cdf$F_CO2_KM_D, # kmol CO2 day-1
                        E_N2O_NMOL_D = cdf$E_N2O_NMOL_D, # nmol N2O m-2 day-1
                        F_N2O_M_D = cdf$F_N2O_M_D) # mol N2O day-1

```

```{r, echo=FALSE, results="hide"}

# Calculate the estimates
cat("\n\nCalculate CDF estimates using the local mean variance estimator\n")
if(exists("warn.df")) rm("warn.df")
time.initial <- Sys.time()
CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont) # 30 minutes on VM
Sys.time() - time.initial
```

```{r, echo=FALSE}

# Check for warning messages and print them if any exist
if(exists("warn.df")) {
   cat("\nWarning messages generated during the call to cont.analysis:\n")
   warnprnt()
} else {
   cat("\nNo warning messages were generated during the call to cont.analysis.\n")
}

# # Write CDF estimates as a comma-separated value (csv) file
# eval(parse(text = paste0("write.csv(CDF_Estimates$CDF, 'NLA_2017_CDF_Estimates_Local_Mean_",
#    gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))
# 
# # Create a PDF file containing plots of the CDF estimates
# eval(parse(text = paste0("cont.cdfplot('CDF_Estimates_Local_Mean_", gsub("-", "", Sys.Date()), ".pdf', CDF_Estimates$CDF)")))

# Write percentile estimates as a csv file
write.csv(CDF_Estimates$Pct, 
          paste0(localPath, 
                 "/Environmental Protection Agency (EPA)/",
                 "ORD NLA17 Dissolved Gas - Documents/",
                 "inputData/populationEstimates/",
                 "NLA_2017_Percentile_Estimates_Local_Mean_JB",
                 gsub("-", "", Sys.Date()), ".csv"),
          row.names=FALSE)

```

# Calculate extent estimate of N2O source/sink
```{r}
# Can use same sites, subpop, and design frames from above. 
# Add N2O source/sink column to dg
cdf$N2O_SRC_SNK = with(cdf, ifelse(N2O_SAT_RATIO > 1, "source", "sink"))
data.cat <-  data.frame(siteID=cdf$SITE_ID,
                        N2O_SRC_SNK = cdf$N2O_SRC_SNK)

Extent_Estimate <- cat.analysis(sites, subpop, design, data.cat)

print(Extent_Estimate)
```

```{r, echo=FALSE}

# Check for warning messages and print them if any exist
if(exists("warn.df")) {
   cat("\nWarning messages generated during the call to cat.analysis:\n")
   warnprnt()
} else {
   cat("\nNo warning messages were generated during the call to cat.analysis.\n")
}

# # Write extent estimates as a comma-separated value (csv) file
write.csv(Extent_Estimate, 
          paste0(localPath, 
                 "/Environmental Protection Agency (EPA)/",
                 "ORD NLA17 Dissolved Gas - Documents/",
                 "inputData/populationEstimates/",
                 "NLA_2017_Extent_Estimates",
                 gsub("-", "", Sys.Date()), ".csv"),
          row.names=FALSE)

```

# Calculate CDF estimates using the SRS variance estimator
#

```{r, echo=FALSE, results="hide"}

# Calculate the estimates
# cat("\n\nCalculate CDF estimates using the SRS variance estimator\n")
# if(exists("warn.df")) rm("warn.df")
# CDF_Estimates <- cont.analysis(sites, subpop, design, data.cont,
#    vartype = "SRS")

```

```{r, echo=FALSE}

# # Check for warning messages and print them if any exist
# if(exists("warn.df")) {
#    cat("\nWarning messages generated during the call to cat.analysis:\n")
#    warnprnt()
# } else {
#    cat("\nNo warning messages were generated during the call to cat.analysis.\n")
# }
# 
# # Write CDF estimates as a comma-separated value (csv) file
# eval(parse(text = paste0("write.csv(CDF_Estimates$CDF, 'NLA_2017_CDF_Estimates_SRS_",
#    gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))
# 
# # Create a PDF file containing plots of the CDF estimates
# eval(parse(text = paste0("cont.cdfplot('CDF_Estimates_SRS_", gsub("-", "", Sys.Date()), ".pdf', CDF_Estimates$CDF)")))
# 
# # Write percentile estimates as a csv file
# eval(parse(text = paste0("write.csv(CDF_Estimates$Pct, 'NLA_2017_Percentile_Estimates_SRS_",
#    gsub("-", "", Sys.Date()), ".csv', row.names=FALSE)")))

```