### 2021-01-21 scripts/DG_sensitivity_to_measurement_error.Rmd

---
title: "Evaluating sensitivity to measurement error in dissolved gas measurements"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  html_notebook: 
    depth: 4
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
#bibliography: RWM_Endnote_Library.bib
link-citations: yes
---

<style>
.vscroll-plot {
    width: 800px;
    height: 600px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

# Background
In this document we explore a range of models for fitting dissolved $\text{N}_2\text{O}$ data observed as part of the 2017 National Lakes Assessment (NLA).

# Setup R
```{r setup_1, include=FALSE, echo=FALSE}
library(knitr)
opts_knit$set(root.dir = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")

#set directories and R package library
setwd("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
.libPaths("C:/Users/rmartin/Desktop/R/library")
```

```{r load_packages, warning=FALSE, message=FALSE}
library(brms)
library(bayesplot)
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(ggExtra)
library(gridExtra)

options(mc.cores = parallel::detectCores())
options( max.print = 2000 )
```

```{r utility_functions, include=FALSE}
# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}

#Calc geometric mean
gm_mean = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(mean(log(x), na.rm = na.rm))
  } else {
    exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

gm_sd = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(sd(log(x), na.rm = na.rm))
  } else {
    exp(sd(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

#calculate skew
MySkew <- function(x) {
     n <- length(x)
     v <- var(x)
     m <- mean(x)
     third.moment <- (1/(n - 2)) * sum((x - m)^3)
     third.moment/(var(x)^(3/2))
 }

#Center and scale by (1) sd
MyStd <- function(x){ (x-mean(x))/(sd(x)*2)}

#First letter in string to uppercase
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x}
```

# Dissolved gas concentration calculation

The full equation for estimating the concentration (nM) of $N_20$ gas dissolved in water is:

$$C = 10^{-6}B \frac{V_{air}(Mr_h - Mr_r)}{(8.3144598 * K * V_{water}) + H^{\theta} \exp( 2700 * ( \frac{1}{K} - \frac{1}{298.15}) * Mr_h}$$

# Set up model for simulating true value

```{r setvars_and_check}
B <- 99 # barometric pressure (kPa)
V_g <- 20 # volume of reference gas in headspace
V_w <- 40 # volume of water below headspace
T_c <- 23 # temperature ( C )
C_e <- 0.055 # gas mixture ratio in headspace
C_a <- 0.091 # gas mixture ratio in reference
H <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant

C <- 10e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )

C <- C * 1e9 #convert mol to nmol

round( C, 1 )

```

# Simulate measurement error for gas measured in water
```{r simulate_me_Ce, message=F}
C_me <- rep( NA, 1e4 )

C_e_me <- rnorm( 1e4, C_e, 0.001 ) #m.e. in measured gas in headspace

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e_me[ i ] - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

qplot( x = C_e_me ) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_e , color = 'red' ) + 
  ggtitle( "Simulated measurement error in water gas concentration") +
  xlab( "concentration (?)")

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error for gas measured in air
```{r simulate_me_Ca, message=F}
C_me <- rep( NA, 1e4 )

C_a_me <- rnorm( 1e4, C_a, 0.001 ) #m.e. in measured gas in air

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

qplot( x = C_a_me ) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_a , color = 'red' ) + 
  ggtitle( "Simulated measurement error in air gas concentration") +
  xlab( "concentration (?)")

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error in $C_e$ and $C_a$
```{r simulate_me_Ca_Ce, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error in water temperature
```{r simulate_me_T_c, message=F}
C_me <- rep( NA, 1e4 )

T_c_me <- rnorm( 1e4, T_c, 0.1 ) #m.e. in measured water temp

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

qplot( x = T_c_me ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = T_c , color = 'red' ) + 
  ggtitle( "Simulated measurement error in temperature") +
  xlab( "temperature (C)")

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error in barometric pressure
```{r simulate_me_B, message=F}
C_me <- rep( NA, 1e4 )

B_me <- rnorm( 1e4, B, 0.25 ) #m.e. in measured water temp

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B_me[ i ] * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

qplot( x = B_me ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = B , color = 'red' ) + 
  ggtitle( "Simulated measurement error in barometric pressure") +
  xlab( "barometric pressure (kPa)")

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error in water volume
```{r simulate_me_V_w, message=F}
C_me <- rep( NA, 1e4 )

V_w_me <- rnorm( 1e4, V_w, 1 ) #m.e. in measured water volume
V_g_me <- 60 - V_w_me

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B * ( V_g_me[ i ] * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

qplot( x = V_w_me ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = V_w , color = 'red' ) + 
  ggtitle( "Simulated measurement error in water volume") +
  xlab( "water vol (mL)")

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

# Simulate measurement error for all observables
```{r simulate_me_all, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 10e-6 * B_me[ i ] * ( V_g_me[ i ] * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

qplot( x = C_me * 1e9 ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "concentration (nMol)")

qplot( x = C_me * 1e9 - C ) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N20 concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```