### 2021-04-05 scripts/Mung_NLA_Sample_Frame.Rmd
Import and munge NLA sample frame

---
title: "Create NLA Sample Frame"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  html_notebook: 
    toc: yes
    toc_depth: 4
    toc_float: true
    code_folding: show
    font_size: 14
    number_sections: true
    theme: simplex
---
# Background
Steps to import, mung, and save NLA sample frame.

# Setup R
```{r load_packages, message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r utility_functions, include=FALSE}
# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}
```

# Mung and save to file
```{r import_sample_frame, echo=TRUE, paged.print=TRUE}
sframe <- read.csv( file = paste0( localPath,
              "/Environmental Protection Agency (EPA)/",
              "ORD NLA17 Dissolved Gas - Documents/",
              "inputData/NLA_Sample_Frame.csv" ), header = T )

sframe <- sframe %>%
  filter(nla17_sf != "Exclude2017") %>%
  filter(nla17_sf != "Exclude2017_Include2017NH") %>%
  mutate(WSA9 = factor(ag_eco9),
          state = factor(state),
          size_cat = factor(area_cat6)) %>% 
  select(WSA9, state, size_cat, area_ha, perim_km) %>% 
  filter(state != "DC") %>%
  filter(state != "HI") %>%
  droplevels() %>%
  mutate(WSA9 = fct_drop(WSA9)) %>% # remove NA level
  #filter( size_cat != "(0,1]" ) %>%
  mutate(size_cat = recode(size_cat, 
                           "(1,4]" = "min_4" ,
                           "(10,20]" = "10_20",
                           "(20,50]" = "20_50",
                           "(4,10]" = "4_10",
                           ">50" = "50_max")) %>%
  mutate(size_cat = factor(size_cat, 
                           levels = c("min_4", "4_10", "10_20", "20_50", "50_max"),
                           ordered = TRUE))

save(sframe, file = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/modelFiles/sframe.rda")

print(sframe)
```