### 2021-06-09 scripts/ASLO_2021.Rmd
adding scripts addressing wind data, population estimates, and ASLO presentation

---
title: "ASLO 2021"
author: "J. Beaulieu"
date: "6/8/2021"
output: html_document
---

```{r setup, include=FALSE}
# load libraries
library(sf) # spatial data
#library(gstat) # lagged scatterplot, variogram
library(tidyverse) # dplyr, ggplot
#library(janitor) # clean names
#library(spdep) # k-NN and Moran plot
#library(ncf) # ncf::correlogram
#library(lattice) # xyplot
library(USAboundaries)

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}
```

## ASLO 2021

This is a convenient place to generate the stats and figures included in the ASLO 2021 presentation.

## Site Map
```{r}
# Load data
 load(paste0(localPath,
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/dg.2021-02-01.RData"))

# To enable mapping the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple")

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# Load state boundaries
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf) +
  geom_sf(data = states, fill = NA, color = "white", size = 0.1) +
  theme_void() +
  theme(legend.title = element_blank())

ggsave("output/figures/siteMap.tiff")

```

## Population Data

```{r}
# about 15 minutes to load first time
# 6/7/21 <2 minutes
Sys.time()
load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
Sys.time()
```

## Explore structure of all_predictions object
```{r}
head(all_predictions)
```
.row uniquely identifies each waterbody.  There are `r all_predictions %>% distinct(.row) %>% nrow()` waterbodies in the sample frame.  `.draw` reflects the simulation number for each waterbody. I think this is a 'draw from the posterior distribution'.  The data contains `r all_predictions %>% distinct(.draw) %>% nrow()` draws for each waterbody.

## N2O saturation ratio: continuous variable
```{r}
# PLOT N2O SAT RATIO BY WSA9

# point and linerange
all_predictions %>%
  group_by(WSA9, .draw) %>%
  #group_by(.draw) %>%
  summarise( mean_sat = mean(preds_sat)) %>%
  summarise( estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw()

```

## N2O saturation ratio: extent estimate
```{r}
# EXTENT OF WATERBODIES FUNCTIONING AS SINKS BY ECOREGION

# density plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  # below summarise generates 500 proportions per WSA
  summarise(prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
  ggplot(aes(x=prop_sat)) + 
  geom_density() +
  facet_wrap(~WSA9) +
  xlab("Proportion of waterbodies functioning as N2O sinks") +
  theme_bw()
ggsave("output/figures/extentDensityPlot.tiff")

# point and line plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/extentPointLinerange.tiff")

# bar and line plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_bar(stat = "identity", aes(fill = ecoregion)) +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()  
ggsave("output/figures/extentBar.tiff")
```