### 2022-12-20 scripts/ASLO_2021_2.Rmd

---
title: "ASLO 2021"
authors: "J. Beaulieu"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    fig_caption: yes
    depth: 2
    number_sections: true
    code_folding:  hide
editor_options: 
  chunk_output_type: console
---

```{r results='hide', message=FALSE, warning=FALSE}
# load libraries
library(sf) # spatial data
library(gstat) # lagged scatterplot, variogram
library(tidyverse) # dplyr, ggplot
library(janitor) # clean names
library(spdep) # k-NN and Moran plot
library(ncf) # ncf::correlogram
library(lattice) # xyplot
library(micromap) # linked micromaps
library(tictoc) # time operations

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}
```

## ASLO 2021

This is a convenient place to generate the stats and figures included in the ASLO 2021 presentation.

## Site Map
```{r}
# Load data.  Loaded object is named dg.
 load(paste0(localPath,
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/dg.2021-02-01.RData"))

# To enable mapping the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple")

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# Load state boundaries
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf) +
  geom_sf(data = states, fill = NA, color = "white", size = 0.1) +
  theme_void() +
  theme(legend.title = element_blank())

ggsave("output/figures/siteMap.tiff")

```

## Population Data
Roy Martin predicted dissolved N2O concentration (preds_n2O), dissolved equilibrium N2O concentration (preds_n2oeq), and N2O saturation ratio (preds_sat) for all waterbodies in the population.  See his .rmd for more details.  
```{r}
# 12/14/22 <2 minutes
tic()
load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
toc()
```

## Explore structure of all_predictions object
.row uniquely identifies each waterbody.  There are `r all_predictions %>% distinct(.row) %>% nrow()` waterbodies in the sample frame.  `.draw` reflects the simulation number for each waterbody. I think this is a 'draw from the posterior distribution'.  The data contains `r all_predictions %>% distinct(.draw) %>% nrow()` draws for each waterbody.
```{r}
head(all_predictions)
```

## N2O saturation ratio: continuous variable
N2O saturation ratio in the sample data has a long right tail, with a maximum value of ~30.  The distribution of the values predicted to the population also has a long right tail, but features more extreme values (>600).
```{r}
# PLOT N2O SAT RATIO BY WSA9

# point and linerange
all_predictions %>%
  group_by(WSA9, .draw) %>%
  #group_by(.draw) %>%
  summarise(mean_sat = mean(n2osat)) %>%
  summarise(estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw() +
  ggtitle("Mean (+/- 95% CI) predicted N2O saturation ratio")

# density plots of predicted distribution
# huge right tail
p <- all_predictions %>%
  #filter(.draw %in% 1:10) %>% # subset for practice
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(ecoregion = fct_reorder(ecoregion, n2osat)) %>%
  ggplot( aes( x = n2osat, group = .draw, color = .draw ) ) +
  geom_density(show.legend = FALSE) +
  geom_vline(xintercept = 1, color = "red") +
  xlab("N2O saturation ratio") +
  xlim(0,5) + # data range is 0-300, but sample max is 30
  facet_wrap(~ecoregion) +
  theme_bw()

#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
g <- ggplot_gtable(ggplot_build(p))
stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
k <- 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid::grid.draw(g)

ggsave("output/figures/n2oSatDensitySmallRange.tiff", g)

summary(dg$n2o.sat.ratio) # max value in sample is 33

# rplot, but show full x-axis range
# include RUG to better show extreme values
p <- all_predictions %>%
  filter(.draw == 1) %>% # need only one draw, otherwise RUG is misleading
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
  ggplot( aes( x = preds_sat ) ) +
  geom_density(color = "blue") +
  geom_vline(xintercept = 1, color = "red") +
  geom_rug() +
  xlab("N2O saturation ratio") +
  #xlim(0,35) + # data range is 0-600, but sample max is 30
  facet_wrap(~ecoregion, scales = "free") +
  theme_bw()

#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
g <- ggplot_gtable(ggplot_build(p))
stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
k <- 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid::grid.draw(g)

ggsave("output/figures/n2oSatDensityFullRange.tiff", g)

all_predictions %>%
    filter(preds_sat > 40, .draw == 1) %>% # look at one draw
  {table(.$WSA9)} # >40 is uncommon outside of CPL

# # density plots of observed distribution
# huge right tail
dg %>%
  mutate(ecoregion = factor(ag.eco9.nm)) %>%
  mutate(ecoregion = fct_reorder(ecoregion, n2o.sat.ratio)) %>%
  ggplot( aes( x = n2o.sat.ratio ) ) +
  geom_density() +
  xlab("mean N2O saturation ratio") +
  facet_wrap(~ecoregion) +
  theme_bw()

```

## N2O saturation ratio: extent estimate
```{r}
# EXTENT OF WATERBODIES FUNCTIONING AS SINKS BY ECOREGION

# density plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  # below summarise generates 500 proportions per WSA
  summarise(prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
  ggplot(aes(x=prop_sat)) + 
  geom_density() +
  facet_wrap(~WSA9) +
  xlab("Proportion of waterbodies functioning as N2O sinks") +
  theme_bw()
ggsave("output/figures/extentDensityPlot.tiff")

# point and line plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/extentPointLinerange.tiff")

# bar and line plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(preds_sat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_bar(stat = "identity", aes(fill = ecoregion)) +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()  
ggsave("output/figures/extentBar.tiff")
```

# N2O saturation ratio: continuous and extent micromap
```{r}
# summarize means by WSA9 (posterior median for mean and 95% credible intervals)
# this is analogous to spsurvey "population estimates"
ecoreg_stats <- all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise( mean_n2o = mean(preds_n2o),
             mean_n2oeq = mean(preds_n2oeq),
             mean_sat = mean(preds_sat),
             prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
  summarise( post_med_n2o = round(mean(mean_n2o), 1),
             LCI_n2o = round(quantile(mean_n2o, probs = 0.025), 1),
             UCI_n2o = round(quantile(mean_n2o, probs = 0.975), 1),
             post_med_n2oeq = round(mean(mean_n2oeq), 2),
             LCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.025), 2),
             UCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.975), 2),
             post_med_sat = round(mean(mean_sat), 2),
             LCI_sat = round(quantile(mean_sat, probs = 0.025), 2),
             UCI_sat = round(quantile(mean_sat, probs = 0.975), 2),
             estimate = round(mean(prop_sat), 3),
             LCL = round(quantile(prop_sat, probs = 0.025), 3),
             UCL = round(quantile(prop_sat, probs = 0.975), 3),
             n = dplyr::n())

ecoreg_stats <- as.data.frame(ecoreg_stats)
class(ecoreg_stats)
str(ecoreg_stats) # WSA9 is a factor, but some of the stats have attribute of Named num

ecoreg_stats$WSA9 <- as.character(ecoreg_stats$WSA9)
str(ecoreg_stats)

ecoreg_stats

ecoregions <- st_read(paste0(localPath, "\\Environmental Protection Agency (EPA)\\",
                             "ORD NLA17 Dissolved Gas - Documents\\inputData\\aggr_ecoregions_simple.shp"))

wsa9names <- st_set_geometry(ecoregions, NULL) %>%
  select(WSA9, WSA9_NAME)

ecoreg_stats <- left_join(ecoreg_stats, wsa9names)

ecoregions <- st_transform(ecoregions, 5070)
# make spatial polygon dataframe need for micromap
ecoregions_sp1 <- as(ecoregions, "Spatial")

# convert to map_table for linked micromap
ecor.polys <- create_map_table(ecoregions_sp1, 'WSA9')
class(ecor.polys)
str(ecor.polys, max =2) # ID is chr "CPL"...

# Full code:  Use this to produce final linked micromaps of the variables you want
# export as pdf US letter landscape from Plots pane
# export as PNG Width 900 Height 968, maintain aspect ratio

mmplot(stat.data=ecoreg_stats,
       map.data=ecor.polys,
       panel.types=c('dot_legend', 'labels', 'dot_cl', 'dot_cl', 'map'),
       panel.data=list(
         NA,
         'WSA9_NAME',
         list('estimate', 'LCL', 'UCL'),
         list('post_med_sat', 'LCI_sat', 'UCI_sat'),
         NA),
       ord.by = 'estimate', rev.ord = TRUE, grouping=3,
       median.row=FALSE,
       map.link=c('WSA9', 'ID'),

       plot.height=5, # what does this do?
       plot.width=7, #what does this do? 
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
       map.color2='lightgray',

       panel.att=list(list(1, point.type=20, point.size=2,
                           graph.border.color='white',
                           xaxis.text.display=F, xaxis.line.display=F,
                           graph.grid.major=F, 
                           panel.width = 2), # added this argument to ensure the full point is shown

                      list(2,  panel.width=2,#increased panel.width to 1.1 from .8, removed header='NARS Reporting\n Regions',
                           align='left', text.size=0.8),#increased text.size to .8 from .6

                      list(3, header= "Proportion undersaturated\n with 95% CI", 
                           graph.bgcolor='lightgray',
                           #xaxis.ticks=list(6.0, 7.0, 8.0, 9.0, 10.0),
                           #xaxis.labels=list(6.0, 7.0, 8.0, 9.0, 10.0),
                           xaxis.title='Proportion undersaturated', panel.width=2.3),

                      list(4, header="N2O Saturation\n with 95% CI", #N2O Saturation\n with 95% CI
                           graph.bgcolor='lightgray',
                           xaxis.ticks=list(0.8, 1.0, 1.2, 1.4, 1.6),
                           xaxis.labels=list(0.8, 1.0, 1.2, 1.4, 1.6),
                           xaxis.title='N2O Saturation, %',panel.width=2.3),

                      list(5, header='Light Gray Means\n Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                           panel.width=2)),

       print.file = "output/figures/mmapSatExtent.tiff", print.res = 300)

```

# N2O emission rate and flux
Roy predicted N2O equilibrium and concentration to the population.  I used ERA5 to calculate mean wind speed during the index period for each lake in the population.  I need to match wind speed to Roy's predictions.  Waterbodies in Roy's predictions are identified by '.row'.  There are `r all_predictions %>% distinct(.row) %>% summarize(population = n()) %>% pull()` rows.

```{r}

# Read in sample frame
sampFrm <- read.csv(file = paste0(Sys.getenv("USERPROFILE"), 
                                    "/Environmental Protection Agency (EPA)/",
                                    "ORD NLA17 Dissolved Gas - Documents/",
                                    "inputData/NLA_Sample_Frame.csv"), header = TRUE,
                    na.strings = "") %>%
  toEPA()

dim(sampFrm) # 614730, 71

sampFrm <- sampFrm %>%
  filter(nla17.sf != "Exclude2017") %>%
  filter(nla17.sf != "Exclude2017_Include2017NH") %>%
  filter(state != "DC") %>%
  filter(state != "HI")

dim(sampFrm)

# print first 6 lines
sampFrm %>% select(ag.eco9, area.cat6, area.ha) %>% head()

# print first 6 .row values of all_predictions
all_predictions %>%
  filter(.row %in% 1:6) %>%
  group_by(.row) %>%
  slice(1)

# OK, .row values in all_predictions match with row number in sampFrm.
# This same order was preserved in wind data.

```