### 2023-02-08 scripts/population.Rmd

---
title: "Explore model results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load the packages
library(tidyverse)
library(ggplot2)

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")
```

## Population Data

```{r}
# about 15 minutes to load
Sys.time()
load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
Sys.time()
```

## Exploratory analysis
```{r}

# summarize means by WSA9 (posterior median for mean and 95% credible intervals)
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise( mean_n2o = mean(preds_n2o)) %>%
  summarise( posterior_median = round(median(mean_n2o), 1),
    LCI = round(quantile(mean_n2o, probs = 0.025), 1),
    UCI = round(quantile(mean_n2o, probs = 0.975), 1))

# plot N2O conc means by wsa9
all_predictions %>%
  group_by(WSA9, .draw) %>%
  #group_by(.draw) %>%
  summarise( mean_n2o = mean(preds_n2o)) %>%
  summarise( estimate = round(median(mean_n2o), 1),
    LCL = round(quantile(mean_n2o, probs = 0.025), 1),
    UCL = round(quantile(mean_n2o, probs = 0.975), 1)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL, type) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  #scale_y_continuous(limits = c(0.8, 2), breaks = seq(0.8, 2, 0.1)) +
  ylab("mean N2O concentration") +
  coord_flip() + 
  theme_bw()

# plot N2O sat ratio means by WSA9
all_predictions %>%
  group_by(WSA9, .draw) %>%
  #group_by(.draw) %>%
  summarise( mean_sat = mean(preds_sat)) %>%
  summarise( estimate = round(median(mean_sat), 3),
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw()

# plot proportion waterbodies functioning as N2O sinks by WSA9
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(preds_sat < 1) / length(unique(.row))) %>%
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()

```