### 2023-06-21 scripts/dgIndicatorAnalysis.Rmd
manuscript progress

---
title: "Analysis of NLA17 Dissolved Gas Indicator"
authors: "J. Beaulieu, R. Martin, and M. McManus"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    fig_caption: yes
    depth: 2
    number_sections: true
    code_folding:  hide
editor_options: 
  chunk_output_type: console
---

```{r results='hide', message=FALSE, warning=FALSE}
# load libraries
library(sf) # spatial data
library(gstat) # lagged scatterplot, variogram
library(tidyverse) # dplyr, ggplot
library(janitor) # clean names
library(spdep) # k-NN and Moran plot
library(ncf) # ncf::correlogram
library(lattice) # xyplot
library(micromap) # linked micromaps
library(tictoc)
library(spmodel) # esv function

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}
```

# Background
During the 2017 National Lakes Assesment, duplicate dissolved gas samples were collected at a depth of ~0.1m at the index site at each waterbody.  Gas samples were analyzed for CO~2~, CH~4~, and N~2~O concentration via gas chromatography and $\delta^{13}CO_{2}$ and $\delta^{13}CH_{4}$ via cavity ring-down spectroscopy.  The following analysis pertains to the concentration data.

## Getting data and filtering
The dissolved gas data were currated the 'NLA' dissolved gas project in RStudio.  The code can be found at a private github repository (https://github.com/USEPA/NLA).  After aggregating across duplicate samples, the data were written to nla17gasDataAggregated_2020-02-24.txt.  A copy of this data file is stored in documents library associated with the 'ORD NLA17 Dissolved Gas' Private Group on SharePoint (https://usepa.sharepoint.com/sites/ORD_NLA17_Dissolved_Gas).  This file was futher manipulated in "C://Users//JBEAULIE//OneDrive-Environmental Protection Agency (EPA)//gitRepository//DissolvedGasNla//scripts//dataMunge.Rmd" and written out as an .RData object, which is imported below.

```{r}
# Load data
 load(paste0(localPath,
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/dg.2021-02-01.RData"))

```

The data file contains `r dg %>% distinct(site.id, visit.no) %>% summarize(n = n()) %>% pull()` unique sampling trips defined by unique combinations of 'site.id' and 'visit.no'.  There were `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` waterbodies sampled, `r dg %>% filter(visit.no == 2) %>% summarize(n = n()) %>% pull()` of which were sampled twice. The primary response variables are the concentrations of CO~2~, CH~4~, and N~2~O dissolved in the surface water of each waterbody.  Concentration is expressed in both mass based units (umol/L, nmol/L: dissolved.co2, dissolved.ch4, dissolved.n2o) and as a saturation ratio (co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio).  Saturation ratio is defined as the ratio of the measured concentration to that expected if the waterbody were in equilibrium with the atmosphere.  A value of 1 indicates equilibrium, values < 1 indicate undersaturation and values > 1 indicate supersaturation.  Undersaturated waterbodies function as a sink for the atmospheric gas whereas supersaturated waterbodies function as a source.

```{r}
# These are the main response variables.
dg %>% 
  select(dissolved.co2.umol, dissolved.ch4.umol, dissolved.n2o.nmol, co2.sat.ratio, ch4.sat.ratio, n2o.sat.ratio) %>%
  {summary(.)}

names(dg)

dg %>% 
  ggplot(aes(as.factor(x = ag.eco9.nm), y = log(dissolved.n2o.nmol))) + geom_boxplot()

dg %>% 
  ggplot(aes(as.factor(x = ag.eco9.nm), y = (n2o.sat.ratio))) + geom_boxplot()

```

#  Exploratory spatial analysis
## Population estimates by ecoregion
### Dissolved gas concentration
The survey design allows for population estimates of central tendency and variance by ecoregion (and other subpopulations), which can provide insight into spatial patterning.  There is broad overlap in mean +/- 95% CI estimates for dissolved CH~4~ and N~2~O concentration across ecoregions.  The Northern Plains, Temperate Plains, and Western Mountains have low dissolved CO~2~ concentration.

```{r, fig.cap= "Fig. caption: Mean +/- 95% CI for the dissolved gas concentration indicator"}
# Read population estimates calculated with local neighborhood variance estimate.  Code provided by Tom Kincaid 
# (12/12/2019), then modified to accomdate the gas emission rate and flux variables.  See CDF_Estimates_Jb.Rmd.
pop <- read.csv(file = paste0(localPath, 
                              "/Environmental Protection Agency (EPA)/",
                              "ORD NLA17 Dissolved Gas - Documents/",
                              "inputData/populationEstimates/",
                              "NLA_2017_Percentile_Estimates_Local_Mean_JB20201202.csv"),
                as.is = TRUE) %>%
  toEPA()  # enforce formatting convention

# Convert units for dissolved gas concentration indicator
pop <- pop %>%
  mutate(estimate = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
                            estimate * 1000000, # mol/L -> umol/L for co2 and ch4
                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
                            estimate * 1000000000, # mol/L -> nmol/L for n2o
                            estimate)), # all others
         lcb95pct = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
                            lcb95pct * 1000000, # mol/L -> umol/L for co2 and ch4
                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
                            lcb95pct * 1000000000, # mol/L -> nmol/L for n2o
                            lcb95pct)), # all others
         ucb95pct = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
                            ucb95pct * 1000000, # mol/L -> umol/L for co2 and ch4
                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
                            ucb95pct * 1000000000, # mol/L -> nmol/L for n2o
                            ucb95pct))) # all others

# extract population estimates for ecoregion and nation
pop.eco <- pop %>% 
  filter(type %in% c("AG_ECO9_NM", "National"), # extract estimates for ecoregion and nation
         statistic == "Mean") # extract mean estimate

# extract dissolved gas indicator, order by dissolved CH4
pop.eco.diss <- pop.eco %>%
  filter(grepl("dissolved", indicator, ignore.case = TRUE)) %>%
  mutate(subpopulation = factor(subpopulation, 
                             levels = filter(pop.eco, indicator == "DISSOLVED_CH4", subpopulation != "National") %>%
                               arrange(estimate) %>%
                               select(subpopulation) %>%
                               pull() %>%
                               c("National", .)))

# look at dissolved concentration indicator
ggplot(pop.eco.diss, aes(estimate, subpopulation)) +
  geom_point() +
  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct)) +
  facet_wrap(~indicator, scales = "free_x") +
  theme(axis.title = element_blank())

# prep object for linked micromap
pop.eco_stats <- pop.eco.diss %>%
  mutate(subpopulation = as.character(subpopulation)) %>%
  rename(WSA9_NAME = subpopulation)

# read in ecoregion polygons
ecoregions <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple")

wsa9names <- st_set_geometry(ecoregions, NULL) %>%
  select(WSA9, WSA9_NAME)

pop.eco_stats <- left_join(pop.eco_stats, wsa9names)

ecoregions <- st_transform(ecoregions, 5070)
# make spatial polygon dataframe need for micromap
ecoregions_sp1 <- as(ecoregions, "Spatial")

# convert to map_table for linked micromap
ecor.polys <- create_map_table(ecoregions_sp1, 'WSA9')
class(ecor.polys)
str(ecor.polys, max =2) # ID is chr "CPL"...

# Full code:  Use this to produce final linked micromaps of the variables you want
# export as pdf US letter landscape from Plots pane
# export as PNG Width 900 Height 968, maintain aspect ratio

mmplot(stat.data=pop.eco_stats %>% 
         filter(WSA9_NAME != "National",
                indicator == "DISSOLVED_CH4"),
       map.data=ecor.polys,
       panel.types=c('dot_legend', 'labels', 'dot_cl', 'map'),
       panel.data=list(
         NA,
         'WSA9_NAME',
         list('estimate', 'lcb95pct', 'ucb95pct'),
         NA),
       ord.by = 'estimate', rev.ord = TRUE, grouping=3,
       median.row=FALSE,
       map.link=c('WSA9', 'ID'),

       plot.height=5, # what does this do?
       plot.width=5.5, #what does this do? 
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
       map.color2='lightgray',

       panel.att=list(list(1, point.type=20, point.size=2,
                           graph.border.color='white',
                           xaxis.text.display=F, xaxis.line.display=F,
                           graph.grid.major=F, 
                           panel.width = 2), # added this argument to ensure the full point is shown

                      list(2,  panel.width=2,#increased panel.width to 1.1 from .8, removed header='NARS Reporting\n Regions',
                           align='left', text.size=0.8),#increased text.size to .8 from .6

                      list(3, header= "CH4 concentration\n with 95% CI", 
                           graph.bgcolor='lightgray',
                           #xaxis.ticks=list(6.0, 7.0, 8.0, 9.0, 10.0),
                           #xaxis.labels=list(6.0, 7.0, 8.0, 9.0, 10.0),
                           xaxis.title='CH4 concentration (uM)', panel.width=2.3),

                      list(4, header='Light Gray Means\n Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                           panel.width=2)),

       print.file = "output/figures/mmapCh4concentration.tiff", print.res = 300)

```

### Dissolved gas saturation ratios and source/sink status
Dissolved gas saturation ratios are calculated as the ratio of the observed to equilibrium concentration.  The equilibrium concentration (sometimes called the saturation concentration) is the concentration expected if the waterbody were in equilibrium with the atmosphere.  It is calculated from the partial pressure of the gas in the atmosphere, water temperature, and the gas specific solubility.  Saturation ratios >1 indicate the water body is supersaturated and is a source of the gas to the atmosphere.  Saturation ratios <1 indicate that the waterbody is undersaturated and is consuming the gas from the atmosphere.  The saturation ratio can therefore be used to categorize each waterbody as a 'sink' or a 'source'.

``` {r}
# read in source/sink population data
# Read extent estimates generated with CDF_Estimates_Jb.Rmd.  File provided by Tom Kincaid (12/12/2019), then modified to accomodate the gas emission rate, flux variables, and cat.analysis for N2O source/sink.  
n2o.extent <- read.csv(file = paste0(localPath, 
                                     "/Environmental Protection Agency (EPA)/",
                                     "ORD NLA17 Dissolved Gas - Documents/",
                                     "inputData/populationEstimates/",
                                     "NLA_2017_Extent_Estimates20201203.csv"),
                       as.is = TRUE) %>%
  toEPA() %>% # enforce formatting convention
  filter(type %in% c("AG_ECO9_NM", "National"), # extract estimates for ecoregion and nation
         category == "sink") %>% # extract extent estimate for sink
  mutate(subpopulation = factor(subpopulation, # arrange by extent estimate for sink
                                levels = filter(., subpopulation != "National") %>%
                                  arrange(estimate.p) %>%
                                  select(subpopulation) %>%
                                  pull() %>%
                                  c("National", .)))
```

The binary 'source' or 'sink' values can be analyzed at the population level using the cat.analysis function in the spsurvey package. The function provides an 'extent' estimate, which is the proportion of waterbodies in each region that are functioning as a source or sink. Over half of the waterbodies are functioning as N2O sinks in all ecoregions.  At the national scale, `r n2o.extent %>% filter(type == "National") %>% select(estimate.p) %>% pull() %>% round(., 1)`% (95% CI: `r n2o.extent %>% filter(type == "National") %>% select(lcb95pct.p) %>% pull() %>% round(., 1)` - `r n2o.extent %>% filter(type == "National") %>% select(ucb95pct.p) %>% pull() %>% round(., 1)`) of lakes and reservoirs were net N~2~O sinks during the index period, suggesting that US lakes and reservoirs are a net N~2~O sink.  However, the 95% confidence intervals of the mean N~2~O saturation ratio include 1 for all ecoregions except the Northern plains which are undersaturated.  At the national scale, the mean N~2~O saturation ratio indicates supersaturation (`r pop.eco %>% filter(type == "National", statistic == "Mean", indicator == "N2O_SAT_RATIO") %>% select(estimate) %>% pull() %>% round(., 2)`% (95% CI: `r pop.eco %>% filter(type == "National", statistic == "Mean", indicator == "N2O_SAT_RATIO") %>% select(lcb95pct) %>% pull() %>% round(., 2)` - `r pop.eco %>% filter(type == "National", statistic == "Mean", indicator == "N2O_SAT_RATIO") %>% select(ucb95pct) %>% pull() %>% round(., 2)`)), suggesting that US lakes and reservoirs function as a net N~2~O sink.  These findings indicate that high levels of N~2~O saturation in a few waterbodies offset undersaturation in most waterbodies, yielding a mean saturation ratio > 1.

```{r message=FALSE}

## Bubble plot of observed N2O saturation index----------------
# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple")

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf, aes(size = n2o.sat.ratio, color = n2o.src.snk),
          show.legend = "point") +
  ggtitle("N2O saturation ratio") +
  scale_color_manual(values = c("white", "black"), name = "source/sink") +
  scale_size_binned(name = "N2O saturation ratio",
             range = c(1, 8), # custom size range
             breaks = c(0.1, 0.5, 1, 2))  # custom breaks

## horizontal bar chart of N2O sink extent---------------
ggplot(n2o.extent, aes(estimate.p, subpopulation)) +
  geom_bar(stat = "identity", aes(fill = subpopulation, color = subpopulation), alpha = 0.25) +
  geom_errorbarh(aes(xmax = ucb95pct.p, xmin = lcb95pct.p, color = subpopulation), height = 0, size = 1) +
  xlab("Percent of waterbodies functioning as N2O sinks") +
  theme(axis.title.y = element_blank())

## horizontal bar chart of N2O saturation ratio-------------
# extract saturation indicator, order by n2o extent (see above).  This enforces same order for both plots.
pop.eco.sat <- pop.eco %>%
  filter(grepl("sat", indicator, ignore.case = TRUE)) %>%
  mutate(subpopulation = factor(subpopulation, 
                             levels = filter(n2o.extent, subpopulation != "National") %>%
                                  arrange(estimate.p) %>%
                                  select(subpopulation) %>%
                                  pull() %>%
                               as.character() %>%
                                  c("National", .)))

# N2O saturation ratio; horizontal bar chart
ggplot(filter(pop.eco.sat, indicator == "N2O_SAT_RATIO"), aes(estimate, subpopulation)) +
  geom_bar(stat = "identity", aes(fill = subpopulation, color = subpopulation), alpha = 0.25) +
  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct, color = subpopulation), height = 0, size = 1) +
  xlab("N2O saturation ratio") +
  theme(axis.title.y = element_blank()) +
  geom_vline(xintercept = 1)

```

### Emission rates and flux
The emission rate (E~gas~) is the rate at which a gas crosses the air water interface and is expressed on an areal basis (i.e. mass m^-2^ day^-1^).  It is calculated as a function of the dissolved gas concentration and gas transfer velocity (see dataMunge.Rmd).  The 95% confidence interval of the mean N~2~O emission rate is quite large in several ecoregions, encompassing 0 in all but the Northen Plains and Xeric ecoregions.  The mean N~2~O emission rate at the national scale is positive (`r  pop.eco %>% filter(indicator == "E_N2O_NMOL_D", subpopulation == "National", statistic == "Mean") %>% select(estimate) %>% pull() %>% round(., 1)/1000` umol N~2~O m^-2^ day^-1^), although the 95% confidence interval includes 0, similar to the N~2~O saturation ratio pattern.

```{r, fig.cap= "Fig. caption: Mean +/- 95% CI for the emission rate (mass of GHG time-1 area-1)"}
# extract population estimates for ecoregion and nation
pop.eco.emission <- pop.eco %>% 
  filter(grepl("^E", indicator)) %>% # ^ means starts with.  Extract the emission rates 
  # order by N2O
   mutate(subpopulation = factor(subpopulation, 
                             levels = filter(., indicator == "E_N2O_NMOL_D", subpopulation != "National") %>%
                               arrange(estimate) %>%
                               select(subpopulation) %>%
                               pull() %>%
                               c("National", .)))

# look at dissolved concentration indicator
ggplot(filter(pop.eco.emission, indicator == "E_N2O_NMOL_D"), 
       aes(estimate/1000, subpopulation)) + # convert to umol
  geom_point(aes(color = subpopulation, fill = subpopulation), alpha = 0.25, size = 5) +
  geom_errorbarh(aes(xmax = ucb95pct/1000, xmin = lcb95pct/1000, color = subpopulation), height = 0, size = 1) + # convert to umol
  geom_vline(xintercept = 0) +
  xlab("N2O emission rate (umol N2O m-2 day-1") + # converted from nmol to umol (see above)
  theme(axis.title.y = element_blank())

```

The N~2~O flux at a waterbody is the total mass of N~2~O crossing the air-water interface per unit time (e.g. mass N~2~O day^-1^) and is calculated as the product of the emission rate (mass N~2~O m^-2^ day^-1^) and waterbody area (m^2^).  Negative values indicate N2O invasion into the waterbody and positive values indicate N~2~O efflux from the waterbody to the atmosphere.  An assessment of the role of lakes and reservoirs in the atmospheric N~2~O budget must be based on flux, rather than emission rates, because flux accounts for waterbody size.  For example, N~2~O emissions from several small waterbodies could be completely offset by N~2~O uptake by a single large waterbody with a small negative emission rate (e.g. N~2~O sink), yet the mean emission rate for these waterbodies would be > 0.

The mean N~2~O flux per waterbody is close to zero for most ecoregions and the 95% confidence interval includes zero for all ecoregions.  The national scale mean N~2~O flux is slightly negative (`r  pop.eco %>% filter(indicator == "F_N2O_M_D", subpopulation == "National", statistic == "Mean") %>% select(estimate) %>% pull() %>% round(., 2)` mol N~2~O day^-1^), but the 95% CI includes 0.

```{r, fig.cap= "Fig. caption: Mean +/- 95% CI of the GHG flux across the air-water interface (mass GHG / time)"}
# extract population estimates for ecoregion and nation
pop.eco.flux <- pop.eco %>% 
  filter(grepl("^F", indicator)) %>% # ^ means starts with.  Extract the flux rates 
  # order by N2O
   mutate(subpopulation = factor(subpopulation, 
                             levels = filter(., indicator == "F_N2O_M_D", subpopulation != "National") %>%
                               arrange(estimate) %>%
                               select(subpopulation) %>%
                               pull() %>%
                               c("National", .)))

# look at dissolved concentration indicator
ggplot(filter(pop.eco.flux, indicator == "F_N2O_M_D"),
       aes(estimate, subpopulation)) +
  geom_point(aes(color = subpopulation, fill = subpopulation), alpha = 0.25, size = 5) +
  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct, color = subpopulation), height = 0, size = 1) +
  geom_vline(xintercept = 0) +
  xlab("N2O flux (mol N2O day^-1 waterbody^-1") +
  theme(axis.title.y = element_blank())

```

### 13C paper
```{r}
# 13C; horizontal bar chart by ecoregion
pop.eco.13c <- filter(pop.eco, indicator == "DELTA_13_CH4_C")
ggplot(pop.eco.13c, aes(estimate, subpopulation)) +
  geom_bar(stat = "identity", aes(fill = subpopulation, color = subpopulation), alpha = 0.25) +
  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct, color = subpopulation), height = 0, size = 1) +
  theme(axis.title.y = element_blank())

# 13C; horizontal bar chart by origin
pop.origin.13c <- filter(pop, type == "LAKE_ORGN", statistic == "Mean", indicator == "DELTA_13_CH4_C")
ggplot(pop.origin.13c, aes(estimate, subpopulation)) +
  geom_bar(stat = "identity", aes(fill = subpopulation, color = subpopulation), alpha = 0.25) +
  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct, color = subpopulation), height = 0, size = 1) +
  theme(axis.title.y = element_blank())

# 13C; horizontal bar by size
pop.size.ch4 <- filter(pop, type == "AREA_CAT6", statistic == "Mean", indicator == "DISSOLVED_CH4")
ggplot(pop.size.ch4, aes(estimate, subpopulation)) +
  geom_bar(stat = "identity", aes(fill = subpopulation, color = subpopulation), alpha = 0.25) +
  geom_errorbarh(aes(xmax = ucb95pct, xmin = lcb95pct, color = subpopulation), height = 0, size = 1) +
  theme(axis.title.y = element_blank())

# 13CH4 Bubble plots

ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf, aes(size = delta.13.ch4), 
          show.legend = "point") + # improve legend
  ggtitle("13CH4") +
  scale_size(name = "13C-CH4",
             range = c(0.05, 3), # custom size range
             breaks = c(-80, -50, -25, 0)) # custom breaks

```

Similarly, we can look at population level estimates of GHG emission rates (mass of GHG time^-1^ area^-1^).

We can also look at the GHG flux across the air-water interface (mass GHG time^-1^)

## Simple random sample vs local neighborhood variance estimator
Population level estimates of variance can be calculated using a simple random sample or local neighborhood estimators.  The local neighborhood estimator will yield lower variance estimates if there is spatial structure in the data.  Below are plotted the 95% confidence intervals calculated using both methods.  In nearly all cases the local neighborhood estimator yields a more restricted interval.

```{r, message=FALSE}
# calculate 95% CI using local neighborhood estimator
pop.eco.diss <- pop.eco.diss %>%
  mutate(ci95.local = ucb95pct - lcb95pct)

# read in SRS estimator population estimates provided by Tom Kincaid (12/12/2019)
pop.srs <- read.csv(file = paste0(localPath, 
                              "/Environmental Protection Agency (EPA)/",
                              "ORD NLA17 Dissolved Gas - Documents/",
                              "inputData/populationEstimates/",
                              "NLA_2017_Percentile_Estimates_SRS_20191212.csv"),
                as.is = TRUE) %>%
  toEPA()  # enforce formatting convention

# Convert units for dissolved gas concentration indicator
pop.srs <- pop.srs %>%
  mutate(estimate = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
                            estimate * 1000000, # mol/L -> umol/L for co2 and ch4
                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
                            estimate * 1000000000, # mol/L -> nmol/L for n2o
                            estimate)), # all others
         lcb95pct = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
                            lcb95pct * 1000000, # mol/L -> umol/L for co2 and ch4
                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
                            lcb95pct * 1000000000, # mol/L -> nmol/L for n2o
                            lcb95pct)), # all others
         ucb95pct = ifelse(grepl("dissolved.c", indicator, ignore.case = TRUE), # dissolved.co2 or dissolved.ch4
                            ucb95pct * 1000000, # mol/L -> umol/L for co2 and ch4
                           ifelse(grepl("dissolved.n", indicator, ignore.case = TRUE), # dissolved.n2o
                            ucb95pct * 1000000000, # mol/L -> nmol/L for n2o
                            ucb95pct))) # all others

# extract population estimates for ecoregion and nation
# dissolved gas indicator
pop.srs.eco.diss <- pop.srs %>% 
  filter(type %in% c("AG_ECO9_NM", "National"), # extract estimates for ecoregion and nation
         statistic == "Mean") %>% # extract mean estimate
  mutate(ci95.srs = ucb95pct - lcb95pct) %>%
  filter(grepl("dissolved", indicator, ignore.case = TRUE))

# merge local neighborhood and SRS variance estimates
pop.local.srs <- dplyr::full_join(pop.eco.diss,
                                  select(pop.srs.eco.diss, -nresp, -estimate, 
                                         -stderror, -lcb95pct, -ucb95pct)) %>%
  select(-nresp, -estimate, -stderror, -lcb95pct, -ucb95pct) %>%
  pivot_longer(cols = starts_with("ci95"))

# look at dissolved concentration indicator
ggplot(pop.local.srs, aes(value, subpopulation)) +
  geom_point(aes(color = name)) +
  facet_wrap(~indicator, scales = "free_x") +
  xlab("95% confidence interval of mean") +
  theme(axis.title.y = element_blank())
```

## Bubble plots
Bubble plots can be used to visualize patterns in geospatial data.  In the plots below, bubble size reflects the gas concentration and color reflects saturation status (i.e. source or sink).  The code immediately below converts the dataframe to a spatial object and reads in the ecoregion ploygons.  The first bubble plot below is of dissolved CO2 which seems to show a pattern of higher concentrations along the eastern seaboard.
```{r, message=FALSE}
# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple")

# Check CRS
st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
st_crs(ecoR) # 5070
```

```{r, message=FALSE}
# CO2 Bubble plots

ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf, aes(size = dissolved.co2.umol, color = co2.src.snk), 
          show.legend = "point") + # improve legend
  ggtitle("Dissolved CO2 (uM)") +
  scale_color_manual(values = c("white", "black"), name = "source/sink") +
  scale_size(name = "dissolved CO2 (uM)",
             range = c(0.1, 5), # custom size range
             breaks = c(10, 100, 300, 600)) # custom breaks
```

Spatial patterns in dissolved CH4 are not apparent, although all but two sites are supersaturated.

```{r}
ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf %>% filter(!is.na(ch4.src.snk)), 
          aes(size = dissolved.ch4.umol, color = ch4.src.snk), 
          show.legend = "point") +
  ggtitle("Dissolved CH4 (uM)") +
  scale_color_manual(values = c("white", "black"), name = "source/sink") +
  scale_size(name = "dissolved CH4 (uM)",
             range = c(0.1, 10), # custom size range
             breaks = c(0.1, 10, 50, 100)) # custom breaks

ggsave("output/figures/ch4BubblePlot.tiff" )
```

The prevalance of undersaturated dissolved N2O concentrations is an important finding.  The IPCC assumes surface waters are an important source of N2O to the atmosphere.   Webb et al. recently published a paper in PNAS reporting that 70 out of 101 reservoirs sampled in Saskatchewan were N2O sinks.  This paper made quite a splash.  In our study, `r sum(dg$n2o.src.snk == "sink")` out of `r sum(!is.na(dg$dissolved.n2o.nmol))` sampled waterbodies were an N2O sink. It is somewhat interesting to note that the some of the highest concentrations were observed in Indiana/Ohio, where I conducted much of my dissertation research on aquatic N2O emissions.

```{r}
ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf, aes(size = dissolved.n2o.nmol, color = n2o.src.snk),
          show.legend = "point") +
  ggtitle("Dissolved N2O (nM)") +
  scale_color_manual(values = c("white", "black"), name = "source/sink") +
  scale_size(name = "dissolved N2O (nM)",
             range = c(0.1, 10), # custom size range
             breaks = c(1, 10, 25, 50, 100)) # custom breaks
```

## Geopackage
Create a simple feature on entire dg data frame.  The geopackage will contain two simple features, the dissolved gas sampled points/lakes and the nine aggregated ecoregions.
```{r, geopackage, eval=FALSE, echo=FALSE,}

names(dg)
dg_1 <- clean_names(dg)
names(dg_1)
# dropped onoffnet variable as I don't think it occurs in this version of the data frame, dg.2021-01-14.RData
dg_1 <- filter( dg_1, sample_source == "DG")

dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag_eco9, ag_eco9_nm, nitrate_n_result, chla_result, ntl_result, ph_result, ptl_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, pstl_code , area_cat6, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag_eco9 != "AK")

names(dg_1)
str(dg_1)

dg_1 %>% 
  ggplot(aes(as.factor(x = ag_eco9), y = log(dissolved_n2o_nmol))) + geom_boxplot()
# now no longer show a boxplot for NA category

dg_1 %>% 
  ggplot(aes(as.factor(x = ag_eco9), y = (n2o_sat_ratio))) + geom_boxplot()
# now no longer show a boxplot for NA category

#filter in View returns 514 nitrate_n_result equal to zero
dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3)) %>%  
  mutate(log_ntl = round(log(ntl_result),3)) %>% 
  mutate(log_ptl = round(log(ptl_result),3)) %>% 
  mutate(log_chla = round(log(chla_result),1)) %>% 
  mutate(log_si = round(log(silica_result),2)) %>% 
  mutate(log_ca = round(log(calcium_result),2)) %>% 
  mutate(ph_result = round(ph_result,2))
# %>% 
#   mutate(nitrate_n_b = case_when(
#     nitrate_n_result <= 0.0001 ~ "low",
#     nitrate_n_result > 0.0001 ~ "high"))  
# 
# dg_1$nitrate_n_b <- factor(dg_1$nitrate_n_b, levels = c("low", "high"))
# summary(dg_1$nitrate_n_b)

# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)

dg_sf1 <- st_as_sf(dg_1, coords = c("map_lon_dd", "map_lat_dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting 
dim(dg_sf1) # 1184 observations

dg_sf1 <- filter(dg_sf1, sample_source == "DG") # now 1082 obs vs. 1184 obs vs. previously 1185 obs., why change in sample size
st_crs(dg_sf1)

# Bryan suggested EPSG 4326
# Bryan Chastain as GeoServices suggested reducing the number of fields, and I tried that but was still unable to open the geodatabase in ArcMap 
# Based on gis stackexchange I think the "." in variable names is causing the problem
# https://gis.stackexchange.com/questions/319611/arcgis-geopackage-arcmap-drawing-error?noredirect=1
# Even with change to _ still get error message about invalid coordinate system when st_transform(5070) used

# Check CRS
st_crs(ecoR)
names(ecoR)
# changed from 4326 to 5070
ecoR1 <- st_transform(ecoR, 5070) %>% 
  select(OBJECTID, ag.eco9.nm, ag.eco9.nm) # Grab variables of interest
st_crs(ecoR1)

# write out dg.sf and ecoR as two layers in geopackage, gpkg, geodatabase
# 74. Creating a geopackage geodatabase, then adding more layers
# https://github.com/r-spatial/sf/issues/392
# creating a _v1.gpkg having large sample size and should be in Albers
# st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", quiet = FALSE, 
#          delete_layer = TRUE) # needed, else function gives error if object already exists
st_layers("nladg_v1.gpkg")

# N.B. b/c old layer with untransformed was already in geodatabase have to delete it and then append the new layer
st_write(dg_sf1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "dg_sf1", driver = "GPKG", delete_layer = TRUE, append = TRUE, quiet = FALSE)
st_layers("nladg_v1.gpkg")

# https://gis.stackexchange.com/questions/223240/writing-multiple-layers-to-geopackage-using-writeogr-in-r
st_write(ecoR1, dsn = file.path(getwd(), "nladg_v1.gpkg"), layer = "ecoR1", driver = "GPKG", append = TRUE, quiet = FALSE, delete_layer = TRUE) # needed, else function gives error if object already exists
st_layers("nladg_v1.gpkg")

names(dg_sf1)
summary(dg_sf1$dissolved_n2o_nmol) # matches Roy's stats
```

## Summarizing distances
Getting distances between sites overall and by ecoregion to help set distances for lagged scatterplots.  First, Worked out steps on a sample of NLA sites.  Could only get parts of Anna Springsteen's code to work.
```{r, eval=FALSE, echo=FALSE, distances}
names(dg.sf)
head(dg.sf$ag.eco9.nm)
# distinct returns all unique geometry + attribute combinations.  convert to df if not interested in geometry
dg.sf %>% as.data.frame() %>% distinct(ag.eco9.nm) 
# distinct.Spatial(dg.sf, ag.eco9.nm) # this doesn't work either.
distinct(dg, ag.eco9.nm)

# take a small sample of dg.sf to test getting distances
set.seed(1859)
# st_sample only returns a list of coordinates
nla_samp <- st_sample(dg.sf, 10, type = "random", exact = TRUE)
plot(nla_samp)
class(nla_samp)
head(nla_samp)

# sample_n keeps attributes and geometries
nla_samp1 <- sample_n(dg.sf, 10)
plot(nla_samp1, max.plot=1)
class(nla_samp1)
head(nla_samp1$site.id, 10)
distmat <- st_distance(nla_samp1,nla_samp1,by_element = FALSE)
class(distmat) # thought this would return matrix
max(distmat)
min(distmat)
summary(distmat) # think this returns summary of all pairwise distances on each obs, v1-v10
dim(distmat)

rdistmat <- distmat[1:10, ]
print(rdistmat)
rdistmat1 <- as.vector(rdistmat)
round(min(rdistmat1[rdistmat1!=0]))
round(quantile(rdistmat1[rdistmat1!=0], probs=seq(0,1, 0.25), names=F)[2])
round(median(rdistmat1[rdistmat1!=0]))
round(quantile(rdistmat1[rdistmat1!=0], probs=seq(0,1, 0.25), names=F)[4])
round(max(rdistmat1[rdistmat1!=0]))

names(ecoR)
class(ecoR)
# contrast ecoRname <- as.data.frame(select(ecoR,"ag.eco9.nm"))
# geom is sticky with select but not pull
# ecoRname <- as.data.frame(pull(ecoR,"ag.eco9.nm"))
ecoRname1 <- as.character(pull(ecoR,"ag.eco9.nm")) #Springsteen code needs to be character to populate columns of matrix below
class(ecoRname1)
head(ecoRname1, 9)
# initialize summary tables
euc.summary <- matrix(NA, nrow=5, ncol=length(ecoRname1)+1) # initializing blank matrices to fill in the loop
dim(euc.summary)
colnames(euc.summary) <- c("ALL",ecoRname1) #indicates same assignment in each object
rownames(euc.summary) <- c("Min","Q1","Median","Q3","Max")
euc.summary

# Add all distances summary statistics to column 1 of the summary table (i.e. not by region)
euc.summary[,1] <- round(c( min(rdistmat1[rdistmat1!=0]), quantile(rdistmat1[rdistmat1!=0], probs=seq(0,1, 0.25), names=F)[2], median(rdistmat1[rdistmat1!=0]), quantile(rdistmat1[rdistmat1!=0], probs=seq(0,1, 0.25), names=F)[4], max(rdistmat1[rdistmat1!=0]) ),1)

print(euc.summary)

# Unsure that Springsteen function below will work as euc is an spDists created object that I think retained ecoregional & not sure that's so with rdistmat1
# Add within-region summary statistics to summary table 
# for (r in 1:length(ecoRname1)){ # loop through regions and fill in the matrices
#   euc.reg <- euc[grep(ecoRname1[r], rownames(euc)), grep(ecoRname1[r],colnames(euc))] # retrieving only columns and rows in the distance matrix that contain the same region code in their column/row name
#   # retrieving only columns and rows in the distance matrix that contain the same region code in their column/row name
#   euc.summary[, (r+1)] <- round(c( min(euc.reg[euc.reg!=0]), quantile(euc.reg[euc.reg!=0], probs=seq(0,1, 0.25), names=F)[2], median(euc.reg[euc.reg!=0]), quantile(euc.reg[euc.reg!=0], probs=seq(0,1, 0.25), names=F)[4], max(euc.reg[euc.reg!=0]) ),1)
# }

# Distances for all NLA sites
distmat_dg <- st_distance(dg.sf,dg.sf,by_element = FALSE)
rdistmat_dg <- distmat_dg[1:1185, ]
rdistmat1_dg <- as.vector(rdistmat_dg)

# initialize summary tables
euc.summary_dg <- matrix(NA, nrow=5, ncol=length(ecoRname1)+1) # initializing blank matrices to fill in the loop
dim(euc.summary_dg)
colnames(euc.summary_dg) <- c("ALL",ecoRname1) #indicates same assignment in each object
rownames(euc.summary_dg) <- c("Min","Q1","Median","Q3","Max")

euc.summary_dg[,1] <- round(c( min(rdistmat1_dg[rdistmat1_dg!=0]), quantile(rdistmat1_dg[rdistmat1_dg!=0], probs=seq(0,1, 0.25), names=F)[2], median(rdistmat1_dg[rdistmat1_dg!=0]), quantile(rdistmat1_dg[rdistmat1_dg!=0], probs=seq(0,1, 0.25), names=F)[4], max(rdistmat1_dg[rdistmat1_dg!=0]) ),1)

print(euc.summary_dg)

# brute force to get 5-number summary of ecoregions cut-n-paste ecoregions to get specific data frames:  cpl, nap, npl, sap, spl, tpl, umw, wmt, xer
names(dg.sf)

count(dg.sf, ag.eco9.nm) %>% print(n = Inf)

# Get ecoregion-specific NLA sites
cpl <- dg.sf %>% filter(ag.eco9.nm == "Coastal Plains")

distmat_cpl <- st_distance(cpl, cpl,by_element = FALSE)
dim(distmat_cpl)
rdistmat_cpl <- distmat_cpl[1:150, ] # change to match # sites/rows
rdistmat1_cpl <- as.vector(rdistmat_cpl)

round(min(rdistmat1_cpl[rdistmat1_cpl!=0]))
round(quantile(rdistmat1_cpl[rdistmat1_cpl!=0], probs=seq(0,1, 0.25), names=F)[2])
round(median(rdistmat1_cpl[rdistmat1_cpl!=0]))
round(quantile(rdistmat1_cpl[rdistmat1_cpl!=0], probs=seq(0,1, 0.25), names=F)[4])
round(max(rdistmat1_cpl[rdistmat1_cpl!=0]))

# GitHub question:  if I save a workspace, .RData, will that be pushed when I push to GitHub?  Or, should data frames go in #.gitignore?
# Jake - we are using a public github repo under the EPA enterprise license.  Anything we push to github can be downloaded by the public.  We should use the shared documents library at our SharePoint site for data.  Code can go to github.
```

## Lagged scatterplots
Note that can use simple feature object, such as cpl, for histogram, but spatial object needed for lagged scatterplots.  I took log10 transform to  get data somewhat unimodal and symmetric.
```{r, eval=FALSE, echo=FALSE, laggedscatterplot}
# cpl
cpl_sp <- as(cpl, "Spatial")
class(cpl_sp)
ggplot(cpl,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(cpl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))

hscat(dissolved.ch4.umol~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))

hscat(dissolved.ch4.umol~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))

ggplot(cpl,aes(dissolved.co2.umol)) + geom_histogram()
ggplot(cpl,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.co2.umol)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))

hscat(dissolved.co2.umol~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))

hscat(log10(dissolved.co2.umol)~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))

ggplot(cpl,aes(dissolved.n2o.nmol)) + geom_histogram()
ggplot(cpl,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.n2o.nmol)~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))

hscat(dissolved.n2o.nmol~1, cpl_sp, c(0, 50000, 150000, 250000, 500000, 1000000, 1500000))

hscat(log10(dissolved.n2o.nmol)~1, cpl_sp, c(0, 10000, 25000, 50000, 150000, 250000, 500000))

# nap # nap object doesn't exist
nap <- dg.sf %>% filter(ag.eco9.nm == "Northern Appalachians")
nap_sp <- as(nap, "Spatial")
class(nap_sp)
ggplot(nap,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(nap,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(dissolved.ch4.umol~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(log10(dissolved.ch4.umol)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))

ggplot(nap,aes(dissolved.co2.umol)) + geom_histogram()
ggplot(nap,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.co2.umol)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(dissolved.co2.umol~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(log10(dissolved.co2.umol)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))

ggplot(nap,aes(dissolved.n2o.nmol)) + geom_histogram()
ggplot(nap,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.n2o.nmol)~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(dissolved.n2o.nmol~1, nap_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(log10(dissolved.n2o.nmol)~1, nap_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))

# npl 
npl <- dg.sf %>% filter(ag.eco9.nm == "Northern Plains")
npl_sp <- as(npl, "Spatial")
class(npl_sp)
ggplot(npl,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(npl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(dissolved.ch4.umol~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(log10(dissolved.ch4.umol)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))

ggplot(npl,aes(dissolved.co2.umol)) + geom_histogram()
ggplot(npl,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.co2.umol)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(dissolved.co2.umol~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(log10(dissolved.co2.umol)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))

ggplot(npl,aes(dissolved.n2o.nmol)) + geom_histogram()
ggplot(npl,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.n2o.nmol)~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(dissolved.n2o.nmol~1, npl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(log10(dissolved.n2o.nmol)~1, npl_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))

# sap
sap <- dg.sf %>% filter(ag.eco9.nm == "Southern Appalachians")
sap_sp <- as(sap, "Spatial")
class(sap_sp)
ggplot(sap,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(sap,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))

hscat(dissolved.ch4.umol~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))

hscat(log10(dissolved.ch4.umol)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))

ggplot(sap,aes(dissolved.co2.umol)) + geom_histogram()
ggplot(sap,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.co2.umol)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))

hscat(dissolved.co2.umol~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))

hscat(log10(dissolved.co2.umol)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))

ggplot(sap,aes(dissolved.n2o.nmol)) + geom_histogram()
ggplot(sap,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.n2o.nmol)~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))

hscat(dissolved.n2o.nmol~1, sap_sp, c(0, 10000, 50000, 150000, 300000, 450000, 650000))

hscat(log10(dissolved.n2o.nmol)~1, sap_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))

# spl
spl <- dg.sf %>% filter(ag.eco9.nm == "Southern Plains")
spl_sp <- as(spl, "Spatial")
class(spl_sp)
ggplot(spl,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(spl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))

hscat(dissolved.ch4.umol~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))

hscat(log10(dissolved.ch4.umol)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))

ggplot(spl,aes(dissolved.co2.umol)) + geom_histogram()
ggplot(spl,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.co2.umol)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))

hscat(dissolved.co2.umol~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))

hscat(log10(dissolved.co2.umol)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))

ggplot(spl,aes(dissolved.n2o.nmol)) + geom_histogram()
ggplot(spl,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.n2o.nmol)~1, spl_sp, c(0, 10000, 50000, 150000, 250000, 350000, 500000))

hscat(dissolved.n2o.nmol~1, spl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 400000))

hscat(log10(dissolved.n2o.nmol)~1, spl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))

# tpl
tpl <- dg.sf %>% filter(ag.eco9.nm == "Temperate Plains")
tpl_sp <- as(tpl, "Spatial")
class(tpl_sp)
ggplot(tpl,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(tpl,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))

hscat(dissolved.ch4.umol~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))

hscat(log10(dissolved.ch4.umol)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))

ggplot(tpl,aes(dissolved.co2.umol)) + geom_histogram()
ggplot(tpl,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.co2.umol)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))

hscat(dissolved.co2.umol~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))

hscat(log10(dissolved.co2.umol)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))

ggplot(tpl,aes(dissolved.n2o.nmol)) + geom_histogram()
ggplot(tpl,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.n2o.nmol)~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))

hscat(dissolved.n2o.nmol~1, tpl_sp, c(0, 10000, 50000, 100000, 200000, 300000, 600000))

hscat(log10(dissolved.n2o.nmol)~1, tpl_sp, c(0, 25000, 50000, 75000, 100000, 125000, 150000))

# umw
umw <- dg.sf %>% filter(ag.eco9.nm == "Upper Midwest")
umw_sp <- as(umw, "Spatial")
class(umw_sp)
ggplot(umw,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(umw,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))

hscat(dissolved.ch4.umol~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))

hscat(log10(dissolved.ch4.umol)~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))

ggplot(umw,aes(dissolved.co2.umol)) + geom_histogram()
ggplot(umw,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.co2.umol)~1, umw_sp, c(0, 10000, 50000, 75000, 100000, 200000, 350000))

hscat(dissolved.co2.umol~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))

hscat(log10(dissolved.co2.umol)~1, umw_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))

ggplot(umw,aes(dissolved.n2o.nmol)) + geom_histogram()
ggplot(umw,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.n2o.nmol)~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))

hscat(dissolved.n2o.nmol~1, umw_sp, c(0, 25000, 50000, 75000, 100000, 250000, 350000))

hscat(log10(dissolved.n2o.nmol)~1, umw_sp, c(0, 25000, 50000, 100000, 200000, 300000, 400000))

# wmt
wmt <- dg.sf %>% filter(ag.eco9.nm == "Western Mountains")
wmt_sp <- as(wmt, "Spatial")
class(wmt_sp)
ggplot(wmt,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(wmt,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))

hscat(dissolved.ch4.umol~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))

hscat(log10(dissolved.ch4.umol)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))

ggplot(wmt,aes(dissolved.co2.umol)) + geom_histogram()
ggplot(wmt,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.co2.umol)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))

hscat(dissolved.co2.umol~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))

hscat(log10(dissolved.co2.umol)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))

ggplot(wmt,aes(dissolved.n2o.nmol)) + geom_histogram()
ggplot(wmt,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.n2o.nmol)~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))

hscat(dissolved.n2o.nmol~1, wmt_sp, c(0, 10000, 50000, 100000, 200000, 350000, 700000))

hscat(log10(dissolved.n2o.nmol)~1, wmt_sp, c(0, 25000, 50000, 100000, 200000, 350000, 700000))

# xer
xer <- dg.sf %>% filter(ag.eco9.nm == "Xeric")
xer_sp <- as(xer, "Spatial")
class(xer_sp)
ggplot(xer,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(xer,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
# 10 km bin/lag only has 2 points plotted so start with 50 km
hscat(log10(dissolved.ch4.umol)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))

hscat(dissolved.ch4.umol~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))

hscat(log10(dissolved.ch4.umol)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))

ggplot(xer,aes(dissolved.co2.umol)) + geom_histogram()
ggplot(xer,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.co2.umol)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))

hscat(dissolved.co2.umol~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))

hscat(log10(dissolved.co2.umol)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))

ggplot(xer,aes(dissolved.n2o.nmol)) + geom_histogram()
ggplot(xer,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.n2o.nmol)~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))

hscat(dissolved.n2o.nmol~1, xer_sp, c(0, 50000, 100000, 200000, 400000, 800000))

hscat(log10(dissolved.n2o.nmol)~1, xer_sp, c(0, 25000, 50000, 750000, 100000, 200000, 400000, 800000))

# all
dg_sp <- as(dg.sf, "Spatial")
class(dg_sp)
ggplot(dg.sf,aes(dissolved.ch4.umol)) + geom_histogram()
ggplot(dg.sf,aes(dissolved.ch4.umol)) + geom_histogram() + scale_x_log10()
hscat(log10(dissolved.ch4.umol)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))

hscat(dissolved.ch4.umol~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))

hscat(log10(dissolved.ch4.umol)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))

ggplot(dg.sf,aes(dissolved.co2.umol)) + geom_histogram()
ggplot(dg.sf,aes(dissolved.co2.umol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.co2.umol)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))

hscat(dissolved.co2.umol~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))

hscat(log10(dissolved.co2.umol)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))

ggplot(dg.sf,aes(dissolved.n2o.nmol)) + geom_histogram()
ggplot(dg.sf,aes(dissolved.n2o.nmol)) + geom_histogram() + scale_x_log10()

hscat(log10(dissolved.n2o.nmol)~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))

hscat(dissolved.n2o.nmol~1, dg_sp, c(0, 50000, 100000, 200000, 400000, 800000, 1500000))

hscat(log10(dissolved.n2o.nmol)~1, dg_sp, c(0, 10000, 25000, 50000, 100000, 200000, 300000, 400000))

```

## k-NN = 1 and Moran plots
As with the lagged scatter plots, I had to take a simple feature object and convert it to a spatial object so I could use the nearest-neighbor and Moran scatter plot functions from the spdep package.  A Moran plot is based on plotting the N~2~O variable, dissolved.n2o, of a focal point on the x-axis and the average dissolved.n2o of the neighboring points on the y-axis.  Moran scatter plots were made for each of the 9 ecoregions and 1 for all of the NLA sites.  For each plot, a neighborhood of points was defined by the maximum distance to first nearest neighbor.  A distance-based spatial relationship was used because many of the ecoregions have several non-contiguous polygons or areas, which prevents a contiguity-based spatial relationship from being used.  The slope of the line through the scatter plot of points measures spatial autocorrelation, with a positive slope indicating positive spatial autocorrelation.

The code below is to be run separately for each ecoregion.  The CPL, TPL, and UMW ecoregions show strong positive spatial autocorrelation.  Negative spatial autocorrelation occurs in the NPL ecoregion because of the maximum outlier at a South Dakota site.  The SPL and XER ecoregions show no evidence of spatial autocorrelation.  Interestingly, it appears that the ecoregions with the highest dissolved.n2o concentrations, CPL, TPL, and UMW, drive the national pattern.
```{r, knn_moran}
names(dg)
# also includ ag.eco9.nm, ag.eco9.nm feature
dg_1 <- dg
names(dg_1)
dg_1 <- clean_names(dg_1)
names(dg_1)
distinct(dg_1, ag_eco9_nm)
# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)

dg_sf1 <- st_as_sf(dg_1, coords = c("map_lon_dd", "map_lat_dd"),crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

st_crs(dg_sf1)
glimpse(dg_sf1)
# Get ecoregion-specific NLA sites
## CPL
cpl <- dg_sf1 %>% filter(ag_eco9 == "CPL" & visit_no == 1 & sample_source == "DG")

cpl_sp <- as(cpl, "Spatial")
class(cpl_sp)

ds1coords <- coordinates(cpl_sp)
ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = cpl_sp$site_id)
# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors

dsnb1 <- dnearneigh(ds1coords, d1=0, d2=134761, row.names=cpl_sp$site_id, longlat=FALSE)
summary(dsnb1)

plot(cpl_sp)
plot(ds1nb1, ds1coords, labels = cpl_sp$site_id, add = TRUE)

dsnb1.listw <- nb2listw(dsnb1, style = "W")
summary(dsnb1.listw)

class(cpl)
# cpl_df <- st_drop_geometry(cpl)
# names(cpl_df)
summary(cpl$dissolved_n2o_nmol)

moran.plot(log(cpl$dissolved_n2o_nmol), listw = dsnb1.listw, labels = cpl$site_id)
rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)

## NAP
nap <- dg_sf1 %>% filter(ag_eco9_nm == "NAP" & visit_no == 1 & sample_source == "DG")

nap_sp <- as(nap, "Spatial")
class(nap_sp)

ds1coords <- coordinates(nap_sp)
ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = nap_sp$site_id)
# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors

dsnb1 <- dnearneigh(ds1coords, d1=0, d2=112630, row.names=nap_sp$site_id, longlat=FALSE)
summary(dsnb1)

plot(nap_sp)
plot(ds1nb1, ds1coords, labels = nap_sp$site_id, add = TRUE)

dsnb1.listw <- nb2listw(dsnb1, style = "W")
summary(dsnb1.listw)

class(nap)
# nap_df <- st_drop_geometry(nap)
# names(nap_df)
summary(nap$dissolved_n2o_nmol)

moran.plot(log(nap$dissolved_n2o_nmol), listw = dsnb1.listw, labels = nap$site_id)
rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)

## NPL
npl <- dg_sf1 %>% filter(ag_eco9_nm == "NPL" & visit_no == 1 & sample_source == "DG")

npl_sp <- as(npl, "Spatial")
class(npl_sp)

ds1coords <- coordinates(npl_sp)
ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = npl_sp$site_id)
# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors

dsnb1 <- dnearneigh(ds1coords, d1=0, d2=215395, row.names=npl_sp$site_id, longlat=FALSE)
summary(dsnb1)

plot(npl_sp)
plot(ds1nb1, ds1coords, labels = npl_sp$site_id, add = TRUE)

dsnb1.listw <- nb2listw(dsnb1, style = "W")
summary(dsnb1.listw)

class(npl)
# npl_df <- st_drop_geometry(npl)
# names(npl_df)
summary(npl$dissolved_n2o_nmol)

moran.plot(log(npl$dissolved_n2o_nmol), listw = dsnb1.listw, labels = npl$site_id)
rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)

## SAP
sap <- dg_sf1 %>% filter(ag_eco9_nm == "SAP" & visit_no == 1 & sample_source == "DG")

sap_sp <- as(sap, "Spatial")
class(sap_sp)

ds1coords <- coordinates(sap_sp)
ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = sap_sp$site_id)
# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors

dsnb1 <- dnearneigh(ds1coords, d1=0, d2=128285, row.names=sap_sp$site_id, longlat=FALSE)
summary(dsnb1)

plot(sap_sp)
plot(ds1nb1, ds1coords, labels = sap_sp$site_id, add = TRUE)

dsnb1.listw <- nb2listw(dsnb1, style = "W")
summary(dsnb1.listw)

class(sap)
# sap_df <- st_drop_geometry(sap)
# names(sap_df)
summary(sap$dissolved_n2o_nmol)

moran.plot(log(sap$dissolved_n2o_nmol), listw = dsnb1.listw, labels = sap$site_id)
rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)

## SPL
spl <- dg_sf1 %>% filter(ag_eco9_nm == "SPL" & visit_no == 1 & sample_source == "DG")

spl_sp <- as(spl, "Spatial")
class(spl_sp)

ds1coords <- coordinates(spl_sp)
ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = spl_sp$site_id)
# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors

dsnb1 <- dnearneigh(ds1coords, d1=0, d2=173395, row.names=spl_sp$site_id, longlat=FALSE)
summary(dsnb1)

plot(spl_sp)
plot(ds1nb1, ds1coords, labels = spl_sp$site_id, add = TRUE)

dsnb1.listw <- nb2listw(dsnb1, style = "W")
summary(dsnb1.listw)

class(spl)
# spl_df <- st_drop_geometry(spl)
# names(spl_df)
summary(spl$dissolved_n2o_nmol)

moran.plot(log(spl$dissolved_n2o_nmol), listw = dsnb1.listw, labels = spl$site_id)
rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)

## TPL
tpl <- dg_sf1 %>% filter(ag_eco9_nm == "TPL" & visit_no == 1 & sample_source == "DG")

tpl_sp <- as(tpl, "Spatial")
class(tpl_sp)
nrow(zerodist(tpl_sp)) # indicates 1 pair of identical coordinates

(zerodist(tpl_sp)) # this gives the row numbers of the pair 30 & 36
View(tpl) # to view rows 30, site_id NLA17_IN-10002, & 36, NLA17_IN-10047

# Warning message from code below:
# In knearneigh(ds1coords, k = 1) : knearneigh: identical points found

ds1coords <- coordinates(tpl_sp)
ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = tpl_sp$site_id)
# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors

dsnb1 <- dnearneigh(ds1coords, d1=0, d2=223893, row.names=tpl_sp$site_id, longlat=FALSE)
summary(dsnb1)

plot(tpl_sp)
plot(ds1nb1, ds1coords, labels = tpl_sp$site_id, add = TRUE)

dsnb1.listw <- nb2listw(dsnb1, style = "W")
summary(dsnb1.listw)

class(tpl)
# tpl_df <- st_drop_geometry(tpl)
# names(tpl_df)
summary(tpl$dissolved_n2o_nmol)

moran.plot(log(tpl$dissolved_n2o_nmol), listw = dsnb1.listw, labels = tpl$site_id)
rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)

## UMW
umw <- dg_sf1 %>% filter(ag_eco9_nm == "UMW" & visit_no == 1 & sample_source == "DG")

umw_sp <- as(umw, "Spatial")
class(umw_sp)
nrow(zerodist(umw_sp)) # indicates 1 pair of identical coordinates

(zerodist(umw_sp)) # this gives the row numbers of the pair 136 & 158
View(umw) # to view rows 136, site_id NLA17_WI-10027, & 158, NLA17_WI-10127

# Warning message from code below:
# In knearneigh(ds1coords, k = 1) : knearneigh: identical points found

ds1coords <- coordinates(umw_sp)
ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = umw_sp$site_id)
# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors

dsnb1 <- dnearneigh(ds1coords, d1=0, d2=99590, row.names=umw_sp$site_id, longlat=FALSE)
summary(dsnb1)

plot(umw_sp)
plot(ds1nb1, ds1coords, labels = umw_sp$site_id, add = TRUE)

dsnb1.listw <- nb2listw(dsnb1, style = "W")
summary(dsnb1.listw)

class(umw)
# umw_df <- st_drop_geometry(umw)
# names(umw_df)
summary(umw$dissolved_n2o_nmol)

moran.plot(log(umw$dissolved_n2o_nmol), listw = dsnb1.listw, labels = umw$site_id)
rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)

## WMT
wmt <- dg_sf1 %>% filter(ag_eco9_nm == "WMT" & visit_no == 1 & sample_source == "DG")

wmt_sp <- as(wmt, "Spatial")
class(wmt_sp)

# nrow(zerodist(wmt_sp)) # indicates 1 pair of identical coordinates
# (zerodist(wmt_sp)) # this gives the row numbers of the pair 136 & 158
# View(wmt) # to view rows 136, site_id NLA17_WI-10027, & 158, NLA17_WI-10127

# Warning message from code below:
# In knearneigh(ds1coords, k = 1) : knearneigh: identical points found

ds1coords <- coordinates(wmt_sp)
ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = wmt_sp$site_id)
# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors

dsnb1 <- dnearneigh(ds1coords, d1=0, d2=295987, row.names=wmt_sp$site_id, longlat=FALSE)
summary(dsnb1)

plot(wmt_sp)
plot(ds1nb1, ds1coords, labels = wmt_sp$site_id, add = TRUE)

dsnb1.listw <- nb2listw(dsnb1, style = "W")
summary(dsnb1.listw)

class(wmt)
# wmt_df <- st_drop_geometry(wmt)
# names(wmt_df)
summary(wmt$dissolved_n2o_nmol)

moran.plot(log(wmt$dissolved_n2o_nmol), listw = dsnb1.listw, labels = wmt$site_id)
rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)

## XER
xer <- dg_sf1 %>% filter(ag_eco9_nm == "XER" & visit_no == 1 & sample_source == "DG")

xer_sp <- as(xer, "Spatial")
class(xer_sp)

# nrow(zerodist(xer_sp)) # indicates 1 pair of identical coordinates
# (zerodist(xer_sp)) # this gives the row numbers of the pair 136 & 158
# View(xer) # to view rows 136, site_id NLA17_WI-10027, & 158, NLA17_WI-10127

# Warning message from code below:
# In knearneigh(ds1coords, k = 1) : knearneigh: identical points found

ds1coords <- coordinates(xer_sp)
ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = xer_sp$site_id)
# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors

dsnb1 <- dnearneigh(ds1coords, d1=0, d2=259280, row.names=xer_sp$site_id, longlat=FALSE)
summary(dsnb1)

plot(xer_sp)
plot(ds1nb1, ds1coords, labels = xer_sp$site_id, add = TRUE)

dsnb1.listw <- nb2listw(dsnb1, style = "W")
summary(dsnb1.listw)

class(xer)
# xer_df <- st_drop_geometry(xer)
# names(xer_df)
summary(xer$dissolved_n2o_nmol)

moran.plot(log(xer$dissolved_n2o_nmol), listw = dsnb1.listw, labels = xer$site_id)
rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)

## NLA
nla <- dg_sf1 %>% filter(visit_no == 1 & sample_source == "DG")

nla_sp <- as(nla, "Spatial")
class(nla_sp)

nrow(zerodist(nla_sp)) # indicates 2 pairs of identical coordinates
(zerodist(nla_sp)) # this gives the row numbers of the pair (192 & 210:  NLA17_IN-10002 & NLA17_IN-10047) and the pair (1028 & 1054:  NLA17_WI-10027, & 158, NLA17_WI-10127)
# View(nla)

# Warning message from code below:
# In knearneigh(ds1coords, k = 1) : knearneigh: identical points found

ds1coords <- coordinates(nla_sp)
ds1nb1 <- knn2nb(knearneigh(ds1coords, k=1), row.names = nla_sp$site_id)
# using the k=1 object to find the minimum distance at which all sites have a distance-based neighbor
dsnb1_dist <- unlist(nbdists(ds1nb1,ds1coords))
summary(dsnb1_dist)# use max distance from summary to assign distance to create neighbors

dsnb1 <- dnearneigh(ds1coords, d1=0, d2=259280, row.names=nla_sp$site_id, longlat=FALSE)
summary(dsnb1)

plot(nla_sp)
plot(ds1nb1, ds1coords, labels = nla_sp$site_id, add = TRUE)

dsnb1.listw <- nb2listw(dsnb1, style = "W")
summary(dsnb1.listw)

class(nla)
# nla_df <- st_drop_geometry(nla)
# names(nla_df)
summary(nla$dissolved_n2o_nmol)

moran.plot(log(nla$dissolved_n2o_nmol), listw = dsnb1.listw, labels = nla$site_id)
rm(ds1coords, ds1nb1, dsnb1, dsnb1.listw, dsnb1_dist)
```

## Correlograms: ncf and pgmirness packages
```{r, correlogram}

# Much of code below is based on example comparing correlograms among packages ncf, pgmirness, and spdep described in this chapter
# https://link.springer.com/chapter/10.1007/978-3-030-01989-1_5

coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)

dg_sf1 <- st_as_sf(dg_1, coords = c("map_lon_dd", "map_lat_dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

st_crs(dg_sf1)
names(dg_sf1)
nap_visit1 <- dg_sf1 %>% filter(ag_eco9_nm == "NAP" & visit_no == 1)
class(nap_visit1)
st_crs(nap_visit1)

nap_sp_visit1 <- as(nap_visit1, "Spatial")
class(nap_sp_visit1)

# trying ncf package which uses resampling to specify number of null permutations under the null to assess significance levels.  See if significance persist after increas resamp from 100 to 1000, which is the default.
# The x-intercept in the results is the distance at which objects are no more similar than expected by chance alone across the region.
# Run on CPL as that is giving correlation greater than 1, which according to Fletcher and Fortin might be due to small number of pairs of points used in that first distance bin or to outliers.  Perhaps, cluster of high Delaware sites (n visit 1 = 5) causes the correlation to be greater than 1.
# Looks like 71 pairs of points being used.
# With a lag, or bin distance, of 50 km only go out 10 bins, 500 km, and do a Bonferroni adjustment of 0.05/10 = 0.005
names(dg_1)
dg_cplvisit1 <- dg_1 %>% filter(ag_eco9 == "CPL" & visit_no == 1)
# N.B. default is resamp = 1000 so that can take 10 minutes for n =1090 sites
class(dg_cplvisit1)

correlo_n2o_ncf <- ncf::correlog(x = dg_cplvisit1$map_lon_dd, y = dg_cplvisit1$map_lat_dd, z = log(dg_cplvisit1$dissolved_n2o_nmol), increment = 50, resamp = 1000, latlon = TRUE)
names(correlo_n2o_ncf)
#plot(correlo_n2o_ncf) # works when pgmiress commented out

# B/c returns a list of list have to access with [[]]
print(correlo_n2o_ncf[["n"]])
print(correlo_n2o_ncf[["mean.of.class"]])
print(correlo_n2o_ncf[["correlation"]])
head(correlo_n2o_ncf[["mean.of.class"]])
plot(correlo_n2o_ncf[["mean.of.class"]], correlo_n2o_ncf[["correlation"]])

class(correlo_n2o_ncf)
#ncf::plot.correlog(correlo_n2o_ncf)

# CPL outlier check by working through all 10 pairwise removal of cluster of 5 high Delaware sites
out_cplvisit1 <- dg_1 %>% filter(ag_eco9 == "CPL" & visit_no == 1 & (site_id != "NLA17_DE-10003" & site_id != "NLA17_DE-10004"))

out_cplcorrelo_n2o <- ncf::correlog(x = out_cplvisit1$map_lon_dd, y = out_cplvisit1$map_lat_dd, z = log(out_cplvisit1$dissolved_n2o_nmol), increment = 50, resamp = 1000, latlon = TRUE)

#plot(out_cplcorrelo_n2o)
#rm(out_cplvisit1, out_cplcorrelo_n2o)

# trying pgirmess package uses library spdep including moran.test and geary.test functions.  Distances are euclidean and in the same units as the spatial coordinates.  As run below, using geographical coordinates in decimal degrees so are distance classes retuned in decimal degrees?
# what if run on a spatial points data frame?
# Uses normality assumption, which is faster, to assess significance
dg_napvisit1 <- dg_1 %>% filter(ag_eco9 == "NAP" & visit_no == 1)
coords_nap <- cbind(dg_napvisit1$map_lon_dd, dg_napvisit1$map_lat_dd)
correl_napn2o_pg <- pgirmess::correlog(coords_nap, log(dg_napvisit1$dissolved_n2o_nmol), method = "Moran", nbclass = NULL, alternative = "two.sided")

head(round(correl_napn2o_pg, 2))
tail(round(correl_napn2o_pg, 2))
plot(correl_napn2o_pg)

# apply to SpatialPointsDataFrame that has projected coordinates and believe returning distance classes in meters
coords_sp_nap <- coordinates(nap_sp_visit1)
correl_spnapn2o_pg <- pgirmess::correlog(coords_sp_nap, log(nap_sp_visit1@data$dissolved_n2o_nmol), method = "Moran", nbclass = 30, alternative = "two.sided")

head(round(correl_spnapn2o_pg, 2), 30)
tail(round(correl_spnapn2o_pg ,2))
plot(correl_spnapn2o_pg)
```

## Correlograms by Ecoregions
Doug Anderson wrote functions to deal with the list of list output from the correlog function in the ncf package.  He also wrote functions so that the correlogram of dissolved NO2 gas could be plotted by ecoregion using base plot, cowplot, or ggplot2 code.
```{r, correlogram_ecoregions}
# Original code from Doug Anderson at Neptune from file:  Neptune_correlogram_TDS05_Anderson.Rmd

# set the theme of the ggplot
theme_set(theme_bw()) # the theme I prefer is theme_bw()

## Can define a new function to plot
plot_correlog <- function(ncf_correlog, 
                          # names of list components
                          n_col = "n", 
                          mean_col = "mean.of.class",
                          cor_col = "correlation",
                          xint_col = "x.intercept",
                          pval_col = "p",
                          name_col = NULL, #can change to column that has names of interest
                          alpha = 0.05, # two sided significance level
                          # plot labels and adjustments
                          plot_title = "Correlogram", #can change to NULL
                          plot_subtitle = waiver(),
                          y_label = "Correlation",
                          x_label = "Distance (mean-of-class)",
                          pt_size = I(2), pt_color = "black",
                          ln_size = I(0.5), ln_color = "black", 
                          ln_type = "solid",
                          title_size = 14, subtitle_size = 14, 
                          axis_title_size = 14, axis_text_size = 14, 
                          title_color = "black", 
                          subtitle_color = "black",
                          axis_title_color = "black", 
                          axis_text_color = "black",
                          title_bold = TRUE, axis_bold = FALSE,
                          xaxis_color = "gray", xaxis_type = "solid",
                          theme_style = theme_bw()){
  # ncf_correlog is a list object exported from ncf::correlog() but can also be a dataframe with named columns

  # prepare a data set for plotting based on the list components or data frame columns
  dat <- data.frame(
    n = ncf_correlog[[n_col]],
    mean = ncf_correlog[[mean_col]],
    cor = ncf_correlog[[cor_col]],
    # xint = ncf_correlog[[xint_col]],
    pval = ncf_correlog[[pval_col]]
  )
  # create a variable to help plot the hollow and solid points
  # If a permutation test was performed, values significant at a nominal (two-sided) 5%-level 
  # will be represented by filled circles and non-significant values by open circles.
  dat$sig <- ifelse(dat$pval * 2 < alpha, TRUE, FALSE)

  ## Start the plot
  g <- ggplot(dat) +
    theme_style +
    geom_hline(yintercept = 0, color = xaxis_color, linetype = xaxis_type) +
    # geom_point(aes(x=mean, y=cor, shape=sig), size=pt_size, color=pt_color) +
    # geom_line(aes(x=mean, y=cor, group=1), size=ln_size, color=ln_color, linetype=ln_type) +
    lemon::geom_pointline(aes(x=mean, y=cor, shape=sig, group=1),
                          size = pt_size, color = pt_color, 
                          linetype = ln_type, linesize = ln_size, 
                          linecolor = ln_color) +
    scale_shape_manual(values=c("TRUE"=19, "FALSE"=1),
                       labels=c("TRUE"="Sig.", "FALSE"="Not Sig."),
                       guide=FALSE) +
    labs(x=x_label, y=y_label)

  ## add a title
  if(!is.null(name_col)){#include a title if column for name is provided or list object has a list item for name
    if(!is.null(plot_title)){ #if plot_title is a string, separate the plot_title with the other name by a ':'
      nm <- unique(as.character(ncf_correlog[[name_col]]))
      g <- g + 
        ggtitle(label = paste(plot_title, ": ", nm, sep=""), subtitle = plot_subtitle)
    } else{ #if plot_title is null, only include the name
      nm <- unique(as.character(ncf_correlog[[name_col]]))
      g <- g + 
        ggtitle(label = nm, subtitle = plot_subtitle)
    }
  } else{ # if there is no name column, only include the plot_title of interest (can be NULL)
    g <- g + 
      ggtitle(label = plot_title, subtitle = plot_subtitle) 
  }

  ## modify theme
  g <- g + 
    theme(plot.title = element_text(size=title_size, 
                                    hjust=0.5, #change to a value between 0 (left) and 1 (right)
                                    face=ifelse(title_bold == TRUE, "bold", "plain"), 
                                    color=title_color),
          plot.subtitle = element_text(size=subtitle_size, 
                                       hjust=0.5, 
                                       color=subtitle_color),
          axis.title = element_text(size=axis_title_size, 
                                    face=ifelse(axis_bold == TRUE, "bold", "plain"), 
                                    color=axis_title_color),
          axis.text = element_text(size=axis_text_size, #?theme
                                   color=axis_text_color), #?element_text
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
  return(g)
}

dg_1 <- select(dg,site.id:map.lon.dd, dissolved.n2o.nmol, ag.eco9.nm, ag.eco9.nm, ONOFFNET, nitrate.n.result,chla.result.vol, max.bf, temp.surf, temp.bottom, o2.surf, o2.bottom, lake.orgn, state, size.cat)
names(dg_1)
dg_1 <- clean_names(dg_1)
names(dg_1)
dg_1 <- dg_1 %>% mutate(log_n2o = round(log(dissolved_n2o_nmol),3))
## Subset dg_1 by visit_no == 1
dg_visit1 <- dg_1 %>% dplyr::filter(visit_no == 1)

## Perform the same test for each ag.eco9.nm using Z = "log_n2o"
## Start by defining a function
ncf_correlog <- function(dat, # data to send to ncf::correlog
                         z_col, # name of the column in the dataset for water quality parameter
                         i = 50, # increment to be used in ncf::correlog
                         N = 1000, # number of resamples in ncf::correlog
                         x_col = "map_lon_dd", # name of the column in the dataset for longitude
                         y_col = "map_lat_dd", # name of the column in the dataset for latitude
                         lat_lon = TRUE, # if the x and y coordinates for ncf::correlog are lat and lon
                         rem_missing = FALSE, # argument of ncf::correlog
                         run_quietly = FALSE){ # argument of ncf::correlog
  library(ncf)

  X = dat[[x_col]] #grabs the x_col from data.frame
  Y = dat[[y_col]] #grabs the y_col from data.frame
  Z = dat[[z_col]] #grabs the z_col from data.frame

  ## Use ncf::correlog
  crlg <- ncf::correlog(x = X, y = Y, z = Z, 
                        increment = i, resamp = N, latlon = lat_lon, 
                        na.rm = rem_missing, quiet = run_quietly)
  return(crlg)
}

## Split the dataframe by each ag.eco9.nm: turns the data frame into a list 
##   where each list item is a data frame filtered to that ag.eco9.nm
## lapply applies the function ncf_correlog on each list object (a data frame)
## the final product is a list of correlog objects
## this is similar to looping
## setting seed so get same result from permutations ncf::correlog function
set.seed(2035)
log_n2o_correlog_list <- dg_visit1 %>%
  split(.$ag_eco9_nm) %>% # '.' means use the data frame that preceded this line
  lapply(function(x) ncf_correlog(dat = x, z_col = "log_n2o")) 
#ncf_correlog is the function name, 
#dat is the paramater that takes a data.frame,
#z_col is the parameter that takes a string name for a column of a water quality paramater

## look at the structure: list of 9 items where each item is a 
##   list containing 6 items: n, mean, correlation, x.int, p, call
str(log_n2o_correlog_list)

## Function to remove the list item 'call' which is of a different length 
##   than the other items (n, mean, correlation, x.int, and p)
## This function will also convert each list item to a dataframe
crlg_to_dat <- function(ncf_correlog, # ncf_correlog is a list object that was output from ncf::correlog
                        n_col = "n", 
                        mean_col = "mean.of.class",
                        cor_col = "correlation",
                        xint_col = "x.intercept",
                        pval_col = "p"){ 
  dat <- data.frame(
    n = ncf_correlog[[n_col]],
    mean = ncf_correlog[[mean_col]],
    cor = ncf_correlog[[cor_col]],
    xint = ncf_correlog[[xint_col]],
    pval = ncf_correlog[[pval_col]]
  )
  return(dat)
}

## log_n20_correlog_list is already a list so there is no need to split
log_n2o_correlog_dats <- lapply(log_n2o_correlog_list, 
                                function(x) crlg_to_dat(ncf_correlog = x))
str(log_n2o_correlog_dats) # each item in the list is now a dataframe

## combine into one dataframe to grab the ag.eco9.nm (notice the name no longer has an s at the end)

log_n2o_correlog_dat <- log_n2o_correlog_dats %>%
  dplyr::bind_rows(., .id = "ag.eco9.nm")

str(log_n2o_correlog_dat)
head(log_n2o_correlog_dat)
## Make individual plots, the following object is a list containing a ggplot for each
log_n2o_correlog_plots <- log_n2o_correlog_dat %>%
  split(.$ag.eco9.nm) %>% # splits the data frame by ag.eco9.nm, which creates a list with 9 data frames
  lapply(function(x) plot_correlog(ncf_correlog = x, # use the data.frame for each ag.eco9.nm to make a plot
                                   plot_title = NULL,
                                   n_col = "n", 
                                   mean_col = "mean",
                                   cor_col = "cor",
                                   xint_col = "xint",
                                   pval_col = "pval",
                                   name_col = "ag.eco9.nm",
                                   title_size = 14, 
                                   subtitle_size = 14, 
                                   axis_title_size = 14, 
                                   axis_text_size = 14))
## if hard to see writing run again but increase "title_size", "subtitle_size", "axis_title_size", "axis_text_size" as needed

## Look at each piece individually or scroll through
length(log_n2o_correlog_plots) # shows 9 items
# log_n20_correlog_plots
log_n2o_correlog_plots$CPL #or log_n20_correlog_plots[["CPL"]] or log_n2o_correlog_plots[[1]]
# log_n2o_correlog_plots$NAP
# log_n2o_correlog_plots$NPL
# log_n2o_correlog_plots$SAP
# log_n2o_correlog_plots$SPL
# log_n2o_correlog_plots$TPL
# log_n2o_correlog_plots$UMW
# log_n2o_correlog_plots$WMT
# log_n2o_correlog_plots$XER

## cowplot package (or other) can be used to group them into a matrix plot
log_n2o_correlog_plotgrid <- 
  cowplot::plot_grid(plotlist=log_n2o_correlog_plots, ncol=3)
print(log_n2o_correlog_plotgrid) # click 'Zoom' in plots quadrant, takes a few seconds to render
## can save the plot using 
# cowplot::ggsave(filename = "FILENAME.png", #saves a png version of the figure to the working folder
#                 plot = log_n20_correlog_plotgrid)
## can change .png to other forms
## can play with scale, width, and height to adjust the output of the figure saved to the png file

## alternatively, can make a facet plot since log_n20_correlog_dat is a dataframe in the correct format
names(log_n2o_correlog_dat) #"n"    "mean" "cor"  "xint" "pval" "ag.eco9.nm"

## create column to shape point, sig == TRUE means two-sided test < 0.05
log_n2o_correlog_dat$sig <- ifelse(log_n2o_correlog_dat$pval * 2 < 0.05,
                                   TRUE, FALSE) 
head(log_n2o_correlog_dat)
## can change ag.eco9.nm to a factor to arrange what order the plots are shown, otherwise default arranges them alphabetically
# log_n20_correlog_dat$ag.eco9.nm <- factor(log_n20_correlog_dat, 
#                                     levels = c("CPL", "NAP", "NPL", "SAP", "SPL", "TPL", "UMW", "WMT", "XER"))

## make facet plot
log_n2o_correlog_facetplot <- ggplot(log_n2o_correlog_dat) +
  geom_hline(yintercept = 0, color = "gray", linetype = "solid") +
  # geom_point(aes(x=mean, y=cor, shape=sig), size=I(2), color="black") +
  # geom_line(aes(x=mean, y=cor, group=1), size=I(0.5), color="black", linetype="solid") +
  lemon::geom_pointline(aes(x=mean, y=cor, shape=sig, group=1), 
                        size=I(2), color = "black", 
                        linetype = "solid", linesize = I(0.5), 
                        linecolor = "black") +
  scale_shape_manual(values=c("TRUE"=19, "FALSE"=1), 
                     labels=c("TRUE"="Sig.", "FALSE"="Not Sig."),
                     guide=FALSE) +
  facet_wrap(~ag.eco9.nm, ncol=3, scales="free") +
  labs(x="Distance (mean-of-class)", y="Correlation") +
  theme_bw() +
  theme(plot.title = element_text(size=14, hjust=0.5, 
                                  face="bold", color="black"),
        plot.subtitle = element_text(size=14, hjust=0.5, color="black"),
        axis.title = element_text(size=14, face="plain", color="black"),
        axis.text = element_text(size=14, color="black"),
        strip.text = element_text(size=14, color="black", face="bold"),
        strip.background = element_rect(fill="gray", color="black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
print(log_n2o_correlog_facetplot)
## can save the plot using 
# cowplot::ggsave(filename = "FILENAME2.png", #saves a png version of the figure to the working folder
#                 plot = log_n20_correlog_facetplot) 
## can change .png to other forms
## can play with scale, width, and height to adjust the output of the figure saved to the png file

```

## Semivariograms
Variograms display the variability among pairs of points at various separation distances.  I was unable to fit a model to the dissolved gas concentration variograms using the default fitting parameters in gstat.  The variograms suggest a lack of spatial structure, but I suspect the model parameterization can be improved.

```{r, warning=FALSE, message=FALSE}

# Variogram function requires sp object
dg.sp <- as(dg.sf, Class = "Spatial")

# CO2 variogram
co2.v <- variogram(dissolved.co2.umol~1, dg.sp)
co2.v.fit <- fit.variogram(co2.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
# co2.v.fit # chose sperical
co2.v.fit <- fit.variogram(co2.v, vgm("Sph")) # refit spherical, several warnings (?)

# CH4 variogram
ch4.v <- variogram(dissolved.ch4.umol~1, dg.sp)
ch4.v.fit <- fit.variogram(ch4.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
# ch4.v.fit # chose exp
ch4.v.fit <- fit.variogram(ch4.v, vgm("Exp")) # refit exponential, no convergence

# N2O variogram
n2o.v <- variogram(dissolved.n2o.nmol~1, dg.sp)
n2o.v.fit <- fit.variogram(n2o.v, vgm(c("Exp", "Mat", "Sph"))) # allow to choose best model.  Not sure which model warnings are associated with.
# n2o.v.fit # chose exp
n2o.v.fit <- fit.variogram(n2o.v, vgm("Exp")) # refit exponential, no convergence

# Plot variograms and models (if fit)
plot(co2.v, model = co2.v.fit, main = "CO2 semivariogram")
plot(ch4.v, main = "CH4 semivariogram")
plot(n2o.v, main = "N2O semivariogram")

```

## Random Semivariograms
Empirical semivariograms were calculated for:  pH, total nitrogen (ntl), total phosphorous (ptl), and chloropyll a (chla).  In addition, semivariograms were calculated for the dissolved gasses:  DO, N2O, CH4, and CO2. We downloaded Chapter 8 code from https://asdar-book.org/ for generating random variograms.  An initial empirical, or sample, variogram is calculated and plotted for pH, with that object saved.  Then pH values are randomly sampled and assigned to the coordinates so that a random semivariogram can be calculated.  That is done 100 times, and the gray lines of the random semivariograms are plotted with the blue line of the empirical semivariogram. Semivariogram for pH shows evidence of spatial autocorrelation out to ~ 500 km compared to randomized semivariograms.

The water quality variables of pH, tnl, tpl, and chla show strong evidence of spatial autocorrelation given the structure of their semivariograms versus randomized semivariograms.  The dissolved gasses, DO, N2O, CH4, and CO2 show none to modest spatial autocorrelation. The empirical semivariogram for surface dissolved oxygen is approximately a horizontal line, showing no evidence of spatial autocorrelation over the Euclidean distances that the lakes were sampled.  CH4 shows some autocorrelation to ~ 1000 km, but with no sill or leveling off the semivariance.
```{r, random_semivar}
names(dg)
dg_1 <- clean_names(dg)

# also includ ag.eco9.nm, ag.eco9.nm feature plus variables from Jake's 02/25/2020 email
# added ntl.lab.result, ptl.lab.result on 06/18/2020
# added physical variables, temperature and depth, on 06/23/2020
# added silica and calcium based on Leon-Palmero 2020 paper 06/24/2020
# made nitrate_n_result into binary variable nitrate_n_b b/c 1/4th of observation have 0.000 value 06/25/2020
# added ph_result 12/14/2020
names(dg_1)
# dropped onoffnet variable as I don't think it occurs in this version of the data frame, dg.2021-01-14.RData

dg_1 <- filter( dg_1, sample_source == "DG")
dg_1 <- select(dg_1,site_id:date_col, map_lat_dd, map_lon_dd, dissolved_co2_umol, dissolved_ch4_umol, dissolved_n2o_nmol, co2_sat_ratio, ch4_sat_ratio, n2o_sat_ratio, ag_eco9, ag_eco9_nm, nitrate_n_result, chla_result, ntl_result, ph_result, ptl_result,max_bf, temp_surf, temp_bottom, o2_surf, o2_bottom, lake_orgn, pstl_code , area_cat6, silica_result, calcium_result, visit_no) %>% filter(visit_no == 1 & ag_eco9 != "AK")

names(dg_1)
str(dg_1)

# To enable spatial analysis of the data, the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg_1$map_lon_dd, latitude = dg_1$map_lat_dd)

dg_sf1 <- st_as_sf(dg_1, coords = c("map_lon_dd", "map_lat_dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

st_crs(dg_sf1)
## pH
hist(dg_sf1$ph_result)

z_ph_sf1 = variogram(ph_result~1, dg_sf1)
plot(z_ph_sf1)
# plot(z_ph_sf1, plot.numbers = TRUE)
# plot(z_ph_sf1, ylab = "NLA ph Semivariance")
# plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters)", ylab = "NLA ph semivariance")
# z_ph_plot1 <- plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters) ")

print(xyplot(gamma ~ dist, z_ph_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ph_result)
                     v = variogram(random ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 0.8), xlab = 'distance', ylab = 'pH semivariance'
))

# used Cressie's robust variogram b/c skewness and not wanting to transform pH as it is already on a log scale
z_cressie_ph <- variogram(ph_result~1, dg_sf1, cressie = TRUE)
plot(variogram(ph_result~1, dg_sf1, cressie = TRUE))

print(xyplot(gamma ~ dist, z_cressie_ph, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ph_result)
                     v = variogram(random ~ 1, dg_sf1, cressie = TRUE)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 0.8), xlab = 'distance', ylab = 'pH semivariance'
))

## NTL
summary(dg_sf1$ntl_result) #no NA's

filter(dg_1,(!is.na(ntl_result))) %>% ggplot(., aes(ntl_result)) + geom_histogram()

# log transformation ~ unimodal & symmetric
filter(dg_1,(!is.na(ntl_result))) %>% ggplot(., aes(log(ntl_result))) + geom_histogram()

# hist(dg_sf1$ntl_result)
# hist(log(dg_sf1$ntl_result))

# default cutoff and width and no transformation
z_lntl_sf1 = variogram(log(ntl_result) ~ 1, dg_sf1)
plot(z_lntl_sf1, ylab = "NLA NTL Semivariance")
# set ylim for random semivariograms based on this plot

# cutoff and width from King et al. paper still not working even with tranform
z1_ntl_king = variogram(log(ntl_result) ~1, dg_sf1, cutoff=2500, width=20)
plot(z1_ntl_king)

# plot(z_ph_sf1, plot.numbers = TRUE)
# plot(z_ph_sf1, ylab = "NLA ph Semivariance")
# plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters)", ylab = "NLA ph semivariance")
# z_ph_plot1 <- plot(z_ph_sf1, cex = 0.5, type = "b", pch = 20, asp = 1, xlab = "Distance (meters) ")

print(xyplot(gamma ~ dist, z_lntl_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ntl_result)
                     v = variogram(log(random) ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 1.2), xlab = 'distance', ylab = 'NTL semivariance'
))

## PTL
summary(dg_sf1$ptl_result) # no NA's

filter(dg_sf1,(!is.na(ptl_result))) %>% ggplot(., aes(ptl_result)) + geom_histogram()

# log transformation ~ unimodal & symmetric
filter(dg_sf1,(!is.na(ptl_result))) %>% ggplot(., aes(log(ptl_result))) + geom_histogram()

# hist(log(dg_sf1$ntl_result))

z_lptl_sf1 = filter(dg_sf1, (!is.na(ptl_result))) %>% variogram(log(ptl_result)~1, .)
plot(z_lptl_sf1, ylab = "NLA TP Semivariance")
# set ylim of random semivariograms based on this plot
# plot(z_ph_sf1, plot.numbers = TRUE)

# data frame with no NA's
# dg_sf2 <- filter(dg_sf1, (!is.na(ptl_result)))

print(xyplot(gamma ~ dist, z_lptl_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$ptl_result)
                     v = variogram(log(random) ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 2), xlab = 'distance', ylab = 'TP semivariance'
))

## CHLA
summary(dg_sf1$chla_result) # no NA's

filter(dg_sf1,(!is.na(chla_result))) %>% ggplot(., aes(chla_result)) + geom_histogram()

# log transformation not that great but use for now
filter(dg_sf1,(!is.na(chla_result))) %>% ggplot(., aes(log(chla_result))) + geom_histogram()

# hist(log(dg_sf1$ntl_result))

z_lchla_sf1 = filter(dg_sf1, (!is.na(chla_result))) %>% variogram(log(chla_result)~1, .)
plot(z_lchla_sf1, ylab = "NLA CHLA Semivariance")
# set ylim of random semivariograms based on this plot
# plot(z_ph_sf1, plot.numbers = TRUE)

# data frame with no NA's
# dg_sf2 <- filter(dg_sf1, (!is.na(chla_result)))

print(xyplot(gamma ~ dist, z_lchla_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$chla_result)
                     v = variogram(log(random) ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 2.5), xlab = 'distance', ylab = 'CHLa semivariance'
))

## DO
summary(dg_sf1$o2_surf) # have 4 NA's

filter(dg_1,(!is.na(o2_surf))) %>% ggplot(., aes(o2_surf)) + geom_histogram()

# log probably too strong of transformation but use for now
filter(dg_1,(!is.na(o2_surf))) %>% ggplot(., aes(log(o2_surf))) + geom_histogram()

# hist(log(dg_sf1$ntl_result))

z_ldo_sf1 = filter(dg_sf1, (!is.na(o2_surf))) %>% variogram(log(o2_surf)~1, .)
# plot(z_ldo_sf1)
# plot(z_ph_sf1, plot.numbers = TRUE)
plot(z_ldo_sf1, ylab = "NLA Surface DO Semivariance")

# data frame with no NA's
dg_sf2 <- filter(dg_sf1, (!is.na(o2_surf)))

print(xyplot(gamma ~ dist, z_ldo_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf2$random = sample(dg_sf2$o2_surf)
                     v = variogram(log(random) ~ 1, dg_sf2)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 0.12), xlab = 'distance', ylab = 'NLA DO semivariance'
))

## N2O
summary(dg_sf1$dissolved_n2o_nmol) # no NA's

filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(dissolved_n2o_nmol)) + geom_histogram()

# unimodal, but not symmetrical
filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(log(dissolved_n2o_nmol))) + geom_histogram()

z_n2o_sf1 <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% variogram(log(dissolved_n2o_nmol) ~1, .)
plot(z_n2o_sf1, ylab = "N2O Semivariance")

# change cutoff and width
z_n2o = filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% variogram(log(dissolved_n2o_nmol)~1, ., cutoff = 750000, width = 25000)
# plot(z_n2o, ylab = "N2O Semivariance", plot.numbers = TRUE)
plot(z_n2o, ylab = "NLA N2O Semivariance")

print(xyplot(gamma ~ dist, z_n2o, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
    panel = function(x, y, ...) {
        for (i in 1:100) {
            dg_sf1$random = sample(dg_sf1$dissolved_n2o_nmol)
            v = variogram(log(random) ~ 1, dg_sf1, cutoff = 750000, width = 25000)
            llines(v$dist, v$gamma, col = 'grey')
        }
        panel.xyplot(x, y, ...)
    },
    ylim = c(0, 0.35), xlab = 'distance', ylab = 'N2O semivariance'
))

## CH4
summary(dg_1$dissolved_ch4_umo) # no NA's

filter(dg_1,(!is.na(dissolved_ch4_umol))) %>% ggplot(., aes(dissolved_ch4_umol)) + geom_histogram()

# log transformation looks good, ~ unimodal & symmetric
filter(dg_1,(!is.na(dissolved_ch4_umol))) %>% ggplot(., aes(log(dissolved_ch4_umol))) + geom_histogram()

z_lch4_sf1 = filter(dg_sf1, (!is.na(dissolved_ch4_umol))) %>% variogram(log(dissolved_ch4_umol)~1, .)
plot(z_lch4_sf1, ylab = "NLA CH4 Semivariance")
# change ylim below in random semivariogram based on this plot
# plot(z_ph_sf1, plot.numbers = TRUE)

# data frame with no NA's for random sampling below
# dg_sf2 <- filter(dg_sf1, (!is.na(dissolved_ch4_umol)))

print(xyplot(gamma ~ dist, z_lch4_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$dissolved_ch4_umol)
                     v = variogram(log(random) ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 2), xlab = 'distance', ylab = 'NLA CH4 semivariance'
))

## CO2
summary(dg_1$dissolved_co2_umo) # no NA's

filter(dg_1,(!is.na(dissolved_co2_umol))) %>% ggplot(., aes(dissolved_co2_umol)) + geom_histogram()

# log transformation, unimodal and symmetric
filter(dg_1,(!is.na(dissolved_co2_umol))) %>% ggplot(., aes(log(dissolved_co2_umol))) + geom_histogram()

# hist(log(dg_sf1$ntl_result))

z_lco2_sf1 = filter(dg_sf1, (!is.na(dissolved_co2_umol))) %>% variogram(log(dissolved_co2_umol)~1, .)
plot(z_lco2_sf1, ylab = "NLA CO2 Semivariance")
# change ylim in random semivariogram based on this plot
# plot(z_ph_sf1, plot.numbers = TRUE)

# data frame with no NA's
# dg_sf2 <- filter(dg_sf1, (!is.na(dissolved_co2_umol)))

print(xyplot(gamma ~ dist, z_lco2_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
             panel = function(x, y, ...) {
                 for (i in 1:100) {
                     dg_sf1$random = sample(dg_sf1$dissolved_co2_umol)
                     v = variogram(log(random) ~ 1, dg_sf1)
                     llines(v$dist, v$gamma, col = 'grey')
                 }
                 panel.xyplot(x, y, ...)
             },
             ylim = c(0, 1.2), xlab = 'distance', ylab = 'NLA CO2 semivariance'
))

```

## Random Ecoregional Semivariogram N2O for ASLO Abstract
I used ecoregional-specific data frames to feed into random semivariogram function.  
```{r, ecoregion_semivar}

## N2O
summary(dg_sf1$dissolved_n2o_nmol) # no NA's

filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% ggplot(., aes(dissolved_n2o_nmol)) + geom_histogram()

# unimodal, but not symmetrical
filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="XER") %>%  ggplot(., aes(log(dissolved_n2o_nmol))) + geom_histogram()

# zcpl <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="CPL")
# zcpl_variog <- variogram(log(dissolved_n2o_nmol) ~1, zcpl)
# plot(zcpl_variog, ylab = "N2O zCPL Semivariance")
# the 2-step process above matches the piped process below
xer_n2o_sf1 <- filter(dg_sf1,!is.na(dissolved_n2o_nmol)) %>%
  filter(ag_eco9 =="XER") %>% 
  variogram(log(dissolved_n2o_nmol) ~1, .)
plot(xer_n2o_sf1, ylab = "N2O XER Semivariance")

xer <- filter(dg_sf1, (!is.na(dissolved_n2o_nmol))) %>% filter(ag_eco9 =="XER")

print(xyplot(gamma ~ dist, wmt_n2o_sf1, pch = 3, type = 'b', lwd = 2, col = 'darkblue',
    panel = function(x, y, ...) {
        for (i in 1:100) {
            xer$random = sample(xer$dissolved_n2o_nmol)
            v = variogram(log(random) ~ 1, xer)
            llines(v$dist, v$gamma, col = 'grey')
        }
        panel.xyplot(x, y, ...)
    },
    ylim = c(0, 0.35), xlab = 'distance', ylab = 'XER N2O semivariance'
))

```

## Linked micromaps
The ecoreg_stats summary is for population estimates.  The code at the bottom, ecoreg_pred, I believe gives summary stats for the predicted distributions, but I would double-check with Roy. There is a quick-and-dirty micromap code used to look trying linked micromap plots on different variables.  When you settle on the variables you want to use then the full micromap code produces the pretty linked micromap plots.
```{r, linked_micromap}
# Read data object containing predictions for population
Sys.time()
load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
Sys.time()

# summarize means by WSA9 (posterior median for mean and 95% credible intervals)
# this is analogous to spsurvey "population estimates"
ecoreg_stats <- all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise( mean_n2o = mean(preds_n2o),
             mean_n2oeq = mean(preds_n2oeq),
             mean_sat = mean(preds_sat))%>%
  summarise( post_med_n2o = round(median(mean_n2o), 1),
    LCI_n2o = round(quantile(mean_n2o, probs = 0.025), 1),
    UCI_n2o = round(quantile(mean_n2o, probs = 0.975), 1),
     post_med_n2oeq = round(median(mean_n2oeq), 2),
    LCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.025), 2),
    UCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.975), 2),
     post_med_sat = round(median(mean_sat), 2),
    LCI_sat = round(quantile(mean_sat, probs = 0.025), 2),
    UCI_sat = round(quantile(mean_sat, probs = 0.975), 2),
    n = dplyr::n())

ecoreg_stats <- as.data.frame(ecoreg_stats)
class(ecoreg_stats)
str(ecoreg_stats) # WSA9 is a factor, but some of the stats have attribute of Named num

ecoreg_stats$WSA9 <- as.character(ecoreg_stats$WSA9)
str(ecoreg_stats)

ecoreg_stats

ecoregions <- st_read(paste0(localPath, "\\Environmental Protection Agency (EPA)\\",
                             "ORD NLA17 Dissolved Gas - Documents\\inputData\\aggr_ecoregions_simple.shp"))

wsa9names <- st_set_geometry(ecoregions, NULL) %>%
  select(WSA9, WSA9_NAME)

ecoreg_stats <- left_join(ecoreg_stats, wsa9names)

ecoregions <- st_transform(ecoregions, 5070)
# make spatial polygon dataframe need for micromap
ecoregions_sp1 <- as(ecoregions, "Spatial")

# convert to map_table for linked micromap
ecor.polys <- create_map_table(ecoregions_sp1, 'WSA9')
class(ecor.polys)
str(ecor.polys, max =2) # ID is chr "CPL"...

# Quick & Dirty micromap code:  use this to test out micromap plot of different summary statistics to get their x-axis values
mmplot(stat.data=ecoreg_stats,
       map.data=ecor.polys,
       panel.types=c('dot_legend', 'labels', 'dot_cl', 'dot_cl', 'map'),
       panel.data=list(
         NA,
         'WSA9_NAME',
         list('post_med_n2o', 'LCI_n2o', 'UCI_n2o'),
         list('post_med_sat', 'LCI_sat', 'UCI_sat'),
         NA),
       ord.by='post_med_n2o', grouping=3,
       median.row=FALSE,
       map.link=c('WSA9', 'ID'))

# Full code:  Use this to produce final linked micromaps of the variables you want
# export as pdf US letter landscape from Plots pane
# export as PNG Width 900 Height 968, maintain aspect ratio
tiff(filename = "output/figures/mmplot.tiff",
     res = 800, width = 9.00, height = 9.68, units = "in" )

mmplot(stat.data=ecoreg_stats,
       map.data=ecor.polys,
       panel.types=c('dot_legend', 'labels', 'dot_cl', 'dot_cl', 'map'),
       panel.data=list(
         NA,
         'WSA9_NAME',
         list('post_med_n2o', 'LCI_n2o', 'UCI_n2o'),
         list('post_med_sat', 'LCI_sat', 'UCI_sat'),
         NA),
       ord.by = 'post_med_n2o', rev.ord = TRUE, grouping=3,
       median.row=FALSE,
       map.link=c('WSA9', 'ID'),

       plot.height=7, plot.width=7, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
       map.color2='lightgray',

       panel.att=list(list(1, point.type=20, point.size=2,
                           graph.border.color='white',
                           xaxis.text.display=F, xaxis.line.display=F,
                           graph.grid.major=F),

                      list(2, header='NARS Reporting\n Regions', panel.width=1.25,#increased panel.width to 1.1 from .8
                           align='left', text.size=1),#increased text.size to .8 from .6

                      list(3, header='N2O Concentration\n with 95% CI',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=list(6.0, 7.0, 8.0, 9.0, 10.0),
                           xaxis.labels=list(6.0, 7.0, 8.0, 9.0, 10.0),
                           xaxis.title='N2O mg/L', panel.width=1.25),

                      list(4, header='N2O Saturation\n with 95% CI',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=list(0.8, 1.0, 1.2, 1.4, 1.6),
                           xaxis.labels=list(0.8, 1.0, 1.2, 1.4, 1.6),
                           xaxis.title='N2O Saturation, %',panel.width=1.25),

                      list(5, header='Light Gray Means\n Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                           panel.width=2)))

dev.off()

# To get ecoregional distributions of the predictions I think this is the code we would use based on our conversation with Roy, but I would double-check with him.

ecoreg_preds <- all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise( post_med_n2o = round(median(mean_n2o), 1),
    LCI_n2o = round(quantile(mean_n2o, probs = 0.025), 1),
    UCI_n2o = round(quantile(mean_n2o, probs = 0.975), 1),
     post_med_n2oeq = round(median(mean_n2oeq), 2),
    LCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.025), 2),
    UCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.975), 2),
     post_med_sat = round(median(mean_sat), 2),
    LCI_sat = round(quantile(mean_sat, probs = 0.025), 2),
    UCI_sat = round(quantile(mean_sat, probs = 0.975), 2),
    n = dplyr::n())
```
# Exploratory analysis of N2O data
The mean N2O saturation ratio for the country is `r pop.eco.sat %>% filter(type == "National", subpopulation == "National", indicator == "N2O_SAT_RATIO") %>% select(estimate) %>% pull() %>% round(2)` (95% CI = `r pop.eco.sat %>% filter(type == "National", subpopulation == "National", indicator == "N2O_SAT_RATIO") %>% select(lcb95pct) %>% pull() %>% round(2)` - `r pop.eco.sat %>% filter(type == "National", subpopulation == "National", indicator == "N2O_SAT_RATIO") %>% select(ucb95pct) %>% pull() %>% round(2)`), suggesting that US waterbodies are a new N2O sink during the NLA index period.  This finding is consistent with a recent report that 67% of 101 agricultural ponds in Canada were undersaturated in N2O during the summer (Webb et al. 2019, PNAS).

There are several sources of N2O in surface waters.  The overlying atmosphere contains trace quantities of N2O (~300 ppb) that dissolve into surface waters as a function of barometric pressure and water temperature.  N2O is also produced and consumed in aquatic ecosystems via denitrification.  Denitrification is the reduction of oxidized forms of N (i.e. NO3) to more reduced forms (i.e. N2).  It is a form of anaerobic respiration where oxidized forms of N serve as the electron acceptor in the absence of more energetically favorable alternatives, such as oxygen.  N2O concentrations/fluxes are therefore somewhat dependent upon the availability of oxidized N and low dissolved oxygen.  N2O can also be produced via nitrification.  Nitrification is the aerobic oxidation of NH4 to more oxidized forms.  N2O can be produced as a byproduct during the process and rates are somewhat dependent upon the availability of ammonium.  See Beaulieu et al. 2011, 2014, and 2015 for more information.

The effect of denitrification and nitrification on dissolved N2O concentrations near the air-water interface is somewhat mediated by hydrology.  Denitrification is an anaerobic process that occurs predominantly in the sediment or anoxic hypolimnion.  Therefore, N2O must advect or diffuse from the sediment/hyoplimion to the water surface before it can affect dissolved N2O concentration near the air/water interface (where NLA samples were collected).  Previous invesitgations have interpretted correlations between N2O and lake size and degree of thermal stratification as indicative of hydrologic control (DelSontro et al. 2019, Webb et al. 2019).

DelSontrol et al. (2019) reported that N2O emission rates were positively related to chl a.  The authors (inlcuding me, I'm one of the 'authors'!) speculate that chla is a surrogate for labile carbon.  Denitrification requires carbon, so perhaps high rates of denitrification-N2O production occur in the presence of abundant carbon availability.

We do not yet have depth profiles of DO or temperature.  When these data become available, we can include bottom water DO and strength of thermal stratification in the model.  For now, we can explore correlations with the following variables.

* MeanDUsed - mean depth  
* ph.result - pH  
* nitrate.n.result - NO3-N (mg/L)  
* ntl.lab.result - total nitrogen (mg/L)  
* chla.result.vol - chlorophyll a (ug/L)  
* ptl.lab.result - total phosphorus (mg/L)  
* surftemp - surface water temperature (C)

A correlation matrix of dissolved N2O concentration and potential controlling variables doesn't show any striking relationships, though the strongest relationship is with nitrate.

```{r}
library(corrr)
dg_1 %>%
  select(index_site_depth, dissolved_n2o_nmol, ph_result, nitrate_n_result, ntl_result, chla_result, ptl_result, temp_surf) %>%
  correlate()
```

Categorization and regression analysis (CART), which can accomodate non-linear and interactive relationships, also identified nitrate as the most important predictor variable.  The CART analysis did not identify splits associated with ecoregion.

```{r}
library(tree)
model <- tree(dissolved.n2o.nmol ~ MeanDUsed + nitrate.n.result + ntl.lab.result + chla.result.vol + ptl.lab.result + surftemp + ag.eco9.nm, data = dg)

plot(model)
text(model)
```
# ASLO 2021

## Site Map
```{r}
# Load data
 load(paste0(localPath,
            "/Environmental Protection Agency (EPA)/",
            "ORD NLA17 Dissolved Gas - Documents/",
            "inputData/dg.2021-02-01.RData"))

# To enable mapping the dataframe will be converted to a 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple")

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# Load state boundaries
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

ggplot() +
  geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
  geom_sf(data = dg.sf) +
  geom_sf(data = states, fill = NA, color = "white", size = 0.1) +
  theme_void() +
  theme(legend.title = element_blank())

ggsave("output/figures/siteMap.tiff")

```

## Population Data

```{r}
# 2/8/2023 75sec on VM
tic()
load(paste0(localPath, "\\Environmental Protection Agency (EPA)\\ORD NLA17 Dissolved Gas - Documents\\inputData\\all_predictions.rda"))
toc()
```

## Explore structure of all_predictions object
```{r}
head(all_predictions)
```
.row uniquely identifies each waterbody.  There are `r all_predictions %>% distinct(.row) %>% nrow()` waterbodies in the sample frame.  `.draw` reflects the simulation number for each waterbody. I think this is a 'draw from the posterior distribution'.  The data contains `r all_predictions %>% distinct(.draw) %>% nrow()` draws for each waterbody.

The `size_cat` variable largely uses the NLA size category definitions.  This may not be the best for showing N2O fluxes by lake size bins, however.  The >50 ha (0.5km2) has most of the surface area

```{r}
# total area by size_cat bar chart
all_predictions %>%
  filter(.draw == 1) %>%
  group_by(WSA9, size_cat) %>%
  summarise(tot_area = sum(area_ha)) %>%
  ggplot(aes(x=size_cat, y=tot_area)) +
  geom_bar(stat = "identity") +
  facet_wrap(~WSA9)

```

Lets create a new size category with more even surface area distributions.

```{r}
# 4 minutes on VM. pushes memory up to 65GB [2/8/2023]
tic()
all_predictions <- all_predictions %>%
  #filter(.draw == 1) %>% # testing
  mutate(area_km2 = area_ha / 100, # this works
         size_cat2 = ifelse(area_km2 < 0.1,
                            "<0.1",
                            ifelse(area_km2 >= 0.1 & area_km2 < 1,
                                   "0.1 - 1",
                                   ifelse(area_km2 >=1 & area_km2 <10,
                                          "1 - 10",
                                          ifelse(area_km2 >=10 & area_km2 <100,
                                                 "10 - 100",
                                                 ifelse(area_km2 >=100, 
                                                        ">100",
                                                        NA))))),
         size_cat2 = factor(size_cat2, ordered = TRUE, 
                            levels = c("<0.1", "0.1 - 1", "1 - 10", "10 - 100", ">100"))) %>%
  select(-area_km2)
toc()
gc()

# total area by size_cat2 bar chart
all_predictions %>%
  filter(.draw == 1) %>%
  group_by(WSA9, size_cat2) %>%
  summarise(tot_area = sum(area_ha)) %>%
  ggplot(aes(x=size_cat2, y=tot_area)) +
  geom_bar(stat = "identity") +
  facet_wrap(~WSA9)

```

## Reorder WSA9
Reorder WSA9 to correspond with order of first micromap produced below, which is ordered by extent of N2O saturation.
```{r}
all_predictions <- all_predictions %>%
  mutate(WSA9 = factor(WSA9,
                       levels = c("NPL", "TPL", "UMW", 
                                  "CPL", "SPL", "SAP",
                                  "XER", "NAP", "WMT"),
                       labels = c("Northern Plains", "Temperate Plains",
                                  "Upper Midwest", "Coastal Plains", "Southern Plains", 
                                  "Southern Appalachians", "Xeric", "Northern Appalachians", 
                                  "Western Mountains")))

```

## N2O Concentration
I think we should start with some plots of the primary measurement value.
```{r}

# PLOT N2O CONCENTRATION BY WSA9 AND LAKE SIZE

# N2O CONCENTRATION BY ECOREGION AND SIZE
# point and linerange
all_predictions %>%
  group_by(WSA9, size_cat, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_n2o = mean(n2o)) %>% # 500 means for each WSA9
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_n2o), 3), 
    LCL = round(quantile(mean_n2o, probs = 0.025), 3),
    UCL = round(quantile(mean_n2o, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(-WSA9) %>% # necessary?
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = size_cat, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  ylab("mean N2O concentration (nmol/L)") +
  coord_flip() + 
  theme_bw() +
  facet_wrap(~ecoregion)
ggsave("output/figures/n2oConcentrationByEcoSize.tiff")

# N2O CONCENTRATION BY ECOREGION
# point and linerange
all_predictions %>%
  group_by(WSA9, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_n2o = mean(n2o)) %>% # 500 means for each WSA9
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_n2o), 3), 
    LCL = round(quantile(mean_n2o, probs = 0.025), 3),
    UCL = round(quantile(mean_n2o, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  ylab("mean N2O concentration (nmol/L)") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/n2oConcentrationByEco.tiff")
```

## N2O saturation ratio: continuous variable
N2O saturation ratio in the sample data has a long right tail, with a maximum value of ~30.  The distribution of the values predicted to the population also has a long right tail, but features more extreme values (>600).
```{r}
# PLOT N2O SAT RATIO BY WSA9------------------

# point and linerange: mean +/- 95%CI
all_predictions %>%
  group_by(WSA9, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_sat = mean(n2osat)) %>% # 500 means for each WSA9
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_sat), 3), 
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/n2oSatRatioByEco.tiff")

# Box and whisker
all_predictions %>%
  filter(.draw == 1) %>%
  ggplot( aes( x = WSA9, y = n2osat ) ) +
  geom_boxplot() +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/n2oSatRatioByEcoBoxWhisker.tiff")

# point and linerange: mean AND MEDIAN.+/- 95%CI
all_predictions %>%
  group_by(WSA9, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_sat = mean(n2osat),
            median_sat = median(n2osat)) %>% # 500 means for each WSA9
  # now summarize to 1 statistic per WSA9
  summarise(estimate = round(median(mean_sat), 3), 
            LCL = round(quantile(mean_sat, probs = 0.025), 3),
            UCL = round(quantile(mean_sat, probs = 0.975), 3),
            estimate_md = round(median(median_sat), 3), 
            LCL_md = round(quantile(median_sat, probs = 0.025), 3),
            UCL_md = round(quantile(median_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  #select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot(aes( x = ecoregion, y = estimate )) +
  geom_point(color = "blue") +
  geom_linerange( aes( ymin = LCL, ymax = UCL), color = "blue") +
  geom_point(aes(x = ecoregion, y = estimate_md), color = "red") +
  geom_linerange( aes( ymin = LCL_md, ymax = UCL_md), color = "red") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/n2oSatRatioByEcoMedianAndMean.tiff")

# density plots of predicted distribution, mean, and median
# using xlim to zoom in on values close to 1

dummy <- all_predictions %>%
  group_by(WSA9) %>%
  summarize(mean = mean(n2osat),
            median = median(n2osat)) %>%
  mutate(dens.med = c(2, 2.5, 2.3, 2.05, 1.75, 2.15, 2.2, 1.9, 1.9),
         dens.mean = c(.4, 1.9, 2, .6, .65, .45, 1.5, 1.45, 0.9),
         .draw = 1) # needed to match grouping aesthetic in ggplot call

tic() # 40 seconds for 100 draws
all_predictions %>%
  filter(.draw %in% 1:100) %>% # subset for practice
  #mutate(ecoregion = factor(WSA9)) %>%
  #mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
  ggplot(aes(x = n2osat, group = .draw, color = .draw ) ) +
  geom_density(show.legend = FALSE) +
  geom_vline(xintercept = 1, color = "red") +
  geom_segment(data = dummy,
               aes(x = mean, xend = mean, y = 0,
                   yend = dens.mean),
               linetype = "solid") +
  geom_segment(data = dummy,
               aes(x = median, xend = median, y = 0,
                   yend = dens.med),
               linetype = "dashed") +
  ylab ("density") + 
  xlab(expression(N[2]*O~saturation~ratio)) +
  xlim(0,3) + # data range is 0-600, but sample max is 30
  facet_wrap(~WSA9) +
  theme_bw() +
  theme(legend.position = "none")
toc()
ggsave("output/figures/n2oSatRatioDensityByEcoWithMeanMedian.tiff", width = 8.5, height = 5)

# density plots of predicted distribution
# using xlim to zoom in on values close to 1
p <- all_predictions %>%
  filter(.draw %in% 1:10) %>% # subset for practice
  #mutate(ecoregion = factor(WSA9)) %>%
  #mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
  ggplot(aes(x = n2osat, group = .draw, color = .draw ) ) +
  geom_density(show.legend = FALSE) +
  geom_vline(xintercept = 1, color = "red") +
  xlab(expression(N[2]*O~saturation~ratio)) +
  xlim(0,5) + # data range is 0-600, but sample max is 30
  facet_wrap(~WSA9) +
  theme_bw()

# this bit fills strip with same colors used in micromap (see below)
#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
g <- ggplot_gtable(ggplot_build(p))
stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
k <- 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid::grid.draw(g)

ggsave("output/figures/n2oSatDensitySmallRange.tiff", g, width = 7.93, height = 4.56)

summary(dg$n2o.sat.ratio) # max value in sample is 33

# rplot, but show full x-axis range
# include RUG to better show extreme values
p <- all_predictions %>%
  filter(.draw == 1) %>% # need only one draw, otherwise RUG is misleading
  #mutate(ecoregion = factor(WSA9)) %>%
  #mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
  ggplot( aes( x = n2osat ) ) +
  geom_density(color = "blue") +
  geom_vline(xintercept = 1, color = "red") +
  geom_rug() +
  xlab(expression(N[2]*O~saturation~ratio)) +
  #xlim(0,35) + # data range is 0-600, but sample max is 30
  facet_wrap(~WSA9, scales = "free") +
  theme_bw()

# this bit fills strip with same colors used in micromap (see below)
#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
g <- ggplot_gtable(ggplot_build(p))
stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
k <- 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid::grid.draw(g)

ggsave("output/figures/n2oSatDensityFullRange.tiff", g)

# Take a peak at values >40 in predictions
all_predictions %>%
    filter(n2osat > 40, .draw == 1) %>% # look at one draw
  {table(.$WSA9)} # >40 is uncommon outside of CPL

# # density plots of observed distribution
# huge right tail
dg %>%
  mutate(ecoregion = factor(ag.eco9.nm)) %>%
  mutate(ecoregion = fct_reorder(ecoregion, n2o.sat.ratio)) %>%
  ggplot( aes( x = n2o.sat.ratio ) ) +
  geom_density() +
  xlab("mean N2O saturation ratio") +
  facet_wrap(~ecoregion) +
  theme_bw()

# PLOT N2O SAT RATIO BY SIZE------------------
# point and linerange
all_predictions %>%
  group_by(size_cat, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_sat = mean(n2osat)) %>% # 500 means for each WSA9
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_sat), 3), 
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
  ggplot( aes( x = size_cat, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw() 
ggsave("output/figures/n2oSatRatioByEco.tiff")

# PLOT N2O SAT RATIO BY WSA9 AND SIZE------------------
# point and linerange
all_predictions %>%
  group_by(WSA9, size_cat, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_sat = mean(n2osat)) %>% # 500 means for each WSA9
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_sat), 3), 
    LCL = round(quantile(mean_sat, probs = 0.025), 3),
    UCL = round(quantile(mean_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  #select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = size_cat, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 1, color='blue') +
  ylab("mean N2O saturation ratio") +
  coord_flip() + 
  theme_bw() +
  facet_wrap(~ecoregion)
ggsave("output/figures/n2oSatRatioByEcoAndSize.tiff")

# PLOT N2O* BY SIZE------------------
# point and linerange
all_predictions %>%
  mutate(n2ostar = abs(n2o - n2oeq)) %>%
  group_by(size_cat, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_star = mean(n2ostar)) %>% # 500 means for each WSA9
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_star), 3), 
    LCL = round(quantile(mean_star, probs = 0.025), 3),
    UCL = round(quantile(mean_star, probs = 0.975), 3)) %>% 
  ggplot( aes( x = size_cat, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  ylab("absolute[mean N2O - equilibrium N2O] (nmol/L)") +
  coord_flip() + 
  theme_bw() 
ggsave("output/figures/n2oStarBySize.tiff")

```

## N2O saturation ratio: extent estimate
```{r}
# EXTENT OF WATERBODIES FUNCTIONING AS SINKS BY ECOREGION

# density plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  # below summarise generates 500 proportions per WSA
  summarise(prop_sat = sum(n2osat < 1) / length(unique(.row))) %>%
  ggplot(aes(x=prop_sat)) + 
  geom_density() +
  facet_wrap(~WSA9) +
  xlab("Proportion of waterbodies functioning as N2O sinks") +
  theme_bw()
ggsave("output/figures/extentDensityPlot.tiff")

# point and line plot
all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% # Roy had length(unique(.row)) here, but the unique isn't necessary.  this object is grouped by WSA9
  summarise(estimate = round(median(prop_sat), 3),
    LCL = round(quantile(prop_sat, probs = 0.025), 3),
    UCL = round(quantile(prop_sat, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL )) +
  ylab("Proportion of waterbodies functioning as N2O sinks") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/extentPointLinerange.tiff")

```

## N2O emission rate and flux
### Calculations
Roy predicted N2O equilibrium and concentration to the population.  I used ERA5 to calculate mean wind speed during the index period for each lake in the population.  I need to match wind speed to Roy's predictions.  Waterbodies in Roy's predictions are identified by '.row'.  There are `r all_predictions %>% distinct(.row) %>% summarize(population = n()) %>% pull()` rows.

```{r calculate_emissions}

# Read in wind data
# Mean index period wind speed was calculated on VM via 
# nlaSampleFrameWind.R.  The resulting data object was written out and read in below.
wind <- read.table(paste0(Sys.getenv("USERPROFILE"), 
                                    "/Environmental Protection Agency (EPA)/",
                                    "ORD NLA17 Dissolved Gas - Documents/",
                                    "inputData/windSampleFrame.txt"))
# print first 6 lines
wind %>% head()

# print first 6 .row values of all_predictions
all_predictions %>%
  filter(.row %in% 1:6) %>% # each row is a unique waterbody
  group_by(.row) %>%
  slice(1) # grabs first record

# Row order in wind and all_predictions is inherited from sample frame.  Thus row 1
# of wind corresponds to .row == 1 in all_predictions.  This is a sketchy way to merge
# things, but will do for now.

# add .id to wind which is simply row number
wind <- wind %>% mutate(.id = row_number())

# Merge wind and all_predictions
tic() # 21 seconds on VM
all_predictions <- full_join(all_predictions, wind, by = c(".row" = ".id")) # 40 seconds on laptop
toc()
dim(wind) # 465897
dim(all_predictions) #232948500, 500 observations for each lake
465897*500 #232948500

# ESTIMATE GAS EXCHANGE RATE FOR POPULATION

# Using gas exchange model from:
# Vachon, D., and Y. T. Prairie (2013), The ecosystem size and shape 
# dependence of gas transfer velocity versus wind speed relationships in 
# lakes, Can. J. Fish. Aquat. Sci., 70(12), 1757-1764, doi:10.1139/cjfas-2013-0241.
# Table 2, model B.

# define k600 prediction function
pred.k600 <- function(ws, area) { 
  # predict k600 in cm h-1
  # ws = wind speed (m s-1)
  # area = lake area (km2)
  2.51 + (1.48 * ws) + (0.39 * ws * log10(area))
}

all_predictions <- all_predictions %>%
  mutate(k600.cm.h = pred.k600(ws = ws, area = area_ha/100), # 100ha = 1km2
         # 1000 L to m3. 44ng to 1nmol. ng->ug. ug->mg
         e.n2o.mg.m2.d = (((n2o - n2oeq) * 1000 * 44) / (1000 * 1000)) * 
           (k600.cm.h * (24/100)), #h->day. cm->m
         f.n2o.Mg.y = (e.n2o.mg.m2.d * 365 * (area_ha*10000)) / # day to year. ha to m2.
           (1000*1000*1000)) #mg-g, g-kg, kg-Mg

```
### compared to literature and US GHG inventory
The distribution of N2O emission rates (mg N2O m-2 d-1) for one draw is very similar to that derived from a literature review of lake/reservoir emission rates (DelSontro et al), except for CPL which shows some exceedingly high rates.  Overall, our results look reasonable.

The total N2O flux from US lakes and reservoirs is minor compared to other anthropogenic N2O sources
```{r}
# Compare to literature
# Read data from Webb et al. 2019
webb <- readr::read_csv(paste0(Sys.getenv("USERPROFILE"), 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "literature/n2o/Webb et al. 2019 PNAS/",
                               "SI_data.csv")) %>%
  mutate(e.n2o.mg.m2.d = N2O.umol.m2.d * (44/1000), # umol->mg
         source = "webb")

# Read data from Deemer et al. 2016
deemer <- readxl::read_xls(paste0(Sys.getenv("USERPROFILE"), 
                               "/Environmental Protection Agency (EPA)/",
                               "ORD NLA17 Dissolved Gas - Documents/",
                               "literature/n2o/Deemer et al. 2016 Bioscience SI.xls"),
                           sheet = "Supplementary_Reservoir_GHG_Dat",
                           skip = 1) %>%
  mutate(e.n2o.mg.m2.d = `mg N2O-N m-2 d-1` * (44/28),
         source = "deemer")

# combine NLA and lit data
all_dat <- rbind(all_predictions %>% 
  filter(.draw == 1) %>% # just one iteration
  mutate(source = "nla") %>%
  select(source, e.n2o.mg.m2.d),
  webb %>% select(source, e.n2o.mg.m2.d),
  deemer%>% select(source, e.n2o.mg.m2.d))

# dot plot of N2O emission rates by source
ggplot(all_dat, aes(source, e.n2o.mg.m2.d)) +
  geom_point(position = "jitter")

# summary statistics by source
all_dat %>%
  split(f = list(.$source)) %>% # split into named lists
  map(~ summary(.x)) # apply summary function to each list element

## Comparison to US GHG inventory
# top sources from inventory report in kt / year
n2o <- data.frame(kt = c(1156,88,84,66,60,34,18,14,7),
                     source = c("Agricultural Soil Management", # this includes leaching and runoff
                                "Wastewater Treatment",
                                "Stationary Combustion",
                                "Manure Management",
                                "Mobile Combustion",
                                "Nitric Acid Production",
                                "Adipid Acid Production",
                                "N2O from Product Uses",
                                "Composting"))

n2o <- n2o %>% mutate(Mg = (kt * 1000 * 1000) / 1000,  # 1000 metric tons in a kt, 1000 kg in a metric ton, 1000 kg in a Mg
                      Gg = Mg / 1000)

# National scale number for lakes/reservoirs (this is is Gg)
national.lakes <- all_predictions %>% 
  group_by(.draw) %>% 
  summarise(total.n2o = sum(f.n2o.Mg.y)) %>% 
  summarise(Gg = round(mean(total.n2o), 3) / 1000, # divide by 1000 for Mg to Gg
            Gg_LCL = round(quantile(total.n2o, probs = 0.025), 3) / 1000, # divide by 1000 for Mg to Gg
            Gg_UCL = round(quantile(total.n2o, probs = 0.975), 3) / 1000) %>% # divide by 1000 for Mg to Gg
  mutate(source = "Lakes and Reservoirs")

# Merge lakes and inventory category
inventory <- full_join(n2o, national.lakes) %>%
  arrange(desc(Gg))
inventory

inventory %>% mutate(source = as.factor(source),
                     source = reorder(source, Gg)) %>%
  ggplot(aes(source, Gg)) +
  geom_bar(stat = "identity", color = "orange", fill = "orange") +
  geom_errorbar(aes(x=source, ymin = Gg_LCL, ymax = Gg_UCL), width = 0.5) +
  coord_flip() +
  ylab(expression(N[2]*O~flux~(Gg~year^{-1}))) +
  theme_bw() +
  theme(axis.title.y = element_blank())
ggsave(filename = "output/figures/n2oInventory.tiff", width = 9, height = 5.5)

tiff("output/figures/n2oInventoryPie.tiff", 
     type = "cairo", bg = "transparent") # for transparent background
pie.labels <- c("Agricultural Soil Management",
                "Wastewater Treatment",
                "Stationary Combustion",
                "Manure Management",
                "Mobile Combustion",
                "Nitric Acid Production",
                NA,
                NA,
                NA,
                "Lakes and Reservoirs")
pie(inventory$Gg, labels = pie.labels, col = brewer.pal(8, "Set2"), radius = 0.6, cex = 0.9)
dev.off()

(national.lakes$Gg/sum(inventory$Gg)) * 100 #0.07%
```

### Figures and tables
Below is the distribution of total annual fluxes from the lakes.  This is calculated as lake area * emission rate.

```{r}
# EMISSION RATES---------
# density plot by ecoregion
all_predictions %>%
  filter(.draw %in% 1:10) %>% # subset for practice
  #mutate(ecoregion = factor(WSA9)) %>%
  #mutate(ecoregion = fct_reorder(ecoregion, preds_sat)) %>%
  ggplot(aes(x = e.n2o.mg.m2.d, group = .draw, color = .draw )) +
  geom_density(show.legend = FALSE) +
  #geom_vline(xintercept = 0, color = "red") +
  xlab(expression("N"[2]*"O"~ "emission"~ "rate" ~ "(mg" ~ N[2]*O ~ m^-2 ~ d^-1 ~ ")")) +
  #xlim(0,5) + # data range is 0-600, but sample max is 30
  facet_wrap(~WSA9, scales = "free") +
  theme_bw()

# density plot with RUG for one draw
all_predictions %>% 
  filter(.draw == 1) %>% 
  ggplot(aes(x=f.n2o.Mg.y)) + 
  geom_density() +
  geom_rug() + 
  facet_wrap(~WSA9, scales = "free")

# point and line range by size and ecoregion
all_predictions %>%
  group_by(WSA9, size_cat, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_e.n2o.mg.m2.d = mean(e.n2o.mg.m2.d)) %>% # 500 means for each WSA9 x size cat
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_e.n2o.mg.m2.d), 3), 
    LCL = round(quantile(mean_e.n2o.mg.m2.d, probs = 0.025), 3),
    UCL = round(quantile(mean_e.n2o.mg.m2.d, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  #select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = size_cat, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 0, color='blue') +
  ylab("mean N2O emission rate") +
  coord_flip() + 
  theme_bw() +
  facet_wrap(~ecoregion)
ggsave("output/figures/n2oEmissionRateBySizeEco.tiff", width = 7)

# FLUX-------------
# flux vs area for one draw
all_predictions %>% 
  filter(.draw == 1) %>% 
  ggplot(aes(y=f.n2o.Mg.y, x=area_ha)) + 
  geom_point() + 
  facet_wrap(~WSA9, scales = "free")

# flux by ecoregion
# point and linerange
all_predictions %>%
  group_by(WSA9, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_f = sum(f.n2o.Mg.y)) %>% # 500 means for each WSA9
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_f), 3), 
    LCL = round(quantile(mean_f, probs = 0.025), 3),
    UCL = round(quantile(mean_f, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = ecoregion, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 0, color='blue') +
  ylab("N2O flux (Mg N2O)") +
  coord_flip() + 
  theme_bw()
ggsave("output/figures/n2oFuxByEco.tiff")

# point and line range by size and ecoregion: facetted
all_predictions %>%
  group_by(WSA9, size_cat, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_f.n2o.Mg.y = sum(f.n2o.Mg.y)) %>% # 500 means for each WSA9 x size cat
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_f.n2o.Mg.y), 3), 
    LCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.025), 3),
    UCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.975), 3)) %>% 
  mutate(ecoregion = factor(WSA9)) %>%
  #mutate(type = "MR") %>%
  #select(ecoregion, estimate, LCL, UCL) %>%
  mutate(ecoregion = fct_reorder(ecoregion, estimate)) %>%
  ggplot( aes( x = size_cat, y = estimate ) ) +
  geom_point() +
  geom_linerange( aes( ymin = LCL, ymax = UCL)) +
  geom_hline(yintercept = 0, color='blue') +
  ylab("mean N2O flux (Mg N2O year-1)") +
  coord_flip() + 
  theme_bw() +
  facet_wrap(~ecoregion, scales = "free")
ggsave("output/figures/n2oFluxBySizeEco.tiff", height = 5.4, width = 9.6)

# point and line range by size and ecoregion: single plot
all_predictions %>%
  group_by(WSA9, size_cat, .draw) %>% # group by iteration
  #group_by(.draw) %>%
  summarise(mean_f.n2o.Mg.y = sum(f.n2o.Mg.y)) %>% # 500 means for each WSA9 x size cat
  # now summarize to 1 statistic per WSA9
  summarise( estimate = round(median(mean_f.n2o.Mg.y), 3), 
    LCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.025), 3),
    UCL = round(quantile(mean_f.n2o.Mg.y, probs = 0.975), 3)) %>% 
  mutate(unique = paste(WSA9, size_cat, "_")) %>%
  # mutate(ecoregion = fct_reorder(ecoregion, )) %>%
  ggplot( aes( y = size_cat, x = estimate ) ) +
  geom_point() +
  geom_linerange( aes( xmin = LCL, xmax = UCL)) +
  geom_vline(xintercept = 0, color='blue') +
  xlab(expression(N[2]*O~flux~(Mg~year^{-1}))) +
  #coord_flip() + 
  theme_bw() +
  facet_grid(WSA9~., scales = "free")
ggsave("output/figures/n2oFluxBySizeEcoVerticalPanel.tiff", height = 5.7, width = 2.5)

# bar plot by size category and ecoregion  
all_predictions %>% 
  filter(.draw == 1) %>% 
  group_by(WSA9, size_cat) %>% 
  summarise(total.n2o = sum(f.n2o.Mg.y)) %>% 
  ggplot(aes(x=size_cat, y=total.n2o)) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~WSA9, scales = "free")

# bar plot by size category with error
p <- all_predictions %>% 
  group_by(WSA9, .draw, size_cat2) %>% 
  summarise(total.n2o = sum(f.n2o.Mg.y)) %>% 
  group_by(WSA9, size_cat2) %>%
  summarise(post_med_fn2o = round(mean(total.n2o), 3), 
            LCL_fn2o = round(quantile(total.n2o, probs = 0.025), 3),
            UCL_fn2o = round(quantile(total.n2o, probs = 0.975), 3)) %>%
  ggplot(aes(x=size_cat2, y=post_med_fn2o)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(x=size_cat2, ymin = LCL_fn2o, ymax = UCL_fn2o)) +
    xlab(expression(Lake~Size~Bin~(km^{2}))) +
  ylab(expression(Mg~N[2]*O~year^{-1})) +
  facet_wrap(~WSA9, scales = "free") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90))

# this bit fills strip with same colors used in micromap (see below)
#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
g <- ggplot_gtable(ggplot_build(p))
stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
k <- 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid::grid.draw(g)

ggsave("output/figures/n2oFluxByLakeSizeWithError.tiff", g, width = 9.6, height = 5.4)

# bar plot by size category without error
p <- all_predictions %>% 
  group_by(WSA9, .draw, size_cat2) %>% 
  # tot x size x draw
  summarise(total.n2o = sum(f.n2o.Mg.y),
            obs = n()) %>% # number of lakes per WSA x size
  group_by(WSA9, size_cat2) %>%
  summarise(post_med_fn2o = round(mean(total.n2o), 3), # across draws
            LCL_fn2o = round(quantile(total.n2o, probs = 0.025), 3),
            UCL_fn2o = round(quantile(total.n2o, probs = 0.975), 3),
            obs = mean(obs)) %>% # obs is repeated across draws.  mean gives identity
  ggplot(aes(x=size_cat2, y=post_med_fn2o)) + 
  geom_bar(stat = "identity") + 
  xlab(expression(Lake~Size~Bin~(km^{2}))) +
  ylab(expression(Mg~N[2]*O~year^{-1})) +
  #scale_x_discrete(labels = c("min to 4", "4-10", "10-20", "20-50", ">50")) +
  #geom_errorbar(aes(x=size_cat, ymin = LCL_fn2o, ymax = UCL_fn2o)) +
  facet_wrap(~WSA9, scales = "free_y") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  geom_text(aes(label = scales::comma(obs, accuracy = 1)), # comma for thousands separator, accurary for decimal points
            vjust = -0.5, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.2)))

# this bit fills strip with same colors used in micromap (see below)
#https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set
g <- ggplot_gtable(ggplot_build(p))
stripr <- which(grepl('strip-t', g$layout$name)) # need strip-t, not strip-r
fills <- rep(brewer.pal(5, "Spectral")[c(1,3,5)], 3) # these are the 3 colors used in micromap
k <- 1
for (i in stripr) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}
grid::grid.draw(g)

ggsave("output/figures/n2oFluxByLakeSizeNoError.tiff", g)

# N2O flux vs total surface area in each ecoregion
all_predictions %>%
  group_by(WSA9, WSA9_NAME, .draw) %>%
  summarise(fn2o_Mg = sum(f.n2o.Mg.y),
             total_sa = sum(area_ha)) %>%
  summarise( post_med_fn2o = round(median(fn2o_Mg), 3), 
             LCL_fn2o = round(quantile(fn2o_Mg, probs = 0.025), 3),
             UCL_fn2o = round(quantile(fn2o_Mg, probs = 0.975), 3),
             total_sa = mean(total_sa)) %>%
  ggplot(aes(total_sa, post_med_fn2o)) +
  geom_point()

```

## MICROMAP: N2O saturation ratio continuous, extent, and emission rate

### Set up data
```{r}
# summarize means by WSA9 (posterior median for mean and 95% credible intervals)
# this is analogous to spsurvey "population estimates"
ecoreg_stats <- all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise( mean_n2o = mean(n2o),
             mean_n2oeq = mean(n2oeq),
             mean_sat = mean(n2osat),
             prop_sat = sum(n2osat < 1) / length(unique(.row)),
             mean_en2o = mean(e.n2o.mg.m2.d),
             fn2o_Mg = sum(f.n2o.Mg.y)) %>%
  summarise( post_med_n2o = round(mean(mean_n2o), 1),
             LCI_n2o = round(quantile(mean_n2o, probs = 0.025), 1),
             UCI_n2o = round(quantile(mean_n2o, probs = 0.975), 1),
             post_med_n2oeq = round(mean(mean_n2oeq), 2),
             LCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.025), 2),
             UCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.975), 2),
             post_med_sat = round(mean(mean_sat), 2),
             LCI_sat = round(quantile(mean_sat, probs = 0.025), 2),
             UCI_sat = round(quantile(mean_sat, probs = 0.975), 2),
             estimate = round(mean(prop_sat), 3),
             LCL = round(quantile(prop_sat, probs = 0.025), 3),
             UCL = round(quantile(prop_sat, probs = 0.975), 3),
             post_med_en2o = round(mean(mean_en2o), 3),
             LCL_en2o = round(quantile(mean_en2o, probs = 0.025), 3),
             UCL_en2o = round(quantile(mean_en2o, probs = 0.975), 3),
             post_med_fn2o = round(mean(fn2o_Mg), 3), 
             LCL_fn2o = round(quantile(fn2o_Mg, probs = 0.025), 3),
             UCL_fn2o = round(quantile(fn2o_Mg, probs = 0.975), 3),
             n = dplyr::n())

ecoreg_stats <- as.data.frame(ecoreg_stats)
class(ecoreg_stats)
str(ecoreg_stats) # WSA9 is a factor

ecoreg_stats$WSA9 <- as.character(ecoreg_stats$WSA9)
str(ecoreg_stats)

ecoreg_stats

ecoregions <- st_read(paste0(localPath, "\\Environmental Protection Agency (EPA)\\",
                             "ORD NLA17 Dissolved Gas - Documents\\inputData\\aggr_ecoregions_simple.shp"))

wsa9names <- st_set_geometry(ecoregions, NULL) %>%
  select(WSA9, WSA9_NAME)

ecoreg_stats <- left_join(ecoreg_stats, wsa9names)

ecoregions <- st_transform(ecoregions, 5070)
# make spatial polygon dataframe need for micromap
ecoregions_sp1 <- as(ecoregions, "Spatial")

# convert to map_table for linked micromap
ecor.polys <- create_map_table(ecoregions_sp1, 'WSA9')
class(ecor.polys)
str(ecor.polys, max =2) # ID is chr "CPL"...

```

### N2O SATURATION AND UNDERSATURATION EXTENT
```{r}
# Full code:  Use this to produce final linked micromaps of the variables you want
# export as pdf US letter landscape from Plots pane
# export as PNG Width 900 Height 968, maintain aspect ratio

mmplot(stat.data=ecoreg_stats,
       map.data=ecor.polys,
       panel.types=c('dot_legend', 'labels', 'dot_cl', 'dot_cl', 'map'),
       panel.data=list(
         NA,
         'WSA9_NAME',
         list('estimate', 'LCL', 'UCL'),
         list('post_med_sat', 'LCI_sat', 'UCI_sat'),
         NA),
       ord.by = 'estimate', rev.ord = TRUE, grouping=3,
       median.row=FALSE,
       map.link=c('WSA9', 'ID'),

       plot.height=5, # what does this do?
       plot.width=7, #what does this do? 
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
       map.color2='lightgray',

       panel.att=list(list(1, point.type=20, point.size=2,
                           graph.border.color='white',
                           xaxis.text.display=F, xaxis.line.display=F,
                           graph.grid.major=F, 
                           panel.width = 2), # added this argument to ensure the full point is shown

                      list(2,  panel.width=2,#increased panel.width to 1.1 from .8, removed header='NARS Reporting\n Regions',
                           align='left', text.size=0.8),#increased text.size to .8 from .6

                      list(3, header= "Proportion undersaturated\n with 95% CI", 
                           graph.bgcolor='lightgray',
                           xaxis.ticks=list(.65, 0.7, 0.75),
                           xaxis.labels=list(.65, 0.7, 0.75),
                           xaxis.title='Proportion undersaturated', panel.width=2.3),

                      list(4, header="N2O Saturation\n with 95% CI", #N2O Saturation\n with 95% CI
                           graph.bgcolor='lightgray',
                           xaxis.ticks=list(0.8, 1.0, 1.2, 1.4, 1.6),
                           xaxis.labels=list(0.8, 1.0, 1.2, 1.4, 1.6),
                           xaxis.title='N2O Saturation, %',panel.width=2.3),

                      list(5, header='Light Gray Means\n Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                           panel.width=2)),

       print.file = "output/figures/mmapSatExtent.tiff", print.res = 300)
```

### N2O EMISSION AND FLUX
```{r}
# Full code:  Use this to produce final linked micromaps of the variables you want
# export as pdf US letter landscape from Plots pane
# export as PNG Width 900 Height 968, maintain aspect ratio

mmplot(stat.data=ecoreg_stats,
       map.data=ecor.polys,
       panel.types=c('dot_legend', 'labels', 'dot_cl', 'dot_cl', 'map'),
       panel.data=list(
         NA,
         'WSA9_NAME',
         list('post_med_en2o', 'LCL_en2o', 'UCL_en2o'),
         list('post_med_fn2o', 'LCL_fn2o', 'UCL_fn2o'),
         NA),
       ord.by = 'post_med_en2o', rev.ord = TRUE, grouping=3,
       median.row=FALSE,
       map.link=c('WSA9', 'ID'),

       plot.height=5, # what does this do?
       plot.width=7, #what does this do? 
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
       map.color2='lightgray',

       panel.att=list(list(1, point.type=20, point.size=2,
                           graph.border.color='white',
                           xaxis.text.display=F, xaxis.line.display=F,
                           graph.grid.major=F, 
                           panel.width = 2), # added this argument to ensure the full point is shown

                      list(2,  panel.width=2,
                           align='left', text.size=0.8),

                      list(3, header= "Emission rate\n with 95% CI", 
                           graph.bgcolor='lightgray',
                           #xaxis.ticks=list(6.0, 7.0, 8.0, 9.0, 10.0),
                           #xaxis.labels=list(6.0, 7.0, 8.0, 9.0, 10.0),
                           xaxis.title='Emission rate (mg m-2 d-1)', panel.width=2.3),

                      list(4, header="N2O flux\n with 95% CI", #N2O Saturation\n with 95% CI
                           graph.bgcolor='lightgray',
                           #xaxis.ticks=list(0.8, 1.0, 1.2, 1.4, 1.6),
                           #xaxis.labels=list(0.8, 1.0, 1.2, 1.4, 1.6),
                           xaxis.title="N2O flux (Mg y-1)", panel.width=2.3),

                      list(5, header='Light Gray Means\n Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                           panel.width=2)),

       print.file = "output/figures/mmapEmisFlux.tiff", print.res = 300)
```

# Manuscript

Results and Discussion
In this study we present the first national-scale survey of N2O concentrations and emissions from lakes,  reservoirs, and ponds, hereafter referred to as waterbodies, in the conterminous U.S. (CONUS).  A generalized random-tessellation survey design (Stevens and Olsen, 2004) was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` [weird, I have 1090, but NLA17 report says 1,005?] sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week (Figure 1).   Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate (See NLA website for all data?). `r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 1)` percent of the sampled waterbodies were undersaturated in N2O and were therefore functioning as N2O sinks.  We modeled the data using a bayesion hierchical approach and predicted the dissolved N2O concentration for all  224,916 CONUS waterbodies in the target population.  At the population level, `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% pull()` percent (95% CI: `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% pull()` - `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% pull()`) of CONUS waterbodies were functioning as N2O sink.  The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N2O sinks (mean = `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% filter(WSA9 == "WMT") %>% select(estimate) %>% pull()`%, 95% CI: `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% filter(WSA9 == "WMT") %>% pull()` - `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% filter(WSA9 == "WMT") %>% pull()`) and the Northern Plains had the greatest (mean = `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% filter(WSA9 == "NPL") %>% select(estimate) %>% pull()`%, 95% CI: `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% filter(WSA9 == "NPL") %>% pull()` - `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% filter(WSA9 == "NPL") %>% pull()`)).

This survey design allows for extrapolation of the measurement results to all 224,916 CONUS waterbodies (this is from the 2017 NLA web report) in the target population.