### 2024-04-05 scripts/DG_sensitivity_to_measurement_error.Rmd
adding uncertainty in denominator

---
title: "Evaluating sensitivity to measurement error in dissolved gas measurements"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    df_print: paged
  pdf_document:
    toc: true
  html_notebook:
    depth: 4
    fig_caption: true
    number_sections: true
    toc: true
    toc_float: true
link-citations: true
---

<style>
.vscroll-plot {
    width: 800px;
    height: 600px;
    overflow-y: scroll;
    overflow-x: hidden;
}
</style>

# Background
In this document we explore a range of models for fitting dissolved $\text{N}_2\text{O}$ data observed as part of the 2017 National Lakes Assessment (NLA).

# Setup R
```{r setup_1, include=FALSE, echo=FALSE}
library(knitr)
# opts_knit$set(root.dir = "C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
# 
# #set directories and R package library
# setwd("C:/Users/rmartin/OneDrive - Environmental Protection Agency (EPA)/Documents/AE_Reservoirs/DissolvedGasNla/scripts")
# .libPaths("C:/Users/rmartin/Desktop/R/library")
```

```{r load_packages, warning=FALSE, message=FALSE}
library(brms)
library(bayesplot)
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(readxl)
library(janitor)
library(sf)

options(mc.cores = parallel::detectCores())
options( max.print = 2000 )
```

```{r utility_functions, include=FALSE}
# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}

#Calc geometric mean
gm_mean = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(mean(log(x), na.rm = na.rm))
  } else {
    exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

gm_sd = function(x, na.rm=TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(sd(log(x), na.rm = na.rm))
  } else {
    exp(sd(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
}

#calculate skew
MySkew <- function(x) {
     n <- length(x)
     v <- var(x)
     m <- mean(x)
     third.moment <- (1/(n - 2)) * sum((x - m)^3)
     third.moment/(var(x)^(3/2))
 }

#Center and scale by (1) sd
MyStd <- function(x){ (x-mean(x))/(sd(x)*2)}

#First letter in string to uppercase
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x}
```

# Dissolved gas concentration calculation

The full equation for estimating the concentration (nM) of $N_20$ gas dissolved in water is:

$$C = 1^{-6}B \frac{V_{air}(Mr_h - Mr_r)}{(8.3144598 * K * V_{water}) + H^{\theta} \exp( 2700 * ( \frac{1}{K} - \frac{1}{298.15}) * Mr_h}$$

# Set up model for simulating true value

```{r setvars_and_check}
load(paste0(localPath, "\\Environmental Protection Agency (EPA)",
            "\\ORD NLA17 Dissolved Gas - Documents\\",
            "inputData\\dg.2021-02-01.RData"))

# summary of measured N2O ppm in equilibrated headspace gas
dg %>%
  group_by(sample.source) %>%
  summarize(mean_n2O = mean(n2o.ppm.gc, na.rm = TRUE),
          median_n2o = median(n2o.ppm.gc, na.rm = TRUE)) #mean = 0.36, median = 0.31

#    mean_n2o  median_n2o
#AIR	0.3093312	0.3104941		
#DG	 0.3618681	0.3067609

B <- 99 # barometric pressure (kPa)
V_g <- 35 # volume of reference gas in headspace
V_w <- 105 # volume of water below headspace
T_c <- 23 # temperature ( C )
C_e <- 0.307 # gas mixture ratio in headspace
C_a <- 0.310 # gas mixture ratio in reference
H <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant

C <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )

C <- C * 1e9 #convert mol to nmol

round( C, 1 )

```
## Define GC measurement error
Measurement error will be defined as standard deviation of repeated analysis of analytical standards throughout the period of time that the NLA2017 samples were analyzed.

```{r get_gc_error}
# READ GC DATA----------------
##########################################################################
#       Code is commented out because collaborators may not have access  #
#       to the file paths. The calculated standard deviation is hard     #
#       coded below.
##########################################################################

# Below we read raw GC data into the environment, extract repeat analysis of
# an analytical standard, then calculate the standard deviation of the replicate
# measurements.

# gc.1 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_06_29_STD_ECD_FID_UNK.xlsx", 
#                    trim_ws = TRUE, skip=59) 
# gc.2 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_07_25_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=59)
# gc.3 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_02_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=50)
# gc.4 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_04_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=50)
# gc.5 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_16_Dgas_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=62)
# gc.6 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_21_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=62)
# gc.7 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_31_UNK_STD_ECD_FID2.xlsx", 
#                    trim_ws = TRUE, skip=65)
# gc.8 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_09_14_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=62)
# gc.9 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_09_19_STD_UNK_ECD_FID_TCD.xlsx", 
#                    trim_ws = TRUE, skip=102)
# gc.10 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_09_28_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=50)
# gc.11 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_10_11_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.12 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_10_17_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.13 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_12_20_STD_UNK_ECD_FID.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.14 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_03_AIR_UNK_STD_ECD_FID.xlsx", 
#                     trim_ws = TRUE, skip=48)
# gc.15 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_09_FID_ECD_STD.xlsx", 
#                     trim_ws = TRUE, skip=48)
# gc.16 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_16_UNK_STD_ECD_FID.xlsx", 
#                     trim_ws = TRUE, skip=48)
# gc.17 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_22_UNK_STD_ECD_FID.xlsx", 
#                     trim_ws = TRUE, skip=60)
# gc.18 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_31_STD_UNK_ECD_FID_TCD.xlsx", 
#                     trim_ws = TRUE, skip=102)
# gc.19 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_05_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.20 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_09_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.21 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_12_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.22 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_16_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.23 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_20_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.24 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_04_25_UNK_STD_ECD_FID.xlsx", 
#                     trim_ws = TRUE, skip=72)
# gc.25 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_23_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.26 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_27_UNK_STD_ECD_FID_DP.xlsx", 
#                     trim_ws = TRUE, skip=72)
# gc.27 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_03_05_Trap_6ml.xlsx", 
#                     trim_ws = TRUE, skip=104)
# # Merge gas data
# gc.all <- Reduce(function(...) merge(..., all=T),
#                   list(gc.1, gc.2, gc.3, gc.4, gc.5, gc.6, gc.7, 
#                        gc.8, gc.9, gc.10, gc.11, gc.12, gc.13, gc.14, gc.15, 
#                        gc.16, gc.17, gc.18, gc.19, gc.20, gc.21, gc.22, gc.23, 
#                        gc.24, gc.25, gc.26, gc.27)) %>%
#   janitor::clean_names()
# 
# 
# # Format gas data
# gc.std <- gc.all %>% 
#   select( sample, contains("n2O")) %>%
#   filter(grepl("std", sample, ignore.case = TRUE)) %>% # only keep standards
#   mutate(sample = tolower(sample), # simplify formatting
#          # we have multiple injections of this 1.8 Prax Air standard
#          sample = case_when(grepl("1.8 prax", sample, ignore.case = TRUE) ~ "1.8prax",
#                             TRUE ~ sample))
# 
# # with some invenstigating we could pull out multiple standards used
# # across and within runs to calculate standard deviation, but the 27
# # replicate injections of the 1.8 prax standard is low hanging fruit
# gc.prax <- gc.std %>%
#   filter(sample == "1.8prax")
# 
# gc.prax %>%
#   ggplot(aes(x = n2o_ppm)) +
#   geom_histogram()
# 
# n2o_sd <- sd(gc.prax$n2o_ppm) # 0.007872511ppm
n2o_sd <- 0.007872511

# # begin digging for other standards to work with
# gc.std %>%
#   filter(grepl("3p", sample, ignore.case = TRUE)) %>%
#   ggplot(aes(sample,n2o_ppm)) +
#   geom_point()
# 
# # list items 17 and 18 contain "m3p chk std"
# # see if these are same standards
# map(list(gc.1, gc.2, gc.3, gc.4, gc.5, gc.6, gc.7, 
#          gc.8, gc.9, gc.10, gc.11, gc.12, gc.13, gc.14, gc.15, 
#          gc.16, gc.17, gc.18, gc.19, gc.20, gc.21, gc.22, gc.23, 
#          gc.24, gc.25, gc.26, gc.27),
#     function(x) {
#       x %>% janitor::clean_names(.) %>% 
#         filter(grepl("m3p chk std", sample, ignore.case = TRUE))
#     })
# 
# table(gc.std$sample)

```

## Simulate measurement error in GC measurement of N2O in equilibrated headspace
```{r simulate_me_Ce, message=F}
C_me <- rep( NA, 1e4 ) # "me" is "measured" gas in headspace. empty vector

# this simulates measurement error on GC.
# assume TRUE value is C_e and error is simulated with with sd of replicate GC measurements
C_e_me <- rnorm( 1e4, C_e, n2o_sd ) # me in measured gas in headspace.
                                   # add random error to measured concentration

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

# combine into dataframe for ggplot. qplot, which acceptes vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, C_e_me = C_e_me)

ggplot(C_df, aes(x = C_e_me )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_e , color = 'red' ) + 
  ggtitle( "Simulated measurement error in equilibrated gas concentration") +
  xlab( "concentration (nmol L-1)")

ggplot(C_df, aes(x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof equilibrated headspace") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes(x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof equilibrated headspace") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error for GC measurement of N2O in air
```{r simulate_me_Ca, message=F}
C_me <- rep( NA, 1e4 )

C_a_me <- rnorm( 1e4, C_a, n2o_sd ) #m.e. in measured gas in air

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
Ca_df <- tibble(C_me = C_me, C_a_me = C_a_me)

ggplot(Ca_df, aes(x = C_a_me )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_a , color = 'red' ) + 
  ggtitle( "Simulated measurement error in air gas concentration") +
  xlab( "concentration (?)")

ggplot(Ca_df, aes(x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof N2O in air") +
  xlab( "concentration (nMol)")

ggplot(Ca_df, aes(x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N2O concentration due to GC measurment error \nof N2O in air") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in $C_e$ and $C_a$
```{r simulate_me_Ca_Ce, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e_me[ i ] )
}

# combine into dataframe for ggplot. qplot, which acceptes vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me)

ggplot(C_df, aes(x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof 
           N2O in equilibrated headspace and air") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes(x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to GC measurment error \nof N2O in equilibrated headspace and air") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in water temperature
```{r simulate_me_T_c, message=F}
C_me <- rep( NA, 1e4 )

# How to get standard deviation of temperature measurement?
# calibration certificate reports an uncertainty of +/-0.058C
# take standard deviation of simulated measurements with specified uncertainty
T_c_sd <- sd(runif(n = 1e4, min = T_c - 0.058, max = T_c + 0.058)) # calibration certificate reports an uncertainty of +/-0.058C
T_c_me <- rnorm( 1e4, T_c, T_c_sd ) #m.e. in measured water temp

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which acceptes vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, T_c_me = T_c_me)

ggplot(C_df, aes(x = T_c_me )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = T_c , color = 'red' ) + 
  ggtitle( "Simulated measurement error in temperature") +
  xlab( "temperature (C)")

ggplot(C_df, aes( x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to water temperature \nmeasurment error") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to water temperature \nmeasurment error") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in barometric pressure
```{r simulate_me_B, message=F}
C_me <- rep( NA, 1e4 )

# How to get standard deviation of barometric pressure?
# The manual reports an "accuracy" of +/-3 mm Hg (+/-0.4kPa)
# take standard deviation of simulated measurements with specified uncertainty
B_sd <- sd(runif(n = 1e4, min = B - 0.4, max = B + 0.4)) # calibration certificate reports an uncertainty of +/-0.058C
B_me <- rnorm( 1e4, B, 0.25 ) #m.e. in measured barometric pressure

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, B_me = B_me, B_ru = runif(n = 1e4, min = B - 0.4, max = B + 0.4))

ggplot(C_df, aes( x = B_me )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = B , color = 'red' ) + 
  ggtitle( "Simulated measurement error in barometric pressure") +
  xlab( "barometric pressure (kPa)")

ggplot(C_df, aes( x = B_ru )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = B , color = 'red' ) + 
  ggtitle( "Simulated measurement error in barometric pressure") +
  xlab( "barometric pressure (kPa)")

ggplot(C_df, aes( x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to barometric pressure \nmeasurment error") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to barometric pressure \nmeasurment error") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in water volume
```{r simulate_me_V_w, message=F}
C_me <- rep( NA, 1e4 )

V_w_me <- rnorm( 1e4, V_w, 1 ) #m.e. in measured water volume
V_g_me <- 140 - V_w_me

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B * ( V_g_me[ i ] * ( C_e - C_a ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me, V_w_me = V_w_me)

ggplot(C_df, aes( x = V_w_me )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = V_w , color = 'red' ) + 
  ggtitle( "Simulated measurement error in water volume") +
  xlab( "water vol (mL)")

ggplot(C_df, aes( x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissollved N2O concentration due to water volume \nmeasurment error") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in N2O concentration due to measurment error") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error for all observables
```{r simulate_me_all, message=F}
C_me <- rep( NA, 1e4 )

for( i in 1:length( C_me ) ){
  C_me[ i ] <- 1e-6 * B_me[ i ] * ( V_g_me[ i ] * ( C_e_me[ i ] - C_a_me[ i ] ) / ( 8.3144598 * ( T_c_me[ i ] + 273.15 ) * V_w_me[ i ] ) + H * exp( 2700 * ( 1 / ( T_c_me[ i ] + 273.15 ) - 1 / 298.15 ) ) * C_e )
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_df <- tibble(C_me = C_me)

ggplot(C_df, aes( x = C_me * 1e9 )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors") +
  xlab( "concentration (nMol)")

ggplot(C_df, aes( x = C_me * 1e9 - C )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in dissolved N2O concentration due to cumulative \nmeasurment errors") +
  xlab( "absolute error (nMol)")

# standard deviation of a measurement is:
sd(C_me* 1e9) # 0.15 nmol L-1, as compared to 1.25 nmol L-1 in Webb et al.

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(C_me* 1e9) * 1.96 # +/- 0.302 nmol N2O
```

## Incorporate measurement uncertainty into source/sink status
The standard deviation of dissolved gas concentration measurements is `r sd(C_me* 1e9)`nM which compares well to 1.25nM reported by Webb et al.  The 95% confidence interval is calculated as +/-2sd, which in this case is `r sd(C_me* 1e9) * 1.96`nM.  Therefore any dissolved gas concentration that is within `r sd(C_me* 1e9) * 1.96`nM of the equilibrium value cannot be determined as a source or a sink.

```{r map_table, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

dg <- dg %>%
  mutate(n2o.src.snk.with.error = case_when(
    abs(dissolved.n2o.nmol - sat.n2o.nmol) > sd(C_me* 1e9) * 1.96 ~ n2o.src.snk,
    TRUE ~ "undetermined"))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

## load ecoregions----
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple",
                quiet = TRUE)

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# SET UP CUSTOM COLORS FOR ECOREGIONS
# Custom color pallette for ecoregion polygons. Attempted to mirror
# https://www.epa.gov/national-aquatic-resource-surveys/
# ecoregional-results-national-lakes-assessment-2012
cols <- c("Coastal Plains" = "orange1",
          "Northern Appalachians" = "lightpink1",
          "Northern Plains" = "darksalmon",
          "Southern Appalachians" = "mediumturquoise",
          "Southern Plains" = "khaki4",
          "Temperate Plains" = "forestgreen", 
          "Upper Midwest" = "deepskyblue4",
          "Western Mountains" = "saddlebrown",
          "Xeric" = "lightskyblue4")

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

# To enable spatial analysis of the data, the dataframe will be converted to a 
# 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME)) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.error),
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("grey", "black", "orange"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.1, 10), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```
# Set up model for simulating true saturation concentration

The full equation for estimating the concentration (nM) of $N_20$ gas dissolved in water that is in equilibrium with the atmosphere is:

$$C = 1^{-6}B*C_{air}*exp(2700* \frac{1}{Tc} -  \frac{1}{298.15})$$
```{r c_sat}
C_sat <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))

C_sat <- C_sat * 1e9 #convert mol to nmol

```

## Simulate measurement error in GC measurement of N2O in air
```{r simulate_me_sat_C_a, message=F}
C_me_sat <- rep( NA, 1e4 )

C_a_me <- rnorm( 1e4, C_a, n2o_sd ) #m.e. in measured gas in air

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a_me[i] * B * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat, C_a_me = C_a_me)

ggplot(C_sat_df, aes(x = C_a_me )) +  
  geom_histogram( binwidth = 0.0001 ) + 
  geom_vline( xintercept = C_a , color = 'red' ) + 
  ggtitle( "Simulated measurement error in air gas concentration") +
  xlab( "concentration (ppm)")

ggplot(C_sat_df, aes(x = C_me_sat * 1e9 )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C_sat , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to GC measurment error \nof N2O in air") +
  xlab( "concentration (nMol)")

ggplot(C_sat_df, aes(x = C_me_sat * 1e9 - C_sat )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to GC measurment error \nof N2O in air") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in water temperature
```{r simulate_me_sat_T_c, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a * B * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which acceptes vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat, T_c_me = T_c_me)

ggplot(C_sat_df, aes(x = T_c_me )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = T_c , color = 'red' ) + 
  ggtitle( "Simulated measurement error in temperature") +
  xlab( "temperature (C)")

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline(xintercept = C_sat , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to water temperature \nmeasurment error") +
  xlab( "concentration (nMol)")

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to water temperature \nmeasurment error" ) +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error in barometric pressure
```{r simulate_me_sat_B, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat, B_me = B_me)

ggplot(C_sat_df, aes( x = B_me )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = B , color = 'red' ) + 
  ggtitle( "Simulated measurement error in barometric pressure") +
  xlab( "barometric pressure (kPa)")

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = C_sat , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to barometric pressure \nmeasurment error") +
  xlab( "concentration (nMol)")

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to barometric pressure \nmeasurment error") +
  xlab( "absolute error (nMol)")
```

## Simulate measurement error for all observables
```{r simulate_me_sat_all, message=F}
C_me_sat <- rep( NA, 1e4 )

for( i in 1:length( C_me_sat ) ){
    C_me_sat[i] <- C_a_me[ i ] * B_me[ i ] * 1e-6 * (H  * exp(2700*(1/(T_c_me[ i ] + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_sat_df <- tibble(C_me_sat = C_me_sat)

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 )) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_vline( xintercept = C_sat , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to cumulative \nmeasurment errors") +
  xlab( "concentration (nMol)")

ggplot(C_sat_df, aes( x = C_me_sat * 1e9 - C_sat )) +  
  geom_histogram( binwidth = 0.001 ) + 
  geom_vline( xintercept = 0 , color = 'red' ) + 
  ggtitle( "Uncertainty in saturated N2O concentration due to cumulative \nmeasurment errors") +
  xlab( "absolute error (nMol)")

# standard deviation of a measurement is:
sd(C_me_sat* 1e9) # 0.20 nmol L-1

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(C_me_sat * 1e9) * 1.96 # +/- 0.395 nmol N2O
```