### 2024-05-24 scripts/DG_sensitivity_to_measurement_error.Rmd
Updated formulas and figures; Added scenarios

---
title: "Sensitivity to measurement error in dissolved gas measurements"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 5
    number_sections: true
    keep_html: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, warning=FALSE, message=FALSE}
library(knitr)
library(rmarkdown)
library(bayesplot)
library(kableExtra)
library(ggpubr)
library(tidyverse)
library(tidybayes)
library(ggplot2)
library(ggExtra)
library(gridExtra)
library(readxl)
library(janitor)
library(sf)

# Identify local path for each user
localPath <- Sys.getenv("USERPROFILE")

# Define helper functions
# standardized formatting for column names
toEPA <- function(X1){
  names(X1) = tolower(names(X1))
  names(X1) = gsub(pattern = c("\\(| |#|)|/|-|\\+|:|_"), replacement = ".", x = names(X1))
  X1
}

options(max.print = 2000)
```

# Background

In this document we explore a range of scenarios wherein measurement errors 
could arise in methods for determining the concentration of $\text{N}_2\text{O}$ 
from field and laboratory measurements observed as part of the 2017 National 
Lakes Assessment (NLA).

# Dissolved concentration

The equation for estimating the concentration (nM) of $\text{N}_2\text{O}$ gas
dissolved in water is:

$$C = B^{-6} \cdot \left(\frac{V_g(M_h - M_r)}{(8.3144598 \cdot K \cdot V_w) + (H^{\theta} \cdot e^{2700 \cdot ( \frac{1}{K} - \frac{1}{298.15})} \cdot M_h)} \right)$$

where:

-   $C$ = concentration (nM)
-   $B$ = barometric pressure (kPa)
-   $V_g$ = volume of reference gas in headspace (mL)
-   $M_h$ = gas mixture ratio in the headspace (C_e below)
-   $M_r$ = gas mixture ratio in the reference (C_a_n2o below)
-   $V_w$ = volume of water below the headspace (mL)
-   $K$ = temperature (K) (temp celsius or T_c + 273 below)
-   $H^{\theta}$ = Henry's law constant ($H_{N_2O}$ or $H_{Ar}$ below)

and:

$$\begin{split}
 V & = V_g + V_w \\
  V_g & = V \cdot R_M \\
  V_w & = V \cdot (1 - R_M)
  \end{split}$$

where:

-   $V$ = Total volume of the mixing vessel (e.g. 140mL syringe)
-   $V_g$ = Volume of gas (mL) in the mixing vessel 
-   $V_w$ = Volume of water (mL) in the mixing vessel
-   $R_M$ = Mixing ratio (gas:water)

## Simulate $\text{N}_2\text{O}$

```{r summary_empirical, echo=TRUE}
load(paste0(localPath, "\\Environmental Protection Agency (EPA)",
            "\\ORD NLA17 Dissolved Gas - Documents\\",
            "inputData\\dg.2021-02-01.RData"))

# summary of measured N2O ppm in equilibrated headspace gas
dg %>%
  group_by(sample.source) %>%
  summarize(mean_n2O = mean(n2o.ppm.gc, na.rm = TRUE),
          median_n2o = median(n2o.ppm.gc, na.rm = TRUE)) #mean = 0.36, median = 0.31
```

Simulate a single concentration without measurment error to represent the "true" 
or "known" value for dissolved $\text{N}_2\text{O}$.
```{r dg_calc, echo=TRUE}
# Set volume of syringe/vessel
V <- 140 # 140mL syringe

# Set ratio of gas volume to water volume in the mixing vessel
R_M <- 0.25

B <- 99 # barometric pressure (kPa)
V_g <- V * R_M # volume of reference gas in headspace
V_w <- V * (1 - R_M) # volume of water below headspace
T_c <- 23 # temperature ( C )
M_h <- 0.307 # gas mixture ratio in headspace (Mr_h) * Changes as function of R_M at constant C
M_a_n2o <- 0.310 # gas mixture ratio in reference (Mr_r)
H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
H_ar <- 0.000014 #mol m-3 Pa, range: 0.000013 - 0.000016; Henry's law constant

C <- 1e-6 * B * ( V_g * ( M_h - M_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * M_h )

C <- C * 1e9 #convert mol to nmol

round( C, 1 ) 
```

```{r dg_calc_vary_r, echo=TRUE}
# Set volume of syringe/vessel
V <- 140 # 140mL syringe

# Set ratio of gas volume to water volume in the mixing vessel
R_M <- 0.25

B <- 99 # barometric pressure (kPa)
V_g <- V * R_M # volume of reference gas in headspace
V_w <- V * (1 - R_M) # volume of water below headspace
T_c <- 23 # temperature ( C )
M_h <- 0.307 # gas mixture ratio in headspace (Mr_h)
M_a_n2o <- 0.310 # gas mixture ratio in reference (Mr_r)
H_n2o <- 0.00024 #mol m-3 Pa, range: 0.00018 - 0.00025; Henry's law constant
H_ar <- 0.000014 #mol m-3 Pa, range: 0.000013 - 0.000016; Henry's law constant

C <- 1e-6 * B * ( V_g * ( M_h - M_a_n2o ) / ( 8.3144598 * ( T_c + 273.15 ) * V_w ) + H_n2o * exp( 2700 * ( 1 / ( T_c + 273.15 ) - 1 / 298.15 ) ) * M_h )

C <- C * 1e9 #convert mol to nmol

round( C, 1 ) 
```

## Empricial variation
Estimate variation due to measurement uncertainty using empirical data from
laboratory replicates.
```{r get_gc_error, echo=FALSE}
# READ GC DATA----------------
##########################################################################
#       Code is commented out because collaborators may not have access  #
#       to the file paths. The calculated standard deviation is hard     #
#       coded below.
##########################################################################

# Below we read raw GC data into the environment, extract repeat analysis of
# an analytical standard, then calculate the standard deviation of the replicate
# measurements.

# gc.1 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_06_29_STD_ECD_FID_UNK.xlsx", 
#                    trim_ws = TRUE, skip=59) 
# gc.2 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_07_25_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=59)
# gc.3 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_02_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=50)
# gc.4 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_04_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=50)
# gc.5 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_16_Dgas_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=62)
# gc.6 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_21_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=62)
# gc.7 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_08_31_UNK_STD_ECD_FID2.xlsx", 
#                    trim_ws = TRUE, skip=65)
# gc.8 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_09_14_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=62)
# gc.9 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_09_19_STD_UNK_ECD_FID_TCD.xlsx", 
#                    trim_ws = TRUE, skip=102)
# gc.10 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_09_28_UNK_STD_ECD_FID.xlsx", 
#                    trim_ws = TRUE, skip=50)
# gc.11 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_10_11_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.12 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_10_17_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.13 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_12_20_STD_UNK_ECD_FID.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.14 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_03_AIR_UNK_STD_ECD_FID.xlsx", 
#                     trim_ws = TRUE, skip=48)
# gc.15 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_09_FID_ECD_STD.xlsx", 
#                     trim_ws = TRUE, skip=48)
# gc.16 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_16_UNK_STD_ECD_FID.xlsx", 
#                     trim_ws = TRUE, skip=48)
# gc.17 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_22_UNK_STD_ECD_FID.xlsx", 
#                     trim_ws = TRUE, skip=60)
# gc.18 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_01_31_STD_UNK_ECD_FID_TCD.xlsx", 
#                     trim_ws = TRUE, skip=102)
# gc.19 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_05_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.20 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_09_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.21 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_12_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.22 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_16_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.23 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_20_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.24 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/17_04_25_UNK_STD_ECD_FID.xlsx", 
#                     trim_ws = TRUE, skip=72)
# gc.25 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_23_UNK_STD_ECD_FID_Dgas.xlsx", 
#                     trim_ws = TRUE, skip=67)
# gc.26 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_02_27_UNK_STD_ECD_FID_DP.xlsx", 
#                     trim_ws = TRUE, skip=72)
# gc.27 <- read_excel("L:/Lab/Lablan/GHG/GC/2017Data/18_03_05_Trap_6ml.xlsx", 
#                     trim_ws = TRUE, skip=104)
# # Merge gas data
# gc.all <- Reduce(function(...) merge(..., all=T),
#                   list(gc.1, gc.2, gc.3, gc.4, gc.5, gc.6, gc.7, 
#                        gc.8, gc.9, gc.10, gc.11, gc.12, gc.13, gc.14, gc.15, 
#                        gc.16, gc.17, gc.18, gc.19, gc.20, gc.21, gc.22, gc.23, 
#                        gc.24, gc.25, gc.26, gc.27)) %>%
#   janitor::clean_names()
# 
# 
# # Format gas data
# gc.std <- gc.all %>% 
#   select( sample, contains("n2O")) %>%
#   filter(grepl("std", sample, ignore.case = TRUE)) %>% # only keep standards
#   mutate(sample = tolower(sample), # simplify formatting
#          # we have multiple injections of this 1.8 Prax Air standard
#          sample = case_when(grepl("1.8 prax", sample, ignore.case = TRUE) ~ "1.8prax",
#                             TRUE ~ sample))
# 
# # with some invenstigating we could pull out multiple standards used
# # across and within runs to calculate standard deviation, but the 27
# # replicate injections of the 1.8 prax standard is low hanging fruit
# gc.prax <- gc.std %>%
#   filter(sample == "1.8prax")
# 
# gc.prax %>%
#   ggplot(aes(x = n2o_ppm)) +
#   geom_histogram()
# 
# n2o_sd <- sd(gc.prax$n2o_ppm) # 0.007872511ppm
n2o_sd_epa <- 0.007872511 # from AWBERC GC

# # begin digging for other standards to work with
# gc.std %>%
#   filter(grepl("3p", sample, ignore.case = TRUE)) %>%
#   ggplot(aes(sample,n2o_ppm)) +
#   geom_point()
# 
# # list items 17 and 18 contain "m3p chk std"
# # see if these are same standards
# map(list(gc.1, gc.2, gc.3, gc.4, gc.5, gc.6, gc.7, 
#          gc.8, gc.9, gc.10, gc.11, gc.12, gc.13, gc.14, gc.15, 
#          gc.16, gc.17, gc.18, gc.19, gc.20, gc.21, gc.22, gc.23, 
#          gc.24, gc.25, gc.26, gc.27),
#     function(x) {
#       x %>% janitor::clean_names(.) %>% 
#         filter(grepl("m3p chk std", sample, ignore.case = TRUE))
#     })
# 
# table(gc.std$sample)

```

Measurement error is defined here as the standard deviation of repeated
analysis of analytical standards throughout the period of time that the
NLA2017 samples were analyzed, which was $\sigma$ = `r n2o_sd_epa`.

## Simulate measurment error

Below, potential errors associated with dissolved $\text{N}_2\text{O}$ 
measurement are simulated and assessed separately for GC observations in:

-

    (1) the equilibrated headspace,

-

    (2) the air, and

-

    (3) the headspace + air

Measurement error is then simulated and assessed separately for:

-

    (4) water temperature,

-

    (5) barometric pressure, and

-

    (5) water volume

Finally, all of the sources (1, 2, 4, 5, and 6) are simulated
simultaneously and assessed as the total variation in dissolved
$\text{N}_2\text{O}$. This total error is then added to the point estimates for 
$\text{N}_2\text{O}$ observed in this study, which results in an estimate of
uncertainty due to measurement error. These error from these simulations was also
carried through to estimated of the saturation ratio. The source-sink status for
each site included in this study was then mapped to present a source-sink 
categorization that reflected these estimated uncertainties. The categorization 
was made such that any observation with simulated error that was within the 
central 95th percentile of

Set the number of simulations for error:

```{r set_n_sims}
nsim <- 1e5
```

### Equilibrated headspace
Simulate measurment error in dissolved $\text{N}_2\text{O}$ due solely to variation in 
measurements within the equlibrated headspace.
```{r simulate_C_h_me, echo=TRUE, message=FALSE}
# Create empty vector to fill below with simulated values for N2O observed with 
# error due to variation in the GC measurments of gas in the equilibrated 
# headspace. The subscript, "me", denotes an observed or measured value, which 
# is contrasted with the known value, C, set above.
C_h_me <- rep(NA, nsim)

# The true value for the concentration of equilibrated gas in the headspace, 
# C_e, was set above. This step simulates additive error in this measurement 
# that is centered over the true value. The observed concentration of
# equilibrated gas is simulated as normally distributed variable, with a mean
# C_e, and a standard deviation specified as the observed sd of replicate GC 
# measurements in our lab. 
M_h_me <- rnorm(nsim, M_h, n2o_sd_epa)

for(i in 1:nsim){
  C_h_me[i] <- 1e-6 * B * (V_g * (M_h_me[i] - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h_me[i])
}

# combine into dataframe for ggplot. 
C_h_df <- tibble(C_h_me = C_h_me, M_h_me = M_h_me)
# save absolute error for plot
C_h_me_abs <- (C_h_me * 1e9) - C
```

The results of the simulations are summarized with density histograms
below.

```{r plot_C_h_me, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
ggarrange(
ggplot(C_h_df, aes(x = M_h_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = M_h , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in \nequilibrated headspace (Mh) \nGC measurements") +
  xlab("Concentration (nmol L-1)"),

ggplot(C_h_df, aes(x = C_h_me * 1e9, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
  xlab("Concentration (nMol)"),

ggplot(C_h_df, aes(x = C_h_me_abs, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = 0 , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  xlab("Error (nMol)"),
nrow = 1
)
```

### Ambient air
Error in the GC measurements of $\text{N}_2\text{O}$ in the ambient air ($M_a$) is simulated 
below. The result of the simulation is plotted as a histogram to illustrate the
variation in the measurement. THe resulting effect on the uncertainty of the 
dissolved N2O measurement is also plotted along with the absolute error in 
N2O as a result of error in this measurement.
```{r simulate_me_Ca, message=F}
C_a_me <- rep(NA, nsim)

M_a_n2o_me <- rnorm(nsim, M_a_n2o, n2o_sd_epa) # m.e. in measured gas in air

for(i in 1:nsim){
  C_a_me[i] <- 1e-6 * B * (V_g * (M_h - M_a_n2o_me[i]) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h)
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_a_df <- tibble(C_a_me = C_a_me, M_a_n2o_me = M_a_n2o_me)
# save absolute error for plot
C_a_me_abs<- (C_a_me * 1e9) - C
```

```{r plot_me_Ca, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
ggarrange(
ggplot(C_a_df, aes(x = M_a_n2o_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.001 ) +
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = M_a_n2o , color = 'red' ) +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Simulated variation in \nreference gas (Ma) GC \nmeasurements") +
  xlab( "Concentration (nmol L-1)"),

ggplot(C_a_df, aes(x = C_a_me * 1e9, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.01) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
  xlab("Concentration (nMol)"),

ggplot(C_a_df, aes(x = C_a_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.01 ) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle( "Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1)
```

### Water temperature

```{r simulate_me_T_c, message=F}
C_tc_me <- rep(NA, nsim)

# How to get standard deviation of temperature measurement?
# calibration certificate reports an uncertainty of +/-0.058C
# take standard deviation of simulated measurements with specified uncertainty
T_c_sd <- sd(runif(n = nsim, min = T_c - 0.058, max = T_c + 0.058)) # calibration certificate reports an uncertainty of +/-0.058C
T_c_me <- rnorm(nsim, T_c, T_c_sd) #m.e. in measured water temp

for(i in 1:nsim){
  C_tc_me[i] <- 1e-6 * B * (V_g * (M_h - M_a_n2o) / (8.3144598 * (T_c_me[i] + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15)) * M_h)
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_tc_df <- tibble(C_tc_me = C_tc_me, T_c_me = T_c_me)
# save absolute error for plot
C_tc_me_abs <- (C_tc_me * 1e9) - C
```

```{r plot_me_T_c, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
ggarrange(
ggplot(C_tc_df, aes(x = T_c_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = T_c , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in water \ntemperature measurements \n") +
  xlab("Temperature (C)"),

ggplot(C_tc_df, aes(x = C_tc_me * 1e9, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
  xlab("Concentration (nMol)"),

ggplot(C_tc_df, aes(x = C_tc_me_abs, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = 0 , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  xlab("Error (nMol)"),
nrow = 1)
```

### Barometric pressure

```{r simulate_me_B, message=F}
C_b_me <- rep(NA, nsim)

# How to get standard deviation of barometric pressure?
# The manual reports an "accuracy" of +/-3 mm Hg (+/-0.4kPa), but EPA calibration
# checks performed in the AWBERC metrology lab demonstrate much greater precision.
# below we calculat sd of replicate measurements made with 15 different barometers.
# we then take the mean of the 15 sd values.

#2023 barometer verification in AWBERC metrology lab
B_sd <- mean(
  sd(c(999,998.6)*0.1), # YSI ProDSS
  sd(c(1000.848, 1000.848)*0.1), # EXO 21G104022
  sd(c(1000.982, 1000.982)*0.1), # EXO 18H111572
  sd(c(1000.982, 1000.982)*0.1), # EXO 22F105975
  sd(c(1001.382, 1001.382)*0.1), # EXO 21K100162
  sd(c(1000.98, 1000.85)*0.1), # EXO 21F103579
  sd(c(1000.31, 1000.44)*0.1), # MDS 16D100327
  sd(c(1001.25, 1001.25)*0.1), # MDS 12F102618
  sd(c(1000, 1000)*0.1), # MDS 16D100328
  sd(c(1002, 1002)*0.1), # MDS 02K599 AF
  sd(c(1003,1003)*0.1), # MDS 12J0793 AF
  sd(c(1000.85,1000.85)*0.1), # MDS 16E100831
  sd(c(1000.32, 1000.32)*0.1), # MDS 12F102619
  sd(c(1002, 1002)*0.1)) # MDS 15F104725

B_me <- rnorm(nsim, B, B_sd) #m.e. in measured barometric pressure

for(i in 1:nsim){
  C_b_me[i] <- 1e-6 * B_me[i] * (V_g * (M_h - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h)
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_b_df <- tibble(C_b_me = C_b_me, B_me = B_me)
# save absolute error for plot
C_b_me_abs <- (C_b_me * 1e9) - C
```

```{r plot_me_B, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
ggarrange(
ggplot(C_b_df, aes(x = B_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = B , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in \nbarometric pressure \nmeasurements") +
  xlab("Barometric pressure (kPa)"),

ggplot(C_b_df, aes(x = C_b_me * 1e9, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.0001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
  xlab("Concentration (nMol)"),

ggplot(C_b_df, aes(x = C_b_me_abs, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.0001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = 0 , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  xlab("Error (nMol)"),
nrow = 1)
```

### Water volume

```{r simulate_me_V_w, message=F}
C_w_me <- rep(NA, nsim)

V_w_me <- rnorm(nsim, V_w, 1) #m.e. in measured water volume
V_g_me <- 140 - V_w_me

for(i in 1:nsim){
  C_w_me[i] <- 1e-6 * B * (V_g_me[i] * (M_h - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w_me[i] ) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h)
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_w_df <- tibble(C_w_me = C_w_me, V_w_me = V_w_me)
# save absolute error for plot
C_w_me_abs <- (C_w_me * 1e9) - C
```

```{r plot_me_V_w, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
ggarrange(
ggplot(C_w_df, aes(x = V_w_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.1) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = V_w , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in \nwater volume \nmeasurements") +
  xlab("Water volume (mL)"),

ggplot(C_w_df, aes(x = C_w_me * 1e9, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.00025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle("Resulting uncertainty in \ndissolved N2O estimate \n") +
  xlab("Concentration (nMol)"),

ggplot(C_w_df, aes(x = C_w_me_abs, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.00025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = 0 , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  xlab("Error (nMol)"),
nrow = 1)
```

### All observables
Simulate measurement error for all observables (_i.e._, Equilibrated headspace, 
air, water temperature, and barometric pressure ) and the resulting effects on
dissolved N2O observations
```{r simulate_me_all, message=F}
C_all_me <- rep(NA, nsim)

for(i in 1:nsim){
  C_all_me[i] <- 1e-6 * B_me[i] * (V_g_me[i] * (M_h_me[i] - M_a_n2o_me[i] ) / (8.3144598 * (T_c_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15)) * M_h_me[i])
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_all_df <- tibble(C_all_me = C_all_me)
# save absolute error for plot
C_all_me_abs <- (C_all_me * 1e9) - C
```

```{r plot_me_all, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(C_all_df, aes(x = C_all_me * 1e9, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.05 ) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in dissolved N2O \nestimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(C_all_df, aes(x = C_all_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.05) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

Print the estimated standard deviations of dissolved N2O measurements
and the resulting 95% credible interval for error in measurements.

```{r calc_sd_me_all}
# standard deviation of a measurement is:
sd(C_all_me* 1e9) # 0.20 nmol L-1

# 2 standard deviations
sd(C_all_me * 1e9) * 1.96 # +/- 0.395 nmol N2O
```

```{r calc_sd_me_eq_all_2}
# standard deviation of a measurement is:
round(sd(C_all_me_abs), 3) # 0.20 nmol L-1

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((C + C_all_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

## Summary of results

```{r final_figure_N2O_diss, message=FALSE, fig.align='center', fig.asp=0.6, fig.width=6}
results_dissolved_N2O <- tibble(Eh_me= C_h_me_abs,
                                Air_me = C_a_me_abs,
                                Tc_me = C_tc_me_abs,
                                B_me = C_b_me_abs,
                                W_me = C_w_me_abs,
                                All_me = C_all_me_abs) %>%
  pivot_longer(cols = ends_with("me"), names_to = "Source", values_to = "Error")

results_dissolved_N2O %>%
  mutate(Source = factor(Source)) %>%
  mutate(Source = fct_relevel(Source, c("B_me",
                                        "W_me",
                                        "Tc_me",
                                        "Air_me",
                                        "Eh_me",
                                        "All_me"
                                        ))) %>%
  ggplot(aes(x = Error, y = Source)) +
  stat_pointinterval() +
  scale_y_discrete(labels=c("Eh_me" = "Equilibrated headspace",
                            "Air_me" = "Ambient air",
                            "Tc_me" = "Water temperature",
                            "B_me" = "Barometric pressure",
                            "W_me" = "Water volume",
                            "All_me" = "All sources")) +
  xlab("Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  ylab("") 
```

# Equilibrium concentration

The full equation for estimating the concentration (nM) of ($\text{N}_2\text{O}$) gas
dissolved in water that is in equilibrium with the atmosphere is:

$$C_{eq} = 1^{-6} \cdot B \cdot C_a \cdot H^{\theta} \cdot e^{2700 \cdot \frac{1}{Tc} -  \frac{1}{298.15}}$$
where:

-   $C_{eq}$ = The equilibrium concentration in nM
-   $B$ = barometric pressure (kPa)
-   $C_a$ = The air concentration in nM
-   $H^{\theta}$ = Henry's law constant ($H_{N_2O}$ below)
-   $T_c$ = temperature celsius

## Simulate equilibrium $\text{N}_2\text{O}$

A "known" concentration is simulated:

```{r c_sat}
C_eq <- M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))

C_eq <- C_eq * 1e9 #convert mol to nmol
```

## Simulate measurement error

### Ambient air

Error in the GC measurements of $\text{N}_2\text{O}$ in the ambient air ($M_a$) is simulated 
below. The resulting effects on the uncertainty of the equilibrium N2O 
measurement is plotted along with the absolute error in N2O as a result of this
measurement.

```{r simulate_me_sat_C_a_n2o, message=F}
C_a_eq_me <- rep( NA, nsim )

M_a_n2o_me <- rnorm( nsim, M_a_n2o, n2o_sd_epa ) #m.e. in measured gas in air

for(i in 1:nsim){
    C_a_eq_me[i] <- M_a_n2o_me[i] * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15)))
}

C_a_eq_df <- tibble(C_a_eq_me = C_a_eq_me, M_a_n2o_me = M_a_n2o_me)
# save absolute error for plot
C_a_eq_me_abs <- (C_a_eq_me * 1e9) - C_eq
```

```{r plot_me_sat_C_a_n2o, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
ggarrange(
ggplot(C_a_eq_df, aes(x = M_a_n2o_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = M_a_n2o , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in air \ngas (M_a_n2o) GC \nmeasurements") +
  xlab("Concentration (nmol L-1)"),

ggplot(C_a_eq_df, aes(x = C_a_eq_me * 1e9, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C_eq , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Resulting uncertainty in \nequilibrium N2O estimate \n") +
  xlab("Concentration (nMol)"),

ggplot(C_a_eq_df, aes(x = C_a_eq_me_abs, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = 0 , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
  xlab("Error (nMol)"),
nrow = 1
)
```

### Water temperature

```{r simulate_me_sat_T_c, message=F}
C_tc_eq_me <- rep(NA, nsim)

for(i in 1:nsim){
    C_tc_eq_me[i] <- M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1/298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_tc_eq_df <- tibble(C_tc_eq_me = C_tc_eq_me, T_c_me = T_c_me)
# save absolute error for plot
C_tc_eq_me_abs <- (C_tc_eq_me * 1e9) - C_eq
```

```{r plot_me_sat_T_c, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
ggarrange(
ggplot(C_tc_eq_df, aes(x = T_c_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = T_c , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in water \ntemperature measurements") +
  xlab("Concentration (nmol L-1)"),

ggplot(C_tc_eq_df, aes(x = C_tc_eq_me * 1e9, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C_eq , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Resulting uncertainty in \nequilibrium N2O estimate \n") +
  xlab("Concentration (nMol)"),

ggplot(C_tc_eq_df, aes(x = C_tc_eq_me_abs, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = 0 , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
  xlab("Error (nMol)"),
nrow = 1
)
```

### Barometric pressure

```{r simulate_me_sat_B, message=F}
C_b_eq_me <- rep(NA, nsim)

for(i in 1:nsim){
    C_b_eq_me[i] <- M_a_n2o * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_b_eq_df <- tibble(C_b_eq_me = C_b_eq_me, B_me = B_me)
# save absolute error for plot
C_b_eq_me_abs <- (C_b_eq_me * 1e9) - C_eq
```

```{r plot_me_sat_B, echo=FALSE, fig.align='center', fig.asp=0.35, fig.width=8}
ggarrange(
ggplot(C_b_eq_df, aes(x = B_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = B , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in \nbarometic pressure \nmeasurements") +
  xlab("Barometric pressure (kPa)"),

ggplot(C_b_eq_df, aes(x = C_b_eq_me * 1e9, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.0005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C_eq , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Resulting uncertainty in \nequilibrium N2O estimate \n") +
  xlab("Concentration (nMol)"),

ggplot(C_b_eq_df, aes(x = C_b_eq_me_abs, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.00025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = 0 , color = 'red') +  
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) +
  ggtitle("Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
  xlab("Error (nMol)"),
nrow = 1
)
```

### All observables
Simulate measurement error for all observables (_i.e._, air, water temperature,
and barometric pressure ) and the resulting effects on dissolved N2O 
observations.
```{r simulate_me_sat_all, message=F}
C_all_eq_me <- rep( NA, nsim )

for(i in 1:nsim){
    C_all_eq_me[i] <- M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15)))
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
C_all_eq_df <- tibble(C_all_eq_me = C_all_eq_me)
# save absolute error for plot
C_all_eq_me_abs <- (C_all_eq_me * 1e9) - C_eq
```

```{r plot_me_sat_all, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(C_all_eq_df, aes(x = C_all_eq_me * 1e9, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.025 ) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C_eq , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in equilibrium N2O \nestimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(C_all_eq_df, aes(x = C_all_eq_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_sd_me_sat_all}
# standard deviation of a measurement is:
sd(C_all_eq_me* 1e9) # 0.20 nmol L-1

# 2 standard deviations
sd(C_all_eq_me * 1e9) * 1.96 # +/- 0.395 nmol N2O
```

```{r calc_sd_me_sat_all_2}
# standard deviation of a measurement is:
round(sd(C_all_eq_me_abs), 3) # 0.20 nmol L-1

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((C_eq + C_all_eq_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

## Summary of results

```{r final_figure_N2O_eq, message=FALSE, fig.align='center', fig.asp=0.6, fig.width=6}
results_equilibrium_N2O <- tibble(Air_me = C_a_eq_me_abs,
                                Tc_me = C_tc_eq_me_abs,
                                B_me = C_b_eq_me_abs,
                                All_me = C_all_eq_me_abs) %>%
  pivot_longer(cols = ends_with("me"), names_to = "Source", values_to = "Error")

results_equilibrium_N2O %>%
  mutate(Source = factor(Source)) %>%
  mutate(Source = fct_relevel(Source, c("B_me",
                                        "Tc_me",
                                        "Air_me",
                                        "All_me"
                                        ))) %>%
  ggplot(aes(x = Error, y = Source)) +
  stat_pointinterval() +
  scale_y_discrete(labels=c("Air_me" = "Ambient air",
                            "Tc_me" = "Water temperature",
                            "B_me" = "Barometric pressure",
                            "All_me" = "All sources")) +
  xlab("Absolute error in equilibrium \nN2O estimate (observed - true) \n") +
  ylab("") 
```

# Saturation ratio
(Check this equation in text! Particularly the description of variables)
The full equation for estimating the saturation ratio of ($N_20$) gas
dissolved in water is:

$$S = \left(\frac{(1^{-6} \cdot B ) \biggl( \frac{V_a (Mr_h - Mr_r)}{(8.3144598 \cdot K \cdot V_w) + H^{\theta} \cdot e^{2700 \cdot ( \frac{1}{K} - \frac{1}{298.15})} \cdot Mr_h} \biggr)}{1^{-6} \cdot B \cdot C_a \cdot e^{2700 \cdot \frac{1}{Tc} -  \frac{1}{298.15}}} \right)$$

where:

-   $S$ = Saturatio ratio
-   $B$ = barometric pressure (kPa)
-   $C_a$ = the air concentration in nM
-   $V_a$ = volume of reference gas in air
-   $M_h$ = gas mixture ratio in the headspace
-   $M_r$ = gas mixture ratio in the reference (C_a_n2o below)
-   $V_w$ = volume of water below the headspace
-   $T_c$ = temperature celsius
-   $H^{\theta}$ = Henry's law constant ($H_{N_2O}$ below)

Simulate a "true" saturation ratio.
```{r sat_ratio_true}
S <- (1e-6 * B * (V_g * (M_h - M_a_n2o) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)) * M_h)) /
  (M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)))) 
S
```

## Standard GC
Simulate uncertainty in the saturation ratio due to measurement error associated
with the standard GC, using ambient air as the headspace gas, and the standard 
field thermometer.

### Ambient air, standard thermometer

```{r simulate_me_ratio_all_air, message=F}
S_a_me <- rep( NA, nsim )

for(i in 1:nsim){
S_a_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_me[i] - M_a_n2o_me[i]) / (8.3144598 * (T_c_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15)) * M_h_me[i])) /
  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
S_a_df <- tibble(S_a_me = S_a_me)
# save absolute error for plot
S_a_me_abs <- S_a_me - S
```

```{r plot_me_ratio_all_air, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(S_a_df, aes(x = S_a_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.01) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = S , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(S_a_df, aes(x = S_a_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.01) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_sd_me_ratio_all_air}
# standard deviation of a saturation ratio value s:
sd(S_a_me) # 0.026

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(S_a_me) * 1.96 # +/- 0.05258214

# so any ratio within 0.05 of 1.0 cannot be discerned as a source or sink?
```

```{r calc_sd_me_ratio_all_air_2}
# standard deviation of a measurement is:
round(sd(S_a_me_abs), 3) # 0.20 nmol L-1

# central 95th percentile of absolute error
round(quantile(S_a_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((S + S_a_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

#### Map source-sink status

```{r map_ratio_all_air, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

#dg <- dg %>%
#  mutate(n2o.src.snk.with.all.error = case_when(
#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
#    TRUE ~ n2o.src.snk)) 
dg <- dg %>%
  mutate(n2o.src.snk.with.all.error = case_when(
    abs(n2o.sat.ratio - 1) < quantile(S_a_me_abs, probs = c(0.975)) ~ "undetermined",
    TRUE ~ n2o.src.snk))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.all.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

# sf for plotting
dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting
```

```{r plot_map_ratio_all_air, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
## load ecoregions----
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple",
                quiet = TRUE)

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# SET UP CUSTOM COLORS FOR ECOREGIONS
# Custom color pallette for ecoregion polygons. Attempted to mirror
# https://www.epa.gov/national-aquatic-resource-surveys/
# ecoregional-results-national-lakes-assessment-2012
cols <- c("Coastal Plains" = "orange1",
          "Northern Appalachians" = "lightpink1",
          "Northern Plains" = "darksalmon",
          "Southern Appalachians" = "mediumturquoise",
          "Southern Plains" = "khaki4",
          "Temperate Plains" = "forestgreen", 
          "Upper Midwest" = "deepskyblue4",
          "Western Mountains" = "saddlebrown",
          "Xeric" = "lightskyblue4")

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

# To enable spatial analysis of the data, the dataframe will be converted to a 
# 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
            shape = 1,
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", 
                      values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.5, 8), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

### Pure gas, standard thermometer

Simulate uncertainty in the saturation ratio due to measurement error associated
with the standard GC, using a pure gas (helium) as the headspace gas, 
and the standard field thermometer.

```{r simulate_me_ratio_pg_all, message=F}
# set M_h to a lower value to generate a saturation ratio close to that used above
# M_h will be lower due to dilution with pure headspace gas (e.g. helium, nitrogen)
M_h_pg <- 0.2

# this simulates measurement error on GC.
# assume TRUE value is M_h and error is simulated with with sd of replicate GC measurements
M_h_pg_me <- rnorm(nsim, M_h_pg, n2o_sd_epa) # me in measured gas in headspace.
                                   # add random error to measured concentration

# under this scenario we don't need to subtract M_a_n2o (N2O concentration in air) from M_h (N2O concentration in equilibrated headspace)
S_pg <- (1e-6 * B * (V_g * (M_h_pg - 0) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15) ) * M_h_pg)) /
  (M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))

S_pg_me <- rep(NA, nsim)

for(i in 1:nsim){
S_pg_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_pg_me[i] - 0) / (8.3144598 * (T_c_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15)) * M_h_pg_me[i])) /
  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_me[i] + 273.15) - 1/298.15)))) 
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
S_pg_df <- tibble(S_pg_me = S_pg_me)
# save absolute error for plot
S_pg_me_abs <- S_pg_me - S_pg
```

```{r plot_me_ratio_pg_all, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(S_pg_df, aes(x = S_pg_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.01) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = S , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(S_pg_df, aes(x = S_pg_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.01) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_sd_me_ratio_pg_all}
# standard deviation of a saturation ratio value s:
sd(S_pg_me) # 0.048

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(S_pg_me) * 1.96 # +/- 0.094

# so any ratio within 0.094 of 1.0 cannot be discerned as a source or sink
```

```{r calc_sd_me_ratio_all_pg_2}
# standard deviation of a measurement is:
round(sd(S_pg_me_abs), 3) # 0.20 nmol L-1

# central 95th percentile of absolute error
round(quantile(S_pg_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((S_pg + S_pg_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

#### Map source-sink status

```{r map_ratio_pg, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

#dg <- dg %>%
#  mutate(n2o.src.snk.with.all.error = case_when(
#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
#    TRUE ~ n2o.src.snk)) 
dg <- dg %>%
  mutate(n2o.src.snk.with.all.error = case_when(
    abs(n2o.sat.ratio - 1) < quantile(S_pg_me_abs, probs = c(0.975)) ~ "undetermined",
    TRUE ~ n2o.src.snk))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.all.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

# sf for plotting
dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting
```

```{r plot_map_ratio_pg, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
## load ecoregions----
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple",
                quiet = TRUE)

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# SET UP CUSTOM COLORS FOR ECOREGIONS
# Custom color pallette for ecoregion polygons. Attempted to mirror
# https://www.epa.gov/national-aquatic-resource-surveys/
# ecoregional-results-national-lakes-assessment-2012
cols <- c("Coastal Plains" = "orange1",
          "Northern Appalachians" = "lightpink1",
          "Northern Plains" = "darksalmon",
          "Southern Appalachians" = "mediumturquoise",
          "Southern Plains" = "khaki4",
          "Temperate Plains" = "forestgreen", 
          "Upper Midwest" = "deepskyblue4",
          "Western Mountains" = "saddlebrown",
          "Xeric" = "lightskyblue4")

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

# To enable spatial analysis of the data, the dataframe will be converted to a 
# 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
            shape = 1,
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", 
                      values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.5, 8), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

### Ambient air, high precision thermometer
Simulate uncertainty in the saturation ratio due to measurement error associated
with the standard GC, ambient air as the headspace gas, and using a high 
precision thermometer (e.g., MIMS).

```{r simulate_tc_mims_air_hpt, message=F}
# High precision water batch used to control temperature of standards
T_c_mims_me <- rnorm(nsim, T_c, 0.005) # thermo stated bath precision is +/-0.01 Celsius
# hist(T_c_mims_me) # sd of 0.005 gives approx distribution for the specified precision

T_c_df <- tibble(T_c_me = T_c_me, 
                 T_c_me_abs = T_c_me - T_c,  
                 T_c_mims_me = T_c_mims_me,
                 T_c_mims_me_abs = T_c_mims_me - T_c)
```

The precision gain from the MIMS thermometer relative to the standard field 
thermometer is roughly one order of magnitude.
```{r plot_me_T_c_vs_T_c_MIMS, echo=FALSE, fig.align='center', fig.asp=0.9, fig.width=6}
ggarrange(
ggplot(T_c_df, aes(x = T_c_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = T_c , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in water \ntemperature measurements \nusing standard field thermometer") +
  xlab("Temperature (C)"),
ggplot(T_c_df, aes(x = T_c_mims_me, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = T_c , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Simulated variation in water \ntemperature measurements \n using MIMS thermometer") +
  xlab("Temperature (C)"),
ggplot(T_c_df, aes(x = T_c_me_abs, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = 0 , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Absolute error in water \ntemperature measurements \nusing standard field thermometer") +
  xlab("Error"),
ggplot(T_c_df, aes(x = T_c_mims_me_abs, y = after_stat(density))) +  
  geom_histogram(binwidth = 0.001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = 0 , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle("Absolute error in water \ntemperature measurements \nusing standard field thermometer") +
  xlab("Error"),
nrow = 2,
ncol = 2)
```

Simulate uncertainty in the saturation ratio using ambient air in the 
headspace and high precision temperature measurements
```{r simulate_me_ratio_all_air_hpt, message=F}
S_a_hpt_me <- rep(NA, nsim)

for(i in 1:nsim){
S_a_hpt_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_me[i] - M_a_n2o_me[i]) / (8.3144598 * (T_c_mims_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)) * M_h_me[i])) /
  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) 
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
S_a_hpt_df <- tibble(S_a_hpt_me = S_a_hpt_me)
# save absolute error for plot
S_a_hpt_me_abs <- S_a_hpt_me - S
```

```{r plot_me_ratio_all_air_hpt, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(S_a_hpt_df, aes(x = S_a_hpt_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.01) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = S , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(S_a_hpt_df, aes(x = S_a_hpt_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.01) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_sd_tc_mims_air_hpt}
# standard deviation of a saturation ratio value s:
sd(S_a_hpt_me) # 0.026

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(S_a_hpt_me) * 1.96 # +/- 0.05258214

# so any ratio within 0.05 of 1.0 cannot be discerned as a source or sink?
```

```{r calc_sd_tc_mims_air_hpt_2}
# standard deviation of a measurement is:
round(sd(S_a_hpt_me_abs), 3) # 0.20 nmol L-1

# central 95th percentile of absolute error
round(quantile(S_a_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((S + S_a_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

#### Map source-sink status

```{r map_mims_air_hpt, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

#dg <- dg %>%
#  mutate(n2o.src.snk.with.all.error = case_when(
#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
#    TRUE ~ n2o.src.snk)) 
dg <- dg %>%
  mutate(n2o.src.snk.with.all.error = case_when(
    abs(n2o.sat.ratio - 1) < quantile(S_a_hpt_me_abs, probs = c(0.975)) ~ "undetermined",
    TRUE ~ n2o.src.snk))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.all.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

# sf for plotting
dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting
```

```{r plot_map_mims_air_hpt, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
## load ecoregions----
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple",
                quiet = TRUE)

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# SET UP CUSTOM COLORS FOR ECOREGIONS
# Custom color pallette for ecoregion polygons. Attempted to mirror
# https://www.epa.gov/national-aquatic-resource-surveys/
# ecoregional-results-national-lakes-assessment-2012
cols <- c("Coastal Plains" = "orange1",
          "Northern Appalachians" = "lightpink1",
          "Northern Plains" = "darksalmon",
          "Southern Appalachians" = "mediumturquoise",
          "Southern Plains" = "khaki4",
          "Temperate Plains" = "forestgreen", 
          "Upper Midwest" = "deepskyblue4",
          "Western Mountains" = "saddlebrown",
          "Xeric" = "lightskyblue4")

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

# To enable spatial analysis of the data, the dataframe will be converted to a 
# 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
            shape = 1,
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", 
                      values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.5, 8), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

### Pure gas, high precision thermometer
Simulate uncertainty in the saturation ratio due to measurement error associated
with the standard GC, a pure gas (helium) as the headspace gas, and using a high 
precision thermometer (e.g., MIMS).

```{r simulate_me_ratio_all_pg_hpt, message=F}
S_pg_hpt_me <- rep(NA, nsim)

for(i in 1:nsim){
S_pg_hpt_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_pg_me[i] - 0) / (8.3144598 * (T_c_mims_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)) * M_h_pg_me[i])) /
  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) 
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
S_pg_hpt_df <- tibble(S_pg_hpt_me = S_pg_hpt_me)
# save absolute error for plot
S_pg_hpt_me_abs <- S_pg_hpt_me - S_pg
```

```{r plot_me_ratio_all_pg_hpt, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(S_pg_hpt_df, aes(x = S_pg_hpt_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = S_pg , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(S_pg_hpt_df, aes(x = S_pg_hpt_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_sd_me_ratio_all_pg_hpt}
# standard deviation of a saturation ratio value s:
sd(S_pg_hpt_me) # 0.026

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(S_pg_hpt_me) * 1.96 # +/- 0.05258214

# so any ratio within 0.05 of 1.0 cannot be discerned as a source or sink?
```

```{r calc_sd_me_ratio_all_pg_hpt_2}
# standard deviation of a measurement is:
round(sd(S_pg_hpt_me_abs), 3) # 0.20 nmol L-1

# central 95th percentile of absolute error
round(quantile(S_pg_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((S_pg + S_pg_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

#### Map source-sink status

```{r map_sat_ratio_pg_hpt, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

#dg <- dg %>%
#  mutate(n2o.src.snk.with.all.error = case_when(
#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
#    TRUE ~ n2o.src.snk)) 
dg <- dg %>%
  mutate(n2o.src.snk.with.all.error = case_when(
    abs(n2o.sat.ratio - 1) < quantile(S_pg_hpt_me_abs, probs = c(0.975)) ~ "undetermined",
    TRUE ~ n2o.src.snk))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.all.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

# sf for plotting
dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting
```

```{r plot_map_sat_ratio_pg_hpt, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
## load ecoregions----
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple",
                quiet = TRUE)

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# SET UP CUSTOM COLORS FOR ECOREGIONS
# Custom color pallette for ecoregion polygons. Attempted to mirror
# https://www.epa.gov/national-aquatic-resource-surveys/
# ecoregional-results-national-lakes-assessment-2012
cols <- c("Coastal Plains" = "orange1",
          "Northern Appalachians" = "lightpink1",
          "Northern Plains" = "darksalmon",
          "Southern Appalachians" = "mediumturquoise",
          "Southern Plains" = "khaki4",
          "Temperate Plains" = "forestgreen", 
          "Upper Midwest" = "deepskyblue4",
          "Western Mountains" = "saddlebrown",
          "Xeric" = "lightskyblue4")

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

# To enable spatial analysis of the data, the dataframe will be converted to a 
# 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
            shape = 1,
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", 
                      values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.5, 8), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

## High precision GC

The GC used in the current study is optimized to quantify several
analytes across a broad range of concentrations but greater precision
can be achieved by optimizing the instrument for a single analyte across
a narrower range of concentrations. For example, laboratories
participating in an international analysis of N2O standards used for
atmsopheric monitoring reported a mean standard deviation of
`r mean(c(.00033, .00034, .00031, 0.00030, 0.0016, 0.0015))`, which is
10-fold lower than the standard deviation of N2O measurements in this
study (`r n2o_sd_epa`). Utilizing more precise N2O measurements for the
equilibrated headspace and/or air will reduce uncertainty in the
estimated dissolved N2O concentration, dissolved N2O equilibrium value,
and N2O saturation ratio.

### Pure gas, standard thermometer
```{r simulate_me_ratio_gc_pure_gas_all, message=F}
# set M_h to a lower value to generate a saturation ratio close to that used above
# M_h will be lower due to dilution with pure headspace gas (e.g. helium, nitrogen)
M_h_pg_hpgc <- 0.2

# two standard deviations reported from each of three labs participating in
# a laboratory intercomparison study
n2o_sd_lit <- mean(c(.00033, .00034, .00031, 0.00030, 0.0016, 0.0015)) # from literature

# simulate improved measurement of N2O in equilibrated headspace gas 
# assume TRUE value is C_e and error is simulated with with sd from literature
M_h_pg_hpgc_me <- rnorm(nsim, M_h_pg_hpgc, n2o_sd_lit)

# under this scenario we don't need to subtract C_a_n2o (N2O concentration in air) from C_e (N2O concentration in equilibrated headspace)
S_pg_hpgc <- (1e-6 * B * (V_g * (M_h_pg_hpgc - 0) / (8.3144598 * (T_c + 273.15) * V_w) + H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15) ) * M_h_pg_hpgc)) /
  (M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700*(1/(T_c + 273.15) - 1/298.15))))

S_pg_hpgc_me <- rep(NA, 1e4)

for(i in 1:nsim){
S_pg_hpgc_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_pg_hpgc_me[i] - 0) / (8.3144598 * (T_c_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15) ) * M_h_pg_hpgc_me[i])) /
  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1 / (T_c_me[i] + 273.15) - 1/298.15)))) 
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
S_pg_hpgc_df <- tibble(S_pg_hpgc_me = S_pg_hpgc_me)
# save absolute error for plot
S_pg_hpgc_me_abs <- S_pg_hpgc_me - S_pg_hpgc
```

```{r plot_me_ratio_gc_pure_gas_all, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(S_pg_hpgc_df, aes(x = S_pg_hpgc_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = S , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(S_pg_hpgc_df, aes(x = S_pg_hpgc_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_sd_ratio_gc_pure_gas_all}
# standard deviation of a saturation ratio value s:
sd(S_pg_hpgc_me) # 0.029

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(S_pg_hpgc_me) * 1.96 # +/- 0.0569

# so any ratio within 0.0569 of 1.0 cannot be discerned as a source or sink
```

```{r calc_sd_me_ratio_gc_pure_gas_all_2}
# standard deviation of a measurement is:
round(sd(S_pg_hpgc_me_abs), 3) # 0.20 nmol L-1

# central 95th percentile of absolute error
round(quantile(S_pg_hpgc_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((S_pg_hpgc + S_pg_hpgc_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

#### Map source-sink status

```{r map_sat_ratio_gc_pure_gas, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

#dg <- dg %>%
#  mutate(n2o.src.snk.with.all.error = case_when(
#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
#    TRUE ~ n2o.src.snk)) 
dg <- dg %>%
  mutate(n2o.src.snk.with.all.error = case_when(
    abs(n2o.sat.ratio - 1) < quantile(S_pg_hpgc_me_abs, probs = c(0.975)) ~ "undetermined",
    TRUE ~ n2o.src.snk))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.all.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

# sf for plotting
dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting
```

```{r plot_map_sat_ratiogc_pure_gas, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
## load ecoregions----
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple",
                quiet = TRUE)

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# SET UP CUSTOM COLORS FOR ECOREGIONS
# Custom color pallette for ecoregion polygons. Attempted to mirror
# https://www.epa.gov/national-aquatic-resource-surveys/
# ecoregional-results-national-lakes-assessment-2012
cols <- c("Coastal Plains" = "orange1",
          "Northern Appalachians" = "lightpink1",
          "Northern Plains" = "darksalmon",
          "Southern Appalachians" = "mediumturquoise",
          "Southern Plains" = "khaki4",
          "Temperate Plains" = "forestgreen", 
          "Upper Midwest" = "deepskyblue4",
          "Western Mountains" = "saddlebrown",
          "Xeric" = "lightskyblue4")

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

# To enable spatial analysis of the data, the dataframe will be converted to a 
# 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
            shape = 1,
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", 
                      values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.5, 8), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

### Pure gas, high precision thermometer

```{r simulate_me_ratio_hpgc_pg_hpt_all, message=F}
# set M_h to a lower value to generate a saturation ratio close to that used above
# M_h will be lower due to dilution with pure headspace gas (e.g. helium, nitrogen)
M_h_pg_hpgc_hpt <- 0.2

# two standard deviations reported from each of three labs participating in
# a laboratory intercomparison study
n2o_sd_lit <- mean(c(.00033, .00034, .00031, 0.00030, 0.0016, 0.0015)) # from literature

# simulate improved measurement of N2O in equilibrated headspace gas 
# assume TRUE value is C_e and error is simulated with with sd from literature
M_h_pg_hpgc_hpt_me <- rnorm(nsim, M_h_pg_hpgc, n2o_sd_lit)

S_pg_hpgc_hpt_me <- rep(NA, 1e4)

for(i in 1:nsim){
S_pg_hpgc_hpt_me[i] <- (1e-6 * B_me[i] * (V_g_me[i] * (M_h_pg_hpgc_hpt_me[i] - 0) / (8.3144598 * (T_c_me[i] + 273.15) * V_w_me[i]) + H_n2o * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1 / 298.15) ) * M_h_pg_hpgc_hpt_me[i])) /
  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700*(1 / (T_c_me[i] + 273.15) - 1/298.15)))) 
}

# combine into dataframe for ggplot. qplot, which accepts vectors,
# was deprecated in ggplot 3.4.0
S_pg_hpgc_hpt_df <- tibble(S_pg_hpgc_hpt_me = S_pg_hpgc_hpt_me)
# save absolute error for plot
S_pg_hpgc_hpt_me_abs <- S_pg_hpgc_hpt_me - S_pg_hpgc
```

```{r plot_me_ratio_hpgc_pg_hpt_all, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(S_pg_hpgc_hpt_df, aes(x = S_pg_hpgc_hpt_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = S_pg, color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in saturation ratio \nestimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(S_pg_hpgc_hpt_df, aes(x = S_pg_hpgc_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.005) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_sd_ratio_hpgc_pg_hpt_all}
# standard deviation of a saturation ratio value s:
sd(S_pg_hpgc_hpt_me) # 0.029

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(S_pg_hpgc_hpt_me) * 1.96 # +/- 0.0569

# so any ratio within 0.0569 of 1.0 cannot be discerned as a source or sink
```

```{r calc_sd_me_ratio_hpgc_pg_hpt_all_2}
# standard deviation of a measurement is:
round(sd(S_pg_hpgc_hpt_me_abs), 3) # 0.20 nmol L-1

# central 95th percentile of absolute error
round(quantile(S_pg_hpgc_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((S_pg_hpgc + S_pg_hpgc_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

#### Map source-sink status

```{r map_ratio_hpgc_pg_hpt, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

#dg <- dg %>%
#  mutate(n2o.src.snk.with.all.error = case_when(
#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
#    TRUE ~ n2o.src.snk)) 
dg <- dg %>%
  mutate(n2o.src.snk.with.all.error = case_when(
    abs(n2o.sat.ratio - 1) < quantile(S_pg_hpgc_hpt_me_abs, probs = c(0.975)) ~ "undetermined",
    TRUE ~ n2o.src.snk))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.all.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

# sf for plotting
dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting
```

```{r plot_map_ratio_hpgc_pg_hpt, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
## load ecoregions----
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple",
                quiet = TRUE)

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# SET UP CUSTOM COLORS FOR ECOREGIONS
# Custom color pallette for ecoregion polygons. Attempted to mirror
# https://www.epa.gov/national-aquatic-resource-surveys/
# ecoregional-results-national-lakes-assessment-2012
cols <- c("Coastal Plains" = "orange1",
          "Northern Appalachians" = "lightpink1",
          "Northern Plains" = "darksalmon",
          "Southern Appalachians" = "mediumturquoise",
          "Southern Plains" = "khaki4",
          "Temperate Plains" = "forestgreen", 
          "Upper Midwest" = "deepskyblue4",
          "Western Mountains" = "saddlebrown",
          "Xeric" = "lightskyblue4")

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

# To enable spatial analysis of the data, the dataframe will be converted to a 
# 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
            shape = 1,
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", 
                      values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.5, 8), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

# MIMS

In Membrane Inlet Mass Spectrometry (MIMS), dissolved gases (typically
N2, Ar, O2, N2O) are extracted from a water sample as it passes through
a semi-permeable membrane. The abundance of the extracted gases are then
measured using a quadrapole mass spectrometer. One advantage of this
approach relative to the traditional headspace equilibrium method is
that dissolved gases are extracted without contamination and/or dilution
from a headspace gas (e.g. air, nitrogen, helium). Dilution by a pure
gas isn't much of a concern because N2O detection limits are very low
via GC, but correction for the presence of N2O in the headspace gas
reduces measurement precision.

Dissolved N2O concentration in a MIMS sample can be calculated from the
N2O signal strength and a calibration factor. The calibration factor is
determined by analyzing a water sample with known N2O concentration.
This is most frequently accomplished by equilibrating laboratory water
with air at constant temperature, barometric pressure and N2O partial
pressure. The dissolved N2O concentration of the equilibrated water is
calculated from solubility laws (see [Equilibrium concentration]) and the 
calibration factor, $f$, is:

$$f_{N_2O} = \frac{C^{N_2O}_s}{X^{N_2O}_s}$$
where:

-   $f_{N_2O} = \text{N}_2\text{O} \text{ calibration factor}$
-   $C^{N_2O}_s = \text{Concentration of } \text{N}_2\text{O} \text{ in calibration standard}$
-   $X^{N_2O}_s = \text{Signal strength of } \text{N}_2\text{O} \text{ in calibration standard}$

The N2O concentration of an unknown sample is then calculated as:

$$C^{N_2O}_{\theta} = \frac{X^{N_2O}}{f_{N_2O}}$$

where:

-   $C^{N_2O}_{\theta} = \text{N}_2\text{O} \text{ concentration in sample}$
-   $X^{N_2O}_{\theta} = \text{Signal strength of } \text{N}_2\text{O} \text{ in sample}$ 
-   $f_{N_2O} = \text{N}_2\text{O} \text{ calibration factor}$

The equation simply scales the concentration of the N2O standard to the
ratio of the MIMS signal strength of the unknown and standard. Via
substitution and rearrangement, the equation can also be written as:

$$C^{N_2O}_{\theta} =\frac{X^{N_2O}_{\theta}}{X^{N_2O}_s} \cdot C^{N_2O}_s$$

MIMS provides greater precision when measuring gas ratios as compared to
measuring a single gas. High precision gas ratio data can be converted
to concentration data with the independent measurement of concentration
of one of the component gases. N2O:Ar ratios are particularly useful for
environmental samples because Ar is biologically inert, therefore its
concentration can be calculated from solubility, water temperature,
barometric pressure, and atmospheric partial pressure (essentially
constant across the globe). As with single gas measurements, N2O:Ar
ratios are converted to N2O concentration using a calibration factor
derived from measurement of laboratory water equilibrated with air:

$$\begin{aligned}
f^\prime_{N_2O} =\frac{R^{C}_s}{R^{X}_s} \\
R^{C}_s = \frac{C^{N_2O}_s}{C^{Ar}_s} \\
R^{X}_s = \frac{X^{N_2O}_s}{X^{Ar}_s}
\end{aligned}$$

where:

-   $f^\prime_{N_2O} = \text{N}_2\text{O} \text{ (ratio) calibration factor}$
-   $R^{C}_s = \text{Concetration ratio of } \text{N}_2\text{O} \text{ to Ar in calibration standard}$
-   $R^{X}_s = \text{Signal ratio of } \text{N}_2\text{O} \text{ to Ar in calibration standard}$

The N2O concentration of an unknown sample is then calculated as:

$$C^{N_2O}_{\theta} = R^{X}_s \cdot f^\prime_{N_2O} \cdot C^{Ar}_s$$
where:

-   $C^{N_2O}_{\theta} = \text{Concentration of } \text{N}_2\text{O} \text{ in the sample}$
-   $C^{Ar}_s = \text{Concentration of } Ar \text{ in the calibration standard}$

As before, the equation above scales the N2O:Ar ratio measured in the
sample to that of the standard. Via substitution and rearrangement, the
equation can be written as:
$$C^{N_2O}_{\theta} = \frac{R^{X}_{\theta}}{R^{X}_s} \cdot R^{C}_s \cdot C^{Ar}_s$$
where:

-   $R^{X}_\theta = \text{Signal ratio of } \text{N}_2\text{O} \text{ to Ar in the sample}$

Concentration of N2O and Ar in the standard, $C^{N_2O}_s$ is calculated from
temperature of the equilibrated water, $Tc$, laboratory
barometric pressure, $B$, and N2O and Ar partial pressure in the atmosphere, 
$C^{N_2O}_a$, $C^{Ar}_a$, respectively (see [Equilibrium concentration]). By 
substitution, the equation to estimate the concentration of N2O in the unknown
sample becomes:

$$\begin{align}
C^{N_2O}_{\theta} = \frac{R^{X}_{\theta}}{R^{X}_s} \times \\ \frac{ 1^{-6} \cdot B \cdot C^{N_2O}_a \cdot (H_{N_2O} \cdot e^{2700 \cdot \frac{1}{Tc} -  \frac{1}{298.15}})}{1^{-6} \cdot B \cdot C^{Ar}_a \cdot (H_{Ar} \cdot e^{1500 \cdot \frac{1}{Tc} -  \frac{1}{298.15}})} \times \\ 1^{-6}B \cdot C^{Ar}_a \cdot (H_{Ar} \cdot e^{1500 \cdot  \frac{1}{Tc} -  \frac{1}{298.15}})
\end{align}$$

where (repeating some variables from above for clarity):

-   $C^{N_2O}_{\theta} = \text{concentration of } \text{N}_2\text{O} \text{ in the sample}$
-   $R^{X}_\theta = \text{Signal ratio of } \text{N}_2\text{O} \text{ to Ar in the sample}$
-   $R^{X}_s = \text{Signal ratio of } \text{N}_2\text{O} \text{ to Ar in calibration standard}$
-   $B = \text{barometric pressure (kPa)}$
-   $Tc = \text{temperature}$
-   $C^{N_2O}_a = \text{concentration of } \text{N}_2\text{O} \text{ in the atmosphere}$
-   $H_{N_2O} = \text{Henry's law constant for } \text{N}_2\text{O}$
-   $C^{Ar}_a = \text{concentration of } Ar \text{ in the atmosphere}$
-   $H_{Ar} = \text{Henry's law constant for } Ar$
-   $Tc = \text{Temperature of the equilibrated water}$

N2O concentration in laboratory air, ($C^{N_2O}_a$), is derived from direct
measurement of N2O via gas chromatography and has a standard deviation
of `r n2o_sd_epa` using the method employed in this study (see "Empirical 
variation" under [Dissolved concentration]) although more precise methods are
available (see [Dissolved $\text{N}_2\text{O}$: high precision GC]). Ar in the lab 
atmosphere ($C^{Ar}_a$) is 9340ppm and is assumed to be constant and perfectly 
known (e.g. zero error). Genther et al (2013) report a standard deviation of
0.022 for the N2O:Ar signal strength ratio but Speir et al (2023) report a CV 
of 0.05% (sd \~ 0.000096). Here we provisionally use the Speir value pending 
confirmation. MIMS uses a very precise thermometer (+/-0.01 C) to control the 
temperature of the water used as the analytical standard (simulated below).
Measurement error in $B$ was quantified above.

## Dissolved concentration

### Standard GC

```{r simulate_me_hpt_MIMS, message=F}
#n2o_ar_standard_signal <- 0.27 # Speir et al. 0.27 pers comm. Genther et al. section 3.4 reports 0.192
#n2o_ar_sample_signal <- 0.23 # just a guess
#C_a_ar <- 9340 # ppm.  invariant
#T_c_mims <- T_c # set up for simulation of MIMS temperature bath (from GC simulations above)

R_X_s <- 0.27 # Speir et al. 0.27 pers comm. Genther et al. section 3.4 reports 0.192 (signal ratio of N2O:Ar in standard)
R_X_theta <- 0.23 # just a guess (signal ratio of N2O:Ar in sample)
C_Ar_a <- 9340 # ppm.  invariant (Concentration of Ar in atmospher)
# T_c_mims <- T_c # set up for simulation of MIMS temperature bath (from GC simulations above)

# this simulates MIMS measurement.
# assume TRUE value is n2o_ar_standard_signal and error is simulated with CV from Speier et al. 2023
# Speier et al reports "...our precision was good among all sample and standard replicated (average CV = 0.05%)."
# Using that CV to model measurement error in the N2O:Ar ratio yields a sd of dissolved N2O that exceeds
# the sd when measured via GC, which is weird.  Speier reported that the CV is for the N2O:Ar ratio.
# Genther et al reports a n2o:Ar sd of 0.022, but this also generated a sd for dissolved N2O which exceeds that 
# measured with GC!  Here we use the Speir data
R_X_s_sd <- 0.0005 * R_X_s # sd = CV * mean. 0.05% = 0.0005
R_X_theta_sd <- 0.0005 * R_X_theta # sd = CV * mean. 0.05% = 0.0005

R_X_s_me <- rnorm(nsim, R_X_s, R_X_s_sd) # me in N2O:Ar of standard  
R_X_theta_me <- rnorm(nsim, R_X_theta, R_X_theta_sd) # me in N2O:Ar of sample

# High precision water batch used to control temperature of standards
T_c_mims_me <- rnorm(nsim, T_c, 0.005) # thermo stated bath precision is +/-0.01 Celsius
# hist(T_c_mims_me) # sd of 0.005 gives approx distribution for the specified precision

# True dissolved N2O concentration
C_mims <- ((R_X_theta / R_X_s) * # N2O:Ar ratios for sample and standard
  ((1e-6 * B * M_a_n2o * (H_n2o * exp( 2700 * (1 / (T_c + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
  (1e-6 * B * C_Ar_a * (H_ar * exp( 1500 * (1 / (T_c + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
  (1e-6 * B * C_Ar_a*(H_ar  * exp(1500 *(1/(T_c + 273.15) - 1/298.15)))))

C_mims <- C_mims * 1e9 #convert mol to nmol

# variation in measured N2O concentration
C_mims_hpt_me <- rep(NA, nsim)

for(i in 1:nsim){
C_mims_hpt_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
  ((1e-6 * B_me[i] * M_a_n2o_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) 
}

# combine into dataframe for ggplot.
C_mims_hpt_df <- tibble(C_mims_hpt_me = C_mims_hpt_me * 1e9)
# save absolute error for plot
C_mims_hpt_me_abs <- (C_mims_hpt_me * 1e9) - C_mims
```

```{r plot_me_hpt_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(C_mims_hpt_df, aes(x = C_mims_hpt_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.01) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C_mims , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in dissolved N2O \nestimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(C_mims_hpt_df, aes(x = C_mims_hpt_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.01) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_me_hpt_MIMS}
# standard deviation of dissolved N2O via MIMS:
sd(C_mims_hpt_me * 1e9) # 0.169 using Speir CV to calculate N2O:Ar sd.   0.15 with GC and headspace equilibrium!

# 1.96 x sd
sd(C_mims_hpt_me * 1e9) * 1.96

# CV of dissolved N2O
(sd(C_mims_hpt_me * 1e9) / mean(C_mims_hpt_me * 1e9)) * 100 # 0.025 = 2.5% vs 12% using Genther sd.
```

```{r calc_me_hpt_MIMS_2}
# standard deviation of a measurement is:
round(sd(C_mims_hpt_me_abs), 3) # 0.20 nmol L-1

# central 95th percentile of absolute error
round(quantile(C_mims_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((C_mims + C_mims_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

### High precision GC

```{r hpgc_M_a_MIMS}
# simulate improved measurement precision of GC measurement of N2O in air
M_a_n2o_lit_me <- rnorm(nsim, M_a_n2o, n2o_sd_lit) # from intn'l labs = 0.00073 vs epa sd = 0.00787
```

```{r simulate_me_hpgc_hpt_MIMS, message=F}
# variation in measured N2O concentration
C_mims_hpgc_hpt_me <- rep(NA, nsim)

for(i in 1:nsim){
C_mims_hpgc_hpt_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
  ((1e-6 * B_me[i] * M_a_n2o_lit_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) 
}

# combine into dataframe for ggplot.
C_mims_hpgc_hpt_df <- tibble(C_mims_hpgc_hpt_me = C_mims_hpgc_hpt_me * 1e9)
# save absolute error for plot
C_mims_hpgc_hpt_me_abs <- (C_mims_hpgc_hpt_me * 1e9) - C_mims
```

```{r plot_me_hpgc_hpt_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(C_mims_hpgc_hpt_df, aes(x = C_mims_hpgc_hpt_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.0025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = C_mims , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in dissolved N2O \nestimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(C_mims_hpgc_hpt_df, aes(x = C_mims_hpgc_hpt_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.0025) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in dissolved \nN2O estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_me_hpgc_hpt_MIMS}
# standard deviation of dissolved N2O via MIMS:
sd(C_mims_hpgc_hpt_me * 1e9) # 0.0166 using Speir CV to calculate N2O:Ar sd. 0.15 with GC and headspace equilibrium!

sd(C_mims_hpgc_hpt_me * 1e9) * 1.96 # sd +/- 1.96 (for conf intervals on error)

# CV of dissolved N2O
sd(C_mims_hpgc_hpt_me * 1e9) / mean(C_mims_hpgc_hpt_me * 1e9) * 100 # 0.0025 = 0.25% CV 
```

```{r calc_me_hpgc_hpt_MIMS_2}
# standard deviation of a measurement is:
round(sd(C_mims_hpgc_hpt_me_abs), 3)

# central 95th percentile of absolute error
round(quantile(C_mims_hpgc_hpt_me_abs, probs = c(0.025, 0.975)), 3)

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((C_mims + C_mims_hpgc_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

## Saturation ratio

Calculate the "true" ratio via MIMS
```{r true_ratio_MIMS, message=F}
# True ratio
S_mims <- ((R_X_theta / R_X_s) * # N2O:Ar ratios for sample and standard
  ( (1e-6 * B * M_a_n2o * (H_n2o * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
  (1e-6 * B * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c + 273.15) - 1 / 298.15)) ))) * #Ar concentration in standard
  (1e-6 * B * C_Ar_a *(H_ar  * exp(1500 * (1 / (T_c + 273.15) - 1 / 298.15))))) / # Ar concentration in standard
  (M_a_n2o * B * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c + 273.15) - 1 / 298.15)))) # N2o saturation value in waterbody 
```

### Standard GC

#### Standard field temperature
This simulation is for field measurment of temperature for the equilibrium concentration with the standard glass thermometer.
```{r simulate_me_ratio_MIMS, message=F}
# ratio simulated with measurement error in mims
S_mims_me <- rep(NA, nsim)

for( i in 1:nsim){
S_mims_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
  ((1e-6 * B_me[i] * M_a_n2o_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 *(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) / # Ar concentration in standard 
  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1/298.15)))) # N2o saturation value in waterbody. Using regular field temp
}

# combine into dataframe for ggplot.
S_mims_df <- tibble(S_mims_me = S_mims_me)
# save absolute error for plot
S_mims_me_abs <- S_mims_me - S_mims
```

```{r plot_me_ratio_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(S_mims_df, aes(x = S_mims_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.0001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = S_mims , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in saturation \nratio estimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(S_mims_df, aes(x = S_mims_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.0001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_me_ratio_MIMS}
# standard deviation of a saturation ratio value s:
sd(S_mims_me) # 0.0011

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(S_mims_me) * 1.96 # +/- 0.00216

# so any ratio within 0.00216 of 1.0 cannot be discerned as a source or sink

# CV of sat
sd(S_mims_me) / mean(S_mims_me) * 100 # 0.13% CV 
```

```{r calc_me_ratio_MIMS_2}
# standard deviation of a measurement is:
round(sd(S_mims_me_abs), 3) # 0.20 nmol L-1

# central 95th percentile of absolute error
round(quantile(S_mims_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((S_mims + S_mims_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O 
```

##### Map source-sink status

```{r map_me_ratio_MIMS, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

#dg <- dg %>%
#  mutate(n2o.src.snk.with.all.error = case_when(
#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
#    TRUE ~ n2o.src.snk)) 
dg <- dg %>%
  mutate(n2o.src.snk.with.all.error = case_when(
    abs(n2o.sat.ratio - 1) < quantile(S_mims_me_abs, probs = c(0.975)) ~ "undetermined",
    TRUE ~ n2o.src.snk))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.all.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

# sf for plotting
dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting
```

```{r plot_map_me_ratio_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
## load ecoregions----
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple",
                quiet = TRUE)

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# SET UP CUSTOM COLORS FOR ECOREGIONS
# Custom color pallette for ecoregion polygons. Attempted to mirror
# https://www.epa.gov/national-aquatic-resource-surveys/
# ecoregional-results-national-lakes-assessment-2012
cols <- c("Coastal Plains" = "orange1",
          "Northern Appalachians" = "lightpink1",
          "Northern Plains" = "darksalmon",
          "Southern Appalachians" = "mediumturquoise",
          "Southern Plains" = "khaki4",
          "Temperate Plains" = "forestgreen", 
          "Upper Midwest" = "deepskyblue4",
          "Western Mountains" = "saddlebrown",
          "Xeric" = "lightskyblue4")

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

# To enable spatial analysis of the data, the dataframe will be converted to a 
# 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
            shape = 1,
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", 
                      values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.5, 8), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

#### High precision temperature

```{r simulate_me_ratio_hpt_MIMS, message=F}
# ratio simulated with measurement error in mims
S_mims_hpt_me <- rep(NA, nsim)

for( i in 1:nsim){
S_mims_hpt_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
  ((1e-6 * B_me[i] * M_a_n2o_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 *(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) / # Ar concentration in standard 
  (M_a_n2o_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1/298.15)))) # N2o saturation value in waterbody. Using regular field temp
}

# combine into dataframe for ggplot.
S_mims_hpt_df <- tibble(S_mims_hpt_me = S_mims_hpt_me)
# save absolute error for plot
S_mims_hpt_me_abs <- S_mims_hpt_me - S_mims
```

```{r plot_me_ratio_hpt_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(S_mims_hpt_df, aes(x = S_mims_hpt_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.0001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = S_mims , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in saturation \nratio estimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(S_mims_hpt_df, aes(x = S_mims_hpt_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.0001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_me_ratio_hpt_MIMS}
# standard deviation of a saturation ratio value s:
sd(S_mims_hpt_me) # 0.0011

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(S_mims_hpt_me) * 1.96 # +/- 0.00216

# so any ratio within 0.00216 of 1.0 cannot be discerned as a source or sink

# CV of sat
sd(S_mims_hpt_me) / mean(S_mims_hpt_me) * 100 # 0.13% CV 
```

```{r calc_me_ratio_hpt_MIMS_2}
# standard deviation of a measurement is:
round(sd(S_mims_hpt_me_abs), 3) # 0.20 nmol L-1

# central 95th percentile of absolute error
round(quantile(S_mims_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((S_mims + S_mims_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O 
```

##### Map source-sink status

```{r map_me_ratio_hpt_MIMS, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

#dg <- dg %>%
#  mutate(n2o.src.snk.with.all.error = case_when(
#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
#    TRUE ~ n2o.src.snk)) 
dg <- dg %>%
  mutate(n2o.src.snk.with.all.error = case_when(
    abs(n2o.sat.ratio - 1) < quantile(S_mims_hpt_me_abs, probs = c(0.975)) ~ "undetermined",
    TRUE ~ n2o.src.snk))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.all.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

# sf for plotting
dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting
```

```{r plot_map_me_ratio_hpt_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
## load ecoregions----
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple",
                quiet = TRUE)

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# SET UP CUSTOM COLORS FOR ECOREGIONS
# Custom color pallette for ecoregion polygons. Attempted to mirror
# https://www.epa.gov/national-aquatic-resource-surveys/
# ecoregional-results-national-lakes-assessment-2012
cols <- c("Coastal Plains" = "orange1",
          "Northern Appalachians" = "lightpink1",
          "Northern Plains" = "darksalmon",
          "Southern Appalachians" = "mediumturquoise",
          "Southern Plains" = "khaki4",
          "Temperate Plains" = "forestgreen", 
          "Upper Midwest" = "deepskyblue4",
          "Western Mountains" = "saddlebrown",
          "Xeric" = "lightskyblue4")

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

# To enable spatial analysis of the data, the dataframe will be converted to a 
# 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
            shape = 1,
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", 
                      values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.5, 8), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

### High precision GC

#### Standard temperature

This simulation is for field measurment of temperature for the equilibrium concentration with the standard glass thermometer.
```{r simulate_me_ratio_hpgc_MIMS, message=F}
S_mims_hpgc_me <- rep(NA, nsim)

for( i in 1:nsim){
S_mims_hpgc_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
  ((1e-6 * B_me[i] * M_a_n2o_lit_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 *(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) / # Ar concentration in standard 
  (M_a_n2o_lit_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_me[i] + 273.15) - 1/298.15)))) # N2o saturation value in waterbody. Using regular field temp
}

# combine into dataframe for ggplot.
S_mims_hpgc_df <- tibble(S_mims_hpgc_me = S_mims_hpgc_me)
# save absolute error for plot
S_mims_hpgc_me_abs <- S_mims_hpgc_me - S_mims
```

```{r plot_me_ratio_hpgc_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(S_mims_hpgc_df, aes(x = S_mims_hpgc_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.0001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = S_mims , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in saturation \nratio estimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(S_mims_hpgc_df, aes(x = S_mims_hpgc_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.0001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_me_ratio_hpgc_MIMS}
# standard deviation of a saturation ratio value s:
sd(S_mims_hpgc_me) # 0.0011

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(S_mims_hpgc_me) * 1.96 # +/- 0.00216

# so any ratio within 0.00216 of 1.0 cannot be discerned as a source or sink

# CV of sat
sd(S_mims_hpgc_me) / mean(S_mims_hpgc_me) * 100 # 0.07% CV 
```

```{r calc_me_ratio_hpgc_MIMS_2}
# standard deviation of a measurement is:
round(sd(S_mims_hpgc_me_abs), 3) # 0.20 nmol L-1

# central 95th percentile of absolute error
round(quantile(S_mims_hpgc_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((S_mims + S_mims_hpgc_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

##### Map source-sink status

```{r map_me_ratio_hpgc_MIMS, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

#dg <- dg %>%
#  mutate(n2o.src.snk.with.all.error = case_when(
#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
#    TRUE ~ n2o.src.snk)) 
dg <- dg %>%
  mutate(n2o.src.snk.with.all.error = case_when(
    abs(n2o.sat.ratio - 1) < quantile(S_mims_hpgc_me_abs, probs = c(0.975)) ~ "undetermined",
    TRUE ~ n2o.src.snk))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.all.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

# sf for plotting
dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting
```

```{r plot_map_me_ratio_hpgc_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
## load ecoregions----
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple",
                quiet = TRUE)

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# SET UP CUSTOM COLORS FOR ECOREGIONS
# Custom color pallette for ecoregion polygons. Attempted to mirror
# https://www.epa.gov/national-aquatic-resource-surveys/
# ecoregional-results-national-lakes-assessment-2012
cols <- c("Coastal Plains" = "orange1",
          "Northern Appalachians" = "lightpink1",
          "Northern Plains" = "darksalmon",
          "Southern Appalachians" = "mediumturquoise",
          "Southern Plains" = "khaki4",
          "Temperate Plains" = "forestgreen", 
          "Upper Midwest" = "deepskyblue4",
          "Western Mountains" = "saddlebrown",
          "Xeric" = "lightskyblue4")

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

# To enable spatial analysis of the data, the dataframe will be converted to a 
# 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
            shape = 1,
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", 
                      values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.5, 8), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

#### High precision temperature

The thermometers used to measure water temperature in this study have an
uncertainty of +/- 0.058C but electronic thermisters are commonly
available (e.g. YSI sondes) with an uncertainty of +/-0.01 Celsius. More
precise water temperature measurements will reduce uncertainty in the
equilibrium N2O conentration and the saturation ratio.

```{r simulate_me_ratio_hpgc_hpt_MIMS, message=F}
S_mims_hpgc_hpt_me <- rep(NA, nsim)

for( i in 1:nsim){
S_mims_hpgc_hpt_me[i] <- (R_X_theta_me[i] / R_X_s_me[i]) * # N2O:Ar ratios for sample and standard
  ((1e-6 * B_me[i] * M_a_n2o_lit_me[i] * (H_n2o * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15)))) / #N2O concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar * exp(1500 * (1 / (T_c_mims_me[i] + 273.15) - 1 / 298.15))))) * #Ar concentration in standard
  (1e-6 * B_me[i] * C_Ar_a * (H_ar  * exp(1500 *(1/(T_c_mims_me[i] + 273.15) - 1/298.15)))) / # Ar concentration in standard 
  (M_a_n2o_lit_me[i] * B_me[i] * 1e-6 * (H_n2o  * exp(2700 * (1 / (T_c_mims_me[i] + 273.15) - 1/298.15)))) # N2o saturation value in waterbody. Using high precision field temp
}

# combine into dataframe for ggplot.
S_mims_hpgc_hpt_df <- tibble(S_mims_hpgc_hpt_me = S_mims_hpgc_hpt_me)

# save absolute error for plot
S_mims_hpgc_hpt_me_abs <- S_mims_hpgc_hpt_me - S_mims
```

```{r plot_me_ratio_hpgc_hpt_MIMS, echo=FALSE, fig.align='center', fig.asp=0.45, fig.width=6}
ggarrange(
ggplot(S_mims_hpgc_hpt_df, aes(x = S_mims_hpgc_hpt_me, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.0001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline(xintercept = S_mims , color = 'red') + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) +
  ggtitle( "Uncertainty in saturation \nratio estimate \n") +
  xlab( "Concentration (nMol)"),

ggplot(S_mims_hpgc_hpt_df, aes(x = S_mims_hpgc_hpt_me_abs, y = after_stat(density))) +  
  geom_histogram( binwidth = 0.0001) + 
  geom_density(aes(y = after_stat(density)), color = 'green') +
  geom_vline( xintercept = 0 , color = 'red' ) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()) + 
  ggtitle( "Absolute error in saturation \nratio estimate (observed - true) \n") +
  xlab( "Error (nMol)"),
nrow = 1
)
```

```{r calc_me_ratio_hpgc_hpt_MIMS}
# standard deviation of a saturation ratio value s:
sd(S_mims_hpgc_hpt_me) # 0.0006603624

# the 95% CI is therefore the computed value +/- 2 standard deviations
sd(S_mims_hpgc_hpt_me) * 1.96 # +/- 0.00129431

# so any ratio within 0.00129 of 1.0 cannot be discerned as a source or sink
```

```{r calc_me_ratio_hpgc_hpt_MIMS_2}
# standard deviation of a measurement is:
round(sd(S_mims_hpgc_hpt_me_abs), 3) # 0.20 nmol L-1

# central 95th percentile of absolute error
round(quantile(S_mims_hpgc_hpt_me_abs, probs = c(0.025, 0.975)), 3) # approx +/- 0.11

# the 95% CI for an observation with measurement error can be computed 
# approximately as the value +/- 2 standard deviations; or the central 95th 
# percentile of the absolute error distribution.
round(quantile((S_mims + S_mims_hpgc_hpt_me_abs), probs = c(0.025, 0.975)), 3) # +/- nmol N2O
```

##### Map source-sink status

```{r map_me_ratio_hpgc_hpt_MIMS, message=F}
# two standard deviations = 0.3019449
# 95% CI is therefore x+/-0.3019449

#dg <- dg %>%
#  mutate(n2o.src.snk.with.all.error = case_when(
#    abs(n2o.sat.ratio - 1) < sd(C_me_ratio) ~ "undetermined",
#    TRUE ~ n2o.src.snk)) 
dg <- dg %>%
  mutate(n2o.src.snk.with.all.error = case_when(
    abs(n2o.sat.ratio - 1) < quantile(S_mims_hpgc_hpt_me_abs, probs = c(0.975)) ~ "undetermined",
    TRUE ~ n2o.src.snk))

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {table(.$n2o.src.snk.with.all.error)} %>%
  kable(digits = 2, caption = "Number of lakes classified as source/sink")

dg %>% 
  filter(sample.source == "DG", visit.no == 1, sitetype == "PROB") %>%
  {prop.table(table(.$n2o.src.snk.with.all.error))} %>%
  kable(digits = 2, caption = "Proportion of lakes classified as source/sink")

# sf for plotting
dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting
```

```{r plot_map_me_ratio_hpgc_hpt_MIMS, echo=FALSE, fig.align='center', fig.width=8, fig.asp=0.6}
## load ecoregions----
# read in ecoregion polygons
ecoR <- st_read(dsn = paste0(localPath, 
                             "/Environmental Protection Agency (EPA)/",
                             "ORD NLA17 Dissolved Gas - Documents/inputData"),
                layer = "aggr_ecoregions_simple",
                quiet = TRUE)

# Check CRS
#st_crs(ecoR) # 3857
ecoR <- st_transform(ecoR, 5070) # convert to CONUS Albers
#st_crs(ecoR) # 5070

# SET UP CUSTOM COLORS FOR ECOREGIONS
# Custom color pallette for ecoregion polygons. Attempted to mirror
# https://www.epa.gov/national-aquatic-resource-surveys/
# ecoregional-results-national-lakes-assessment-2012
cols <- c("Coastal Plains" = "orange1",
          "Northern Appalachians" = "lightpink1",
          "Northern Plains" = "darksalmon",
          "Southern Appalachians" = "mediumturquoise",
          "Southern Plains" = "khaki4",
          "Temperate Plains" = "forestgreen", 
          "Upper Midwest" = "deepskyblue4",
          "Western Mountains" = "saddlebrown",
          "Xeric" = "lightskyblue4")

## Load state boundaries----
states <- USAboundaries::us_states() %>%
  dplyr::filter(!state_name %in% c("Alaska", "District of Columbia", "Hawaii", "Puerto Rico")) %>%
  st_transform(5070) # convert to CONUS Albers

# To enable spatial analysis of the data, the dataframe will be converted to a 
# 'simple features' (sf) object.
# Define coordinates
coords <- data.frame(longitude = dg$map.lon.dd, latitude = dg$map.lat.dd)

dg.sf <- st_as_sf(dg, coords = c("map.lon.dd", "map.lat.dd"), 
                  crs = 4269) %>% # standard for lat/lon
  st_transform(5070) # project to CONUS ALBERS for plotting

## Map
ggplot() +
    geom_sf(data = ecoR, color = NA, aes(fill = WSA9_NAME), alpha = 0.5) +
    geom_sf(data = dg.sf %>% 
              filter(!is.na(n2o.src.snk), sitetype == "PROB", visit.no == 1) %>%
              arrange(desc(n2o.src.snk.with.all.error)), # grey on top of black
            aes(size = dissolved.n2o.nmol, color = n2o.src.snk.with.all.error),
            shape = 1,
            show.legend = "point") +
    geom_sf(data = states, fill = NA, color = "cornsilk3", size = 0.1) +
    # guide argument below removes points from the boxes in the ecoregion legend
    # https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/
    scale_fill_manual("Ecoregion", 
                      values = cols,
                      guide = guide_legend(override.aes = list(shape = NA))) +
    scale_color_manual(values = c("#ca0020", "#0571b0", "#000000"), name = "source/sink") +
    scale_size(name = expression(N[2]*O~(nM)),
               range = c(0.5, 8), # custom size range
               breaks = c(1, 10, 25, 50, 100)) + # custom breaks
    theme(legend.key.size = unit(0.4, "cm"), # size of boxes in legend
          legend.title = element_text(size = 8))
```

# Summary of ratio results

Precision of the N2O saturation ratio can be improved using high
precision GC analytics and N2O-free headspace gas (e.g. helium,
nitrogen) but the greatest improvement is seen when dissolved N2O is
directly measured using MIMS.

```{r final_summary, message=FALSE}
#results <- tibble(ratio_sd = c(0.055,
#                               0.048,
#                               0.029,
#                               0.001207747,
#                               0.001207747,
#                               0.000690139),
#                  method = c("GC",
#                             "GC + helium",
#                             "High precision GC + helium",
#                             "MIMS",
#                             "MIMS + high precision GC",
#                             "MIMS + high precision GC + \nhigh precision water temp"))

#results %>%
#  mutate(lower.95 = C_ratio - (1.96 * ratio_sd),
#         upper.95 = C_ratio + (1.96 * ratio_sd),
#         C_ratio = C_ratio) %>%
#  ggplot(aes(C_ratio, method)) +
#  geom_point() +
#  geom_segment(aes(x=lower.95, xend = upper.95, y=method, yend=method)) +
#  xlab("True saturation ratio (black dot) +/- measurement error (1.96 standard deviations))") +
#  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust = 1))

results <- tibble(GC_me= S_a_me_abs,
                  GC_He_me = S_pg_me_abs,
                  GC_hpt_me = S_a_hpt_me_abs,
                  GC_He_hpt_me = S_pg_hpt_me_abs,
                  HPGC_He_me = S_pg_hpgc_me_abs,
                  HPGC_He_hpt_me = S_pg_hpgc_hpt_me_abs,
                  MIMS_me = S_mims_me_abs,
                  MIMS_hpt_me = S_mims_hpt_me_abs,
                  MIMS_HPGC_me = S_mims_hpgc_me_abs,
                  MIMS_HPGC_HPT_me = S_mims_hpgc_hpt_me_abs) %>%
  pivot_longer(cols = ends_with("me"), names_to = "Method", values_to = "Error")

results %>%
  mutate(Method = factor(Method)) %>%
  mutate(Method = fct_relevel(Method, c("MIMS_HPGC_HPT_me",
                                        "MIMS_HPGC_me",
                                        "MIMS_hpt_me",
                                        "MIMS_me",
                                        "HPGC_He_hpt_me",
                                        "HPGC_He_me",
                                        "GC_He_hpt_me",
                                        "GC_hpt_me",
                                        "GC_He_me",
                                        "GC_me"
                                        ))) %>%
  ggplot(aes(x = Error, y = Method)) +
  stat_pointinterval() +
  scale_y_discrete(labels=c("GC_me" = "Standard GC, ambient air, \n standard field temperature",
                            "GC_He_me" = "Standard GC, helium, \n standard field temperature",
                            "GC_hpt_me" = "Standard GC, ambient air, \n high precision temperature",
                            "GC_He_hpt_me" = "Standard GC, helium, \nhigh precision temperature",
                            "HPGC_He_me" = "High precision GC, helium, \nstandard field temperature",
                            "HPGC_He_hpt_me" = "High precision GC, helium, \nhigh precision temperature",
                            "MIMS_me" = "MIMS: standard GC, \nstandard field temperature",
                            "MIMS_hpt_me" = "MIMS: standard GC, \nhigh precision temperature",
                            "MIMS_HPGC_me" = "MIMS: high precision GC, \nstandard field temperature",
                            "MIMS_HPGC_HPT_me" = "MIMS: high precision GC, \nhigh precision temperature")) +
  xlab("Absolute error in saturation ratio") +
  ylab("")
```

Because the MIMS method is so much more precise than the traditional method, 
the differences between the MIMS-based variations above are difficult to see. 
Below, just the MIMS-based variations are compared.

Precision of the N2O saturation ratio can be improved using high
precision GC analytics and N2O-free headspace gas (e.g. helium,
nitrogen) but the greatest improvement is seen when dissolved N2O is
directly measured using MIMS.

```{r final_summary_MIMS, message=FALSE}
results_MIMS <- tibble(MIMS_me = S_mims_me_abs,
                       MIMS_hpt_me = S_mims_hpt_me_abs,
                       MIMS_HPGC_me = S_mims_hpgc_me_abs,
                       MIMS_HPGC_HPT_me = S_mims_hpgc_hpt_me_abs) %>%
  pivot_longer(cols = ends_with("me"), names_to = "Method", values_to = "Error")

results_MIMS %>%
  mutate(Method = factor(Method)) %>%
  mutate(Method = fct_relevel(Method, c("MIMS_HPGC_HPT_me",
                                        "MIMS_HPGC_me",
                                        "MIMS_hpt_me",
                                        "MIMS_me"
                                        ))) %>%
  ggplot(aes(x = Error, y = Method)) +
  stat_pointinterval() +
  scale_y_discrete(labels=c("MIMS_me" = "MIMS: standard GC, \nstandard field temperature",
                            "MIMS_hpt_me" = "MIMS: standard GC, \nhigh precision temperature",
                            "MIMS_HPGC_me" = "MIMS: high precision GC, \nstandard field temperature",
                            "MIMS_HPGC_HPT_me" = "MIMS: high precision GC, \nhigh precision temperature")) +
  xlab("Absolute error in saturation ratio") +
  ylab("")
```