### 2025-03-03 scripts/Design_Based_Estimates_N2O.Rmd
Updating repo to include only files and directories needed to generate results and manuscript

---
title: "Design-based estimates for N2O from the NLA 2017 Survey"
author: "Roy Martin, Jake Beaulieu, Michael McManus"
date: "`r Sys.Date()`"
output:
  github_document:
    number_sections: TRUE
    df_print: "tibble"
    math_method: 
      engine: webtex
    #  url: https://latex.codecogs.com/svg.image?
    html_preview: TRUE
    keep_html: TRUE
editor_options:
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

# Purpose
Calculate estimates for the NLA 2017 survey with spsurvey and local mean variance estimator for comparison to Bayesian MRP Estimates.
Tom Kincaid created a .csv file to be used for population estimates, which is loaded and munged below.

# Import and munge survey data
```{r setup_data, echo=FALSE}

# Attach the Hmisc and spsurvey packages
library(tidyverse)
library(Hmisc)
library(spsurvey)

# Setup file path
localPath <- Sys.getenv("USERPROFILE")
```

```{r import_data} 
# modified file to include N2O emission rates.  See "Users\JBEAULIE\OneDrive - Environmental Protection Agency (EPA)\gitRepository\DissolvedGasNla\scripts\dataMunge.Rmd" for details
# updated file path
df_survey <- read.csv(file = "./../inputData/populationEstimates/NLA17_DissolvedGases_forPopEst.csv")
df_survey <- df_survey %>% 
  data.frame() %>%
  filter(!is.na(DISSOLVED_N2O))
nr <- nrow(df_survey) # 984

# Reorder levels for the lake size class (AREA_CAT6) variable
levels(df_survey$AREA_CAT6) <- list(
   "(1,4]" = "(1,4]",
   "(4,10]" = "(4,10]",
   "(10,20]" = "(10,20]",
   "(20,50]" = "(20,50]",
   ">50" = ">50")

head(df_survey)

tail(df_survey)

# Create data frame for spsurvey estimates
dframe <- data.frame(
  siteID = df_survey$SITE_ID,
  N2O_SAT_RATIO=df_survey$N2O_SAT_RATIO,
  DISSOLVED_N2O=df_survey$DISSOLVED_N2O,
  National=rep("National", nr),
  LAKE_ORGN=df_survey$LAKE_ORGN,
  AREA_CAT6=df_survey$AREA_CAT6,
  AG_ECO9_NM=df_survey$AG_ECO9_NM,
  State = df_survey$STRATUM,
  wgt = df_survey$WGT_TP,
  xcoord = df_survey$XCOORD,
  ycoord = df_survey$YCOORD)
```

# Survey estimates with spsurvey package
Estimates made using the local mean variance estimator
```{r spsurvey_ests, echo=FALSE, results="hide", include=TRUE}

# Calculate the estimates
#cat("\n\nCalculate CDF estimates using the local mean variance estimator\n")
if(exists("warn.df")) rm("warn.df")
time.initial <- Sys.time()
Design_estimates <- cont_analysis(dframe = dframe, 
                               vars = c("N2O_SAT_RATIO", "DISSOLVED_N2O"),
                               subpops = c("National", "LAKE_ORGN", "AREA_CAT6", "AG_ECO9_NM", "State"), 
                               weight = "wgt", 
                               xcoord = "xcoord", 
                               ycoord = "ycoord",
                               vartype = "Local") # 30 minutes on VM
Sys.time() - time.initial
```

# Summarize estimates and export
```{r summarize_mean_estimates_dissolved, echo=FALSE}
mean_dissolved <- Design_estimates$Mean %>% 
  tibble() %>%
  filter(Indicator == "DISSOLVED_N2O") %>%
  mutate(Estimate = round(Estimate * 1e9, 2),
         StdError = round(StdError * 1e9, 2),
         MarginofError = round(MarginofError * 1e9, 2),
         LCB95Pct = round(LCB95Pct * 1e9, 2),
         UCB95Pct = round(UCB95Pct * 1e9, 2),
         Subpopulation = recode(Subpopulation, # simpler naming conventions for the size categories
                           "(1,4]" = "min_4" ,
                           "(10,20]" = "10_20",
                           "(20,50]" = "20_50",
                           "(4,10]" = "4_10",
                           ">50" = "50_max")) %>%
  select(!Indicator)

mean_dissolved %>% print()
```

```{r save_to_drive_1, eval=FALSE, include=TRUE}
write.csv(mean_dissolved, file = "./../inputData/populationEstimates/Survey_Ests_N20_dissolved.csv")
```

```{r summarize_mean_estimates_saturation, echo=FALSE}
mean_sat_ratio <- Design_estimates$Mean %>% 
  tibble() %>%
  filter(Indicator == "N2O_SAT_RATIO") %>%
  mutate(Estimate = round(Estimate, 2),
         StdError = round(StdError, 2),
         MarginofError = round(MarginofError, 2),
         LCB95Pct = round(LCB95Pct, 2),
         UCB95Pct = round(UCB95Pct, 2),
         Subpopulation = recode(Subpopulation, # simpler naming conventions for the size categories
                           "(1,4]" = "min_4" ,
                           "(10,20]" = "10_20",
                           "(20,50]" = "20_50",
                           "(4,10]" = "4_10",
                           ">50" = "50_max")) %>%
  select(!Indicator)

mean_sat_ratio %>% print()
```

```{r save_to_drive_2, eval=FALSE, include=TRUE}
write.csv(mean_sat_ratio, file = "./../inputData/populationEstimates/Survey_Ests_N20_sat.csv")
```

# Session Info

```{r session}
sessionInfo()
```